### 2024-04-17 paper/manuscript.qmd
add apaquarto doc

---
title: "Is Phenomenological Control related to Illusion Sensitivity"
# If blank, the running header is the title in upper case.
shorttitle: "Illusion Sensitivity and Phenomenological Control"
# Set names and affiliations.
# It is nice to specify everyone's orcid, if possible.
# There can be only one corresponding author, but declaring one is optional.
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    # Roles are optional. 
    # Select from the CRediT: Contributor Roles Taxonomy https://credit.niso.org/
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    roles:
      - Conceptualization
      - Data curation
      - formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: id1
        name: "University of Sussex"
        # group: Clinical Psychology Program
        department: School of Psychology
        # address: 1234 Capital St.
        # city: Albany
        # region: NY
        # postal-code: 12084-1234
  - name: Ana Neves
    orcid: 0000-0000-0000-0002
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations: 
      - ref: id1
  - name: Zoltan Dienes
    orcid: 0000-0001-7454-3161
    roles:
      - Supervision
      - Writing – review & editing
    affiliations: 
      - ref: id1
author-note:
  status-changes: 
    # Example: [Author name] is now at [affiliation].
    affiliation-change: ~
    # Example: [Author name] is deceased.
    deceased: ~
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Example: This study was registered at X (Identifier Y).
    study-registration: ~
    # Acknowledge and cite data/materials to be shared.
    data-sharing: ~
    # Example: This article is based on data published in [Reference].
    # Example: This article is based on the dissertation completed by [citation].  
    related-report: ~
    # Example: [Author name] has been a paid consultant for Corporation X, which funded this study.
    conflict-of-interest: ~
    # Example: This study was supported by Grant [Grant Number] from [Funding Source].
    financial-support: ~
    # Example: The authors are grateful to [Person] for [Reason].
    gratitude: ~
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: ~
abstract: "Blablablabla blablabla todo todo todo."
# Put as many keywords at you like, separated by commmas (e.g., [reliability, validity, generalizability])
keywords: [keyword1, keyword2, keyword3]
# If true, tables and figures are mingled with the text instead of listed at the end of the document.
floatsintext: true
# Numbered lines (.pdf and .docx only)
numbered-lines: true
# File with references
bibliography: bibliography.bib
# Suppress title page
suppress-title-page: false
# Masks references that appear in the masked-citations list
mask: false
masked-citations:
  - schneider2012cattell
  - schneider2015intelligence
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
---

```{r}
#| label: setup
#| include: false
library(conflicted)
library(tidyverse)
library(flextable)
library(ftExtra)
library(officer)
library(knitr)

conflicts_prefer(dplyr::filter, .quiet = TRUE)
conflicts_prefer(flextable::separate_header, .quiet = TRUE)
```

This is my introductory paragraph. The title will be placed above it automatically. *Do not start with an introductory heading* (e.g., "Introduction"). The title acts as your Level 1 heading for the introduction.

See [here](https://quarto.org/docs/authoring/footnotes-and-citations.html) for instructions on setting up citations and references.

A parenthetical citation requires square brackets [@CameronTrivedi2013]. This reference was in my bibliography file. An in-text citation is done like so:

@CameronTrivedi2013 make some important points ...

See [here](https://wjschne.github.io/apaquarto/writing.html#references) for explanations, examples, and citation features exclusive to apaquarto. For example, apaquarto can automatically handle possessive citations:

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

### 2024-04-17 paper/manuscript.qmd
Written manuscript and references

@@
   - name: Ana Neves
-    orcid: 0000-0000-0000-0002
+    orcid: 0009-0006-0020-7599
     roles:
@@
 
-This is my introductory paragraph.
-The title will be placed above it automatically.
-*Do not start with an introductory heading* (e.g., "Introduction").
-The title acts as your Level 1 heading for the introduction.
+Visual Illusions are an interesting type of stimuli highlighting the ease with which our phenomenological conscious experience can become dissociated from physical reality.
+Their robust and reliable effect makes them useful stimuli to explore how perception is constructed and shaped, and several theoretical models have been put forth to explain how they work.
+In particular, illusions have been recently reframed using a predictive coding account of perception [@notredame2014] in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].
 
-See [here](https://quarto.org/docs/authoring/footnotes-and-citations.html) for instructions on setting up citations and references.
+Such computational model(s) propose to conceptualize illusions as stimuli providing weak or conflicting sensory evidence [@sundareswara2008; @gershman2012] that bias perception toward prior knowledge.
+In other words, the weight of priors, in the form of perceptual knowledge about the world (e.g., internalized rules of perspective) is amplified when the sensory input is confusing.
+For instance, in the Müller-Lyer illusion, we “compute” the two (actually identical) lines as being of different lengths because the line flanked with converging fins is misinterpreted as being further away [@notredame2014].
+In this context, measuring sensitivity to illusion can be operationalized as indexing the parameters of the Bayesian inference process (e.g., prior precision).
 
-A parenthetical citation requires square brackets [@CameronTrivedi2013]. This reference was in my bibliography file. An in-text citation is done like so:
+These accounts also provide a compelling framework to explain existing findings reporting interindividual variability in the sensitivity to illusions.
+Indeed, several studies suggest a relationship between psychopathology and lower illusion sensitivity.
+Schizophrenic patients, for instance, constantly exhibit lower scores on visual illusion tasks [@costa2023].
+Proposed explanations, under this framework, emphasise alterations in metacognitive components whereby greater confidence is given to sensory evidence over prior knowledge, which show as reduced susceptibility to the effect of illusions [@notredame2014].
+While evidence for lower visual sensitivity is less established in autistic individuals [@gori2016], reduced biases in the effects of illusions have also been attributed to diminished influence of top-down processes such as prior knowledge [@mitchell2010] and greater emphasis on sensory information [@palmer2017].
+Evidence beyond psychopathology suggests illusion sensitivity variation in the general population, which could be correlated for instance with personality traits such as higher agreeableness and honest-humility [@makowski2023] and open-mindness and flexibility of cognitive control [@shoshina2014].
 
-@CameronTrivedi2013 make some important points ...
+However, the exact nature of this interindividual variability and its potential origin remains unclear.
+The somewhat mixed evidence in the literature regarding its generalizability and strength could be related to the variety of the paradigms used and the type of processes being mobilised [@makowski2023].
+Indeed, traditional methods frequently focus on participant’s experience by prompting them to assess the difference between two identical targets, estimate the target’s physical properties, or adjust the targets to match a reference stimulus [@todorovic2020].
+Relying on metacognitive judgments about one’s subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion sensibility.
+Moreover, paradigms often face challenges in diversifying the illusory effects (i.e., using multiple stimuli to experimentally manipulate the strength of the illusion) and the illusion types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus, Delboeuf which might rely on a different admixture of mechanisms), hindering the potential of obtaining a comprehensive, valid, and reliable measure of illusion sensitivity [@makowski2023].
 
-See [here](https://wjschne.github.io/apaquarto/writing.html#references) for explanations, examples, and citation features exclusive to apaquarto.
-For example, apaquarto can automatically handle possessive citations:
+The “Illusion Game” paradigm [@makowski2023] has been recently developed to measure illusion sensitivity to various illusion types through its behavioural impact (on response time and error rate) in a perceptual decision task (where participants have to respond as fast as possible; e.g., “which of the left or right circles is bigger”).
+The stimuli for different classical illusions are created using the Pyllusion software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
+This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its interference effect, hopefully bypassing some of the metacognitive processes at stake in other paradigms.
+
+Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by metacognitive abilities difference, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and “resist” the illusion in a top-town fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
+If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants’ performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one’s ability to alter one’s subjective experience following suggestions - a mechanism referred to as “phenomenological control”.
+
+The idea that we are endowed with the potential to unconsciously alter our subjective experience and distort reality - even momentarily - to meet the goals at hand is not novel.
+While this phenomenon has been historically often studied under the label of "hypnotisability" - the tendency to alter our conscious experience to match external demands [@lush2021], the term "phenomenological control" (PC) has been recently introduced to disconnect this concept from the potentially negative associations with hypnosis and the misconception that a hypnotic context is necessary for responding to imaginative suggestions [@dienes2022].
+
+To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH**,** @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
+This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has been demonstrated to be compatible with online experiments [@lush2022].
+
+Interestingly, @lush2022 did test for a relationship between PC and illusion sensitivity using the Müller-Lyer illusion (in which the arrangement of the arrowheads flanking two lines makes them appear as having different lengths), and reported evidence in favour of an absence of correlation between the two measures.
+This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types.
+
+# **Methods**
+
+## **Participants**
+
+We aim to recruit around 500 [in line with the sample sizes used in @lush2021; @lush2022] adult English native speakers with a desktop device using Prolific (www.prolific.co).
+Participants will be first presented with an explanatory statement and the consent form, and can proceed by pressing a button to confirm they have read and understood the information.
+This study will be approved by the ethics board of the School of Psychology of the University of Sussex (XX/XXXX).
+
+## **Procedure**
+
+The experiment’s setup follows the born-open principle [@deleeuw2023].
+The online experiment, implemented entirely using JsPsych [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
+Participants’ raw data files are then automatically stored on a European server on OSF; and the deanonymized and ready-to-use data (downloaded from OSF and formatted using a reproducible script) will be made available in the same GitHub repository, alongside the scripts used for the analysis.
+
+Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity).
+Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
+
+**Phenomenological Control Scale (PCS)**.
+Participants will be asked to put on their headphones and await further auditory instructions.
+The PCS procedure starts with a recorded introduction explaining that a series of tests will be applied to evaluate how experiences can be created through imagination.
+This will be followed by 10 suggestions in a fixed order [see @lush2021], such as “now extend your arms ahead of you, with palms facing each other, hands about a foot apart” and “as you sit comfortably in your chair with your eyes closed, a picture of two balls will be displayed on the computer screen”.
+Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale.
+Phenomenological control will be indexed by averaging the scores from the 10 scales.
+
+**Illusion Game**.
+The task is an adaptation of the one used in @makowski2023 to make it shorter and more reliable, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
+It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer, and Vertical-Horizontal.
+The procedure encompasses 2 sets of 80 trials for each illusion type.
+Each set will include, in a random order, the 3 blocks of illusion types, in which trials are separated by a fixation cross, temporally (uniformly sampled duration of 500 - 1000s) and spatially jittered (around the centre of the screen in a radius of a 1 cm) to attenuate its potential usefulness as a reference point.
+After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one’s abilities.
+
+For each illusion type, two continuous dimensions are orthogonally manipulated [see @makowski2021 for details on the rationale and execution], namely task difficulty and illusion strength, so that each trial corresponds to a unique combination.
+Task difficulty corresponds to the difficulty of the perceptual decision (e.g., if the task is to select the longest red line, task difficulty corresponds to how the lines are objectively different).
+Illusion strength corresponds to the degree to which the illusion elements (e.g., the black arrow lines in Müller-Lyer) are interfering with the aforementioned task.
+Note that the illusion effect can be “incongruent” (biasing perception in the direction of the incorrect response) or “congruent” (facilitating, i.e., biasing perception in the direction of the correct response).
+Participants respond with a key arrow (left vs.
+right; or up vs.
+down), and their reaction time (RT) and accuracy are recorded.
+
+Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, separately for the 3 illusion types.
+Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy tradeoff), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current preregistration.
+
+The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.
+
+## **Data Analysis**
+
+The PCS will contain several manipulation check indices to identify problematic participants.
+Participants should not answer "no balls were presented" when queried about the colour they observed on the screen following the negative visual hallucination suggestion, and should execute at least 5 space presses upon instructions to do so.
+
+Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023].
+If the RT is collapsed to the left (i.e., has \> ⅓ of ultra-fast responses - typically \< 200 ms) in the first block, the entire participant will be discarded (suggesting that they did not properly do the text), but if only the second block is bad, then only the second block will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
+In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @thériault2024].
+
+After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the BayesFactor package; @BayesFactor]**.** Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity.
+Data analysis will be carried out using R, using tidyverse [@tidyverse] and easystats [@easystats; @parameters; @correlation; @bayestestR; @datawizard].
+The analysis script and additional information are available at https://github.com/RealityBending/IllusionGamePhenomenologicalControl.
 

### 2024-04-22 paper/manuscript.qmd
Deleted unwanted reference to meta-analysis
abstract and format

@@
 ---
-title: "Is Phenomenological Control related to Illusion Sensitivity"
+title: "Testing the Relationship between Phenomenological Control related to Illusion Sensitivity"
 # If blank, the running header is the title in upper case.
@@
     authorship-agreements: ~
-abstract: "Blablablabla blablabla todo todo todo."
+abstract: "
+Visual illusions highlight how easily our conscious experience can be altered with respect to perceptual reality.
+Despite sharing in-principle mechanisms with phenomenological control, i.e., the ability to alter our perceptual experience to match task demands or expectations, research tying the two remains scarce.
+This study aims to replicate and expand Lush et al.
+(2022) reporting an absence of correlation between phenomenological control (measured using the Phenomenological Control Scale) and illusion sensitivity to different illusion types.
+*[N participants were recruited in an online study.
+Results will added in the final version of the manuscript]*.
+"
 # Put as many keywords at you like, separated by commmas (e.g., [reliability, validity, generalizability])
-keywords: [keyword1, keyword2, keyword3]
+keywords: [illusion sensitivity, visual illusions, phenomenological control, suggestibility, hypnotizability]
 # If true, tables and figures are mingled with the text instead of listed at the end of the document.
@@
 mask: false
-masked-citations:
-  - schneider2012cattell
-  - schneider2015intelligence
 # Language options. See https://quarto.org/docs/authoring/language.html
@@
 In other words, the weight of priors, in the form of perceptual knowledge about the world (e.g., internalized rules of perspective) is amplified when the sensory input is confusing.
-For instance, in the Müller-Lyer illusion, we “compute” the two (actually identical) lines as being of different lengths because the line flanked with converging fins is misinterpreted as being further away [@notredame2014].
+For instance, in the Müller-Lyer illusion, we "compute" the two (actually identical) lines as being of different lengths because the line flanked with converging fins is misinterpreted as being further away [@notredame2014].
 In this context, measuring sensitivity to illusion can be operationalized as indexing the parameters of the Bayesian inference process (e.g., prior precision).
@@
 These accounts also provide a compelling framework to explain existing findings reporting interindividual variability in the sensitivity to illusions.
-Indeed, several studies suggest a relationship between psychopathology and lower illusion sensitivity.
-Schizophrenic patients, for instance, constantly exhibit lower scores on visual illusion tasks [@costa2023].
-Proposed explanations, under this framework, emphasise alterations in metacognitive components whereby greater confidence is given to sensory evidence over prior knowledge, which show as reduced susceptibility to the effect of illusions [@notredame2014].
-While evidence for lower visual sensitivity is less established in autistic individuals [@gori2016], reduced biases in the effects of illusions have also been attributed to diminished influence of top-down processes such as prior knowledge [@mitchell2010] and greater emphasis on sensory information [@palmer2017].
-Evidence beyond psychopathology suggests illusion sensitivity variation in the general population, which could be correlated for instance with personality traits such as higher agreeableness and honest-humility [@makowski2023] and open-mindness and flexibility of cognitive control [@shoshina2014].
+Indeed, several studies suggest a potential link with psychopathology, in particular schizophrenia [@costa2023] and autism [@gori2016], in which the reported lower sensitivity to illusions has been attributed to a diminished influence of top-down processes such as prior knowledge [@mitchell2010] and a greater emphasis on (i.e., precision of) sensory information [@palmer2017].
+Evidence beyond psychopathology also suggests variability in the general population, potentially correlated with personality traits such as agreeableness and honest-humility [@makowski2023], as well as cognitive abilities [@shoshina2014].
 
@@
 The somewhat mixed evidence in the literature regarding its generalizability and strength could be related to the variety of the paradigms used and the type of processes being mobilised [@makowski2023].
-Indeed, traditional methods frequently focus on participant’s experience by prompting them to assess the difference between two identical targets, estimate the target’s physical properties, or adjust the targets to match a reference stimulus [@todorovic2020].
-Relying on metacognitive judgments about one’s subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion sensibility.
+Indeed, traditional methods frequently focus on participant's experience by prompting them to assess the difference between two identical targets, estimate the target's physical properties, or adjust the targets to match a reference stimulus [@todorovic2020].
+Relying on metacognitive judgments about one's subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion sensibility.
 Moreover, paradigms often face challenges in diversifying the illusory effects (i.e., using multiple stimuli to experimentally manipulate the strength of the illusion) and the illusion types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus, Delboeuf which might rely on a different admixture of mechanisms), hindering the potential of obtaining a comprehensive, valid, and reliable measure of illusion sensitivity [@makowski2023].
 
-The “Illusion Game” paradigm [@makowski2023] has been recently developed to measure illusion sensitivity to various illusion types through its behavioural impact (on response time and error rate) in a perceptual decision task (where participants have to respond as fast as possible; e.g., “which of the left or right circles is bigger”).
-The stimuli for different classical illusions are created using the Pyllusion software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
+The "Illusion Game" paradigm [@makowski2023] has been recently developed to measure illusion sensitivity to various illusion types through its behavioural impact (on response time and error rate) in a perceptual decision task (where participants have to respond as fast as possible; e.g., "which of the left or right circles is bigger").
+The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
 This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its interference effect, hopefully bypassing some of the metacognitive processes at stake in other paradigms.
 
-Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by metacognitive abilities difference, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and “resist” the illusion in a top-town fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
-If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants’ performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one’s ability to alter one’s subjective experience following suggestions - a mechanism referred to as “phenomenological control”.
+Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by metacognitive abilities difference, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a top-town fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
+If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".
 
@@
 
-To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH**,** @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
+To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
 This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has been demonstrated to be compatible with online experiments [@lush2022].
@@
 
-# **Methods**
+# Methods
 
-## **Participants**
+## Participants
 
@@
 
-## **Procedure**
+## Procedure
 
-The experiment’s setup follows the born-open principle [@deleeuw2023].
-The online experiment, implemented entirely using JsPsych [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
-Participants’ raw data files are then automatically stored on a European server on OSF; and the deanonymized and ready-to-use data (downloaded from OSF and formatted using a reproducible script) will be made available in the same GitHub repository, alongside the scripts used for the analysis.
+The experiment's setup follows the born-open principle [@deleeuw2023].
+The online experiment, implemented entirely using *JsPsych* [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
+Participants' raw data files are then automatically stored on a European server on OSF; and the deanonymized and ready-to-use data (downloaded from OSF and formatted using a reproducible script) will be made available in the same GitHub repository, alongside the scripts used for the analysis.
 
@@
 
-**Phenomenological Control Scale (PCS)**.
+### Phenomenological Control Scale (PCS)
+
 Participants will be asked to put on their headphones and await further auditory instructions.
 The PCS procedure starts with a recorded introduction explaining that a series of tests will be applied to evaluate how experiences can be created through imagination.
-This will be followed by 10 suggestions in a fixed order [see @lush2021], such as “now extend your arms ahead of you, with palms facing each other, hands about a foot apart” and “as you sit comfortably in your chair with your eyes closed, a picture of two balls will be displayed on the computer screen”.
+This will be followed by 10 suggestions in a fixed order [see @lush2021], such as "now extend your arms ahead of you, with palms facing each other, hands about a foot apart" and "as you sit comfortably in your chair with your eyes closed, a picture of two balls will be displayed on the computer screen".
 Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale.
@@
 
-**Illusion Game**.
+### Illusion Game
+
 The task is an adaptation of the one used in @makowski2023 to make it shorter and more reliable, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
@@
 Each set will include, in a random order, the 3 blocks of illusion types, in which trials are separated by a fixation cross, temporally (uniformly sampled duration of 500 - 1000s) and spatially jittered (around the centre of the screen in a radius of a 1 cm) to attenuate its potential usefulness as a reference point.
-After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one’s abilities.
+After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
 
@@
 Illusion strength corresponds to the degree to which the illusion elements (e.g., the black arrow lines in Müller-Lyer) are interfering with the aforementioned task.
-Note that the illusion effect can be “incongruent” (biasing perception in the direction of the incorrect response) or “congruent” (facilitating, i.e., biasing perception in the direction of the correct response).
+Note that the illusion effect can be "incongruent" (biasing perception in the direction of the incorrect response) or "congruent" (facilitating, i.e., biasing perception in the direction of the correct response).
 Participants respond with a key arrow (left vs.
@@
 Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, separately for the 3 illusion types.
-Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy tradeoff), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current preregistration.
+Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy trade off), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current preregistration.
 
@@
 
-## **Data Analysis**
+## Data Analysis
 
@@
 Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023].
-If the RT is collapsed to the left (i.e., has \> ⅓ of ultra-fast responses - typically \< 200 ms) in the first block, the entire participant will be discarded (suggesting that they did not properly do the text), but if only the second block is bad, then only the second block will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
-In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @thériault2024].
+If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first block, the entire participant will be discarded (suggesting that they did not properly do the text), but if only the second block is bad, then only the second block will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
+In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
 
-After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the BayesFactor package; @BayesFactor]**.** Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity.
-Data analysis will be carried out using R, using tidyverse [@tidyverse] and easystats [@easystats; @parameters; @correlation; @bayestestR; @datawizard].
-The analysis script and additional information are available at https://github.com/RealityBending/IllusionGamePhenomenologicalControl.
+After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the *BayesFactor* package; @BayesFactor].
+Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity.
+Data analysis will be carried out using R, using *tidyverse* [@tidyverse] and *easystats* [@easystats; @parameters; @correlation; @bayestestR; @datawizard].
+The analysis script and additional information are available at *https://github.com/RealityBending/IllusionGamePhenomenologicalControl*.
+
+\pagebreak
 

### 2024-04-26 paper/manuscript.qmd
small changes to writing
Table

@@
       - ref: id1
-  - name: Zoltan Dienes
-    orcid: 0000-0001-7454-3161
-    roles:
-      - Supervision
-      - Writing – review & editing
-    affiliations: 
-      - ref: id1
+  # - name: Zoltan Dienes
+  #   orcid: 0000-0001-7454-3161
+  #   roles:
+  #     - Supervision
+  #     - Writing – review & editing
+  #   affiliations: 
+  #     - ref: id1
 author-note:
@@
 *[N participants were recruited in an online study.
-Results will added in the final version of the manuscript]*.
+Results will be added in the final version of the manuscript]*.
 "
@@
   title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
-  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
+  references-meta-analysis:
 format:
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types.
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see table 1).
+
+```{r}
+# Create the dataframe
+df <- data.frame(
+  Question = "Is there a relationship between trait phenomenological control (PC) and visual illusion (VI) sensitivity?",
+  Hypothesis = "Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS",
+  Sampling_plan = "The goal is to recruit around 500 adult English speakers using Prolific.
+This sample size is based on the ones used in Lush et l., 2021 and Lush et al., 2022 that we aim at replicate.",
+  Analysis_Plan = "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to 'medium')",
+  Rationale = "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+BF10 > 3 will be taken as evidence for a relationship between these two measures.",
+  Interpretation = "The hypothesis that VI sensitivity is independent from PC.",
+  Theory = ""
+)
+
+# Print the dataframe as a table
+print(kable(df), format = "markdown")
+
+```
+
+Table 1. Study Design Table
 

### 2024-04-26 paper/manuscript.qmd
preprint

--- 2024-04-22 paper/manuscript.qmd
@@
     data-sharing: ~
-    # Example: This article is based on data published in [Reference].
     # Example: This article is based on the dissertation completed by [citation].  
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types.
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see @tbl-table1).
+
+::: {#tbl-table1}
+
+| **Question**                                                                                                                                 | **Hypothesis**                                                                                                                            | **Sampling Plan**                                                                                                               | **Analysis Plan**                                                                                                                                                                             | **Rationale for Deciding the Sensitivity of the Test**                                                                                           | **Interpretation Given Different Outcomes**                         | **Theory That Could Be Shown Wrong by the Outcomes**                                                                                                        |
+|----------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------|
+| Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?
+| Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS         | The goal is to recruit around 500 adult English speakers using Prolific.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.
+| Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+BF10 > 3 will be interpreted as evidence for a relationship between these two measures.
+| The hypothesis that VI sensitivity is independent from PC.
+| The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
+|
+
+Study Design Table
+:::
 
@@
 Participants will be first presented with an explanatory statement and the consent form, and can proceed by pressing a button to confirm they have read and understood the information.
-This study will be approved by the ethics board of the School of Psychology of the University of Sussex (XX/XXXX).
+This study has been approved by the ethics board of the School of Psychology of the University of Sussex (ER/ASF25/5).
 
@@
 
+# Results
+
+*This section will be completed after data is collected.*
+
+# Discussion
+
+*This section will be completed after data is collected.*
+
+# Data Availability
+
+All the study materials, experiment, data, and analysis is available on GitHub at https://github.com/RealityBending/IllusionGamePhenomenologicalControl
+
+# Acknowledgments
+
+We would like to thank An Shu Te for her help in setting up the project, Ryan Scott for his help in implementing the phenomenological control scale, and Zoltan Dienes for his input, feedback and guidance.
+
 \pagebreak

### 2024-04-26 paper/manuscript.qmd
cover letter

@@
 However, the exact nature of this interindividual variability and its potential origin remains unclear.
-The somewhat mixed evidence in the literature regarding its generalizability and strength could be related to the variety of the paradigms used and the type of processes being mobilised [@makowski2023].
+The somewhat mixed evidence in the literature regarding its generalizability and strength could be related to the variety of the paradigms used and the type of processes being mobilised [@makowski2021].
 Indeed, traditional methods frequently focus on participant's experience by prompting them to assess the difference between two identical targets, estimate the target's physical properties, or adjust the targets to match a reference stimulus [@todorovic2020].
 Relying on metacognitive judgments about one's subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion sensibility.
-Moreover, paradigms often face challenges in diversifying the illusory effects (i.e., using multiple stimuli to experimentally manipulate the strength of the illusion) and the illusion types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus, Delboeuf which might rely on a different admixture of mechanisms), hindering the potential of obtaining a comprehensive, valid, and reliable measure of illusion sensitivity [@makowski2023].
+Moreover, paradigms often face challenges in diversifying the illusory effects (i.e., using multiple stimuli to experimentally manipulate the strength of the illusion) and the illusion types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus, Delboeuf which might rely on a different admixture of mechanisms), hindering the potential of obtaining a comprehensive, valid, and reliable measure of illusion sensitivity.
 

### 2024-06-04 paper/manuscript.qmd
Updated manuscript

@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see @tbl-table1).
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see @tbl-table1).**Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.
 
@@
 
-The experiment's setup follows the born-open principle [@deleeuw2023].
-The online experiment, implemented entirely using *JsPsych* [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
-Participants' raw data files are then automatically stored on a European server on OSF; and the deanonymized and ready-to-use data (downloaded from OSF and formatted using a reproducible script) will be made available in the same GitHub repository, alongside the scripts used for the analysis.
+The experiment’s setup follows **a modified version** of the born-open principle (De Leeuw, 2023), **with adjustments made to ensure data privacy, confidentiality and transparency**.
+The online experiment, implemented entirely using JsPsych (De Leeuw, 2015), has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
+**Participant’s raw data files will be sent directly to a private OSF repository.
+These files will then undergo pre-processing to anonymize the data by removing any sensitive information, such as Prolific and SONA IDs.
+The anonymized data will then be saved on GitHub, where it will be accessible for analysis, along with the pre-processing script**.
 
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity).
+**Although these variables are not analyzed in the current study, they will be made available to provide a detailed and thorough description of the sample and to ensure consistency with previous studies using the paradigm (Makowski et al., 2023), as well as maximizing reusability**.
 Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
@@
 After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
+**To mitigate the speed/accuracy trade-off, inherent to most decision-making tasks, we will maintain a double-constrained instruction by emphasizing with equal weight to “be as fast as possible without making errors”**.
 
@@
 
-The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.
+The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.**These questionnaires are strategically included to provide a break between the two cognitively taxing blocks of comparable activity and duration and to maximize the reusability of the data by maintaining consistency with paradigms used previously (Makowski et al., 2023)**.
 

### 2024-06-07 paper/R0/manuscript.qmd
edits
@AnafNeves made some edits in the revision, can you transfer to ms and knit

--- 2024-06-04 paper/manuscript.qmd
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see @tbl-table1).**Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1).
+**Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.
 
@@
 This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.
-| Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+| Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the `BayesFactor::BFCorrelation()` function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
 BF10 > 3 will be interpreted as evidence for a relationship between these two measures.

### 2024-06-07 paper/R0/manuscript.qmd
Updated manuscripts

@@
 The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1).
-**Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.
+Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout.
 
@@
 
-The experiment’s setup follows **a modified version** of the born-open principle (De Leeuw, 2023), **with adjustments made to ensure data privacy, confidentiality and transparency**.
+The experiment’s setup follows a modified version of the born-open principle (De Leeuw, 2023), with adjustments made to ensure data privacy, confidentiality and transparency.
 The online experiment, implemented entirely using JsPsych (De Leeuw, 2015), has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
-**Participant’s raw data files will be sent directly to a private OSF repository.
+Participant’s raw data files will be sent directly to a private OSF repository.
 These files will then undergo pre-processing to anonymize the data by removing any sensitive information, such as Prolific and SONA IDs.
-The anonymized data will then be saved on GitHub, where it will be accessible for analysis, along with the pre-processing script**.
+The anonymized data will then be saved on GitHub, where it will be accessible for analysis, along with the pre-processing script.
 
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity).
-**Although these variables are not analyzed in the current study, they will be made available to provide a detailed and thorough description of the sample and to ensure consistency with previous studies using the paradigm (Makowski et al., 2023), as well as maximizing reusability**.
+Although these variables are not analyzed in the current study, they will be made available to provide a detailed and thorough description of the sample and to ensure consistency with previous studies using the paradigm (Makowski et al., 2023), as well as maximizing reusability.
 Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
@@
 After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
-**To mitigate the speed/accuracy trade-off, inherent to most decision-making tasks, we will maintain a double-constrained instruction by emphasizing with equal weight to “be as fast as possible without making errors”**.
+To mitigate the speed/accuracy trade-off, inherent to most decision-making tasks, we will maintain a double-constrained instruction by emphasizing with equal weight to “be as fast as possible without making errors”.
 
@@
 
-The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.**These questionnaires are strategically included to provide a break between the two cognitively taxing blocks of comparable activity and duration and to maximize the reusability of the data by maintaining consistency with paradigms used previously (Makowski et al., 2023)**.
+The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.
+These questionnaires are strategically included to provide a break between the two cognitively taxing blocks of comparable activity and duration and to maximize the reusability of the data by maintaining consistency with paradigms used previously (Makowski et al., 2023).
 

### 2024-06-07 paper/R0/manuscript.qmd
edits

@@
 The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1).
-Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout.
 
@@
 
-The experiment’s setup follows a modified version of the born-open principle (De Leeuw, 2023), with adjustments made to ensure data privacy, confidentiality and transparency.
-The online experiment, implemented entirely using JsPsych (De Leeuw, 2015), has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
-Participant’s raw data files will be sent directly to a private OSF repository.
-These files will then undergo pre-processing to anonymize the data by removing any sensitive information, such as Prolific and SONA IDs.
-The anonymized data will then be saved on GitHub, where it will be accessible for analysis, along with the pre-processing script.
+The experiment’s setup follows of the born-open principle [@deleeuw2023].
+The online experiment, implemented entirely using JsPsych [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
+**Participant’s raw data files (containing identifiers) is automatically stored in a private OSF repository.
+The preprocessing and analysis scripts, as well as the anonymized data, will be available directly on GitHub, ensuring the transparency and reproducibility of all the analysis steps.**
 
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity).
-Although these variables are not analyzed in the current study, they will be made available to provide a detailed and thorough description of the sample and to ensure consistency with previous studies using the paradigm (Makowski et al., 2023), as well as maximizing reusability.
+**Although these variables are directly not analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.**.
 Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
@@
 After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
-To mitigate the speed/accuracy trade-off, inherent to most decision-making tasks, we will maintain a double-constrained instruction by emphasizing with equal weight to “be as fast as possible without making errors”.
+**To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors**.
 
@@
 The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.
-These questionnaires are strategically included to provide a break between the two cognitively taxing blocks of comparable activity and duration and to maximize the reusability of the data by maintaining consistency with paradigms used previously (Makowski et al., 2023).
+**These questionnaires are included as a way of providing a break between the two cognitively taxing blocks and maintain paradigmatic consistency with previous studies [@makowski2023]**.
 

### 2024-06-09 paper/R0/manuscript.qmd
submitted

@@
 To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
-This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has been demonstrated to be compatible with online experiments [@lush2022].
+This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has demonstrated validity in online experiments [@lush2022].
 
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1).
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion paradigm that encompasses other illusion types (see @tbl-table1).
 

### 2025-02-28 paper/R1/manuscript.qmd
Edits to manuscript
Un-bolded old comments, checked everything that said "addressed" and moved to new R1 folder.

--- 2024-06-09 paper/R0/manuscript.qmd
@@
 Their robust and reliable effect makes them useful stimuli to explore how perception is constructed and shaped, and several theoretical models have been put forth to explain how they work.
-In particular, illusions have been recently reframed using a predictive coding account of perception [@notredame2014] in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].
+**In particular, illusions have been reframed using a predictive coding account of perception [@notredame2014] in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].**
 
@@
 Indeed, traditional methods frequently focus on participant's experience by prompting them to assess the difference between two identical targets, estimate the target's physical properties, or adjust the targets to match a reference stimulus [@todorovic2020].
-Relying on metacognitive judgments about one's subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion sensibility.
+Relying on metacognitive judgments about one's subjective experiences adds an additional layer to the measure that might not be desired when attempting to measure illusion **sensitivity**.
 Moreover, paradigms often face challenges in diversifying the illusory effects (i.e., using multiple stimuli to experimentally manipulate the strength of the illusion) and the illusion types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus, Delboeuf which might rely on a different admixture of mechanisms), hindering the potential of obtaining a comprehensive, valid, and reliable measure of illusion sensitivity.
@@
 To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
-This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale and has demonstrated validity in online experiments [@lush2022].
+**This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale (from 0-5) and has demonstrated validity in online experiments [@lush2022].**
 
@@
 The online experiment, implemented entirely using JsPsych [@de2015jspsych], has its code stored on GitHub and will leverage the power of the platform to host the experiment for free.
-**Participant’s raw data files (containing identifiers) is automatically stored in a private OSF repository.
-The preprocessing and analysis scripts, as well as the anonymized data, will be available directly on GitHub, ensuring the transparency and reproducibility of all the analysis steps.**
+Participant’s raw data files (containing identifiers) **are** automatically stored in a private OSF repository.
+The preprocessing and analysis scripts, as well as the anonymized data, will be available directly on GitHub, ensuring the transparency and reproducibility of all the analysis steps.
 
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity).
-**Although these variables are directly not analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.**.
-Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
+**Although these variables are not directly analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.** Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
 
@@
 This will be followed by 10 suggestions in a fixed order [see @lush2021], such as "now extend your arms ahead of you, with palms facing each other, hands about a foot apart" and "as you sit comfortably in your chair with your eyes closed, a picture of two balls will be displayed on the computer screen".
-Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale.
-Phenomenological control will be indexed by averaging the scores from the 10 scales.
+**Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale (from 0-5).** Phenomenological control will be indexed by averaging the scores from the 10 scales.
 
@@
 After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
-**To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors**.
+To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors.
 
@@
 Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, separately for the 3 illusion types.
-Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy trade off), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current preregistration.
+Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy trade off), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current **registered report**.
 
 The two sets of 3 illusion blocks will be separated by 2 short questionnaires acting as a break, namely the IPIP-6 [@sibley2011mini], measuring 6 personality traits with 24 analogue scales items, and the PID-5 [@krueger2011], measuring 5 maladaptive personality traits with 25 Likert scales items.
-**These questionnaires are included as a way of providing a break between the two cognitively taxing blocks and maintain paradigmatic consistency with previous studies [@makowski2023]**.
+These questionnaires are included as a way of providing a break between the two cognitively taxing blocks and maintain paradigmatic consistency with previous studies [@makowski2023].
 
@@
 The PCS will contain several manipulation check indices to identify problematic participants.
-Participants should not answer "no balls were presented" when queried about the colour they observed on the screen following the negative visual hallucination suggestion, and should execute at least 5 space presses upon instructions to do so.
+**The phenomenological control task consists of various auditory and visual exercises.
+In one such exercise, participants receive the following instruction: “Open your eyes.
+You will see only two balls on the screen…just two balls”.
+However, three differently coloured balls are actually displayed.
+If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.
+In another exercise, participants are asked to press the spacebar six times.
+If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.
+Participants will be excluded if they fail at least one of these checks.
+Participants will be excluded if they fail at least one of the attention checks.**
 
 Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023].
-If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first block, the entire participant will be discarded (suggesting that they did not properly do the text), but if only the second block is bad, then only the second block will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
-In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
+**If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).** In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
 

### 2025-02-28 experiment/demographics.js
changes to scripts
questionnaires, demographics and consent now use jsPsychSurvey

@@ -23,6 +23,7 @@ var demographics_browser_info = {
     },
 }
 
+
 var restrict_mobile = {
     timeline: [
         {
@@ -42,36 +43,60 @@ var restrict_mobile = {
 }
 
 // Consent form ========================================================================
-var demographics_consent = {
-    type: jsPsychHtmlButtonResponse,
-    css_classes: ["narrow-text"],
-    stimulus:
+const ConsentForm = {
+    type: jsPsychSurvey,
+    survey_json: function () {
+        // Get URL variables
+        let urlvars = jsPsych.data.urlVariables()
         // Logo and title
-        "<img src='https://blogs.brighton.ac.uk/sussexwrites/files/2019/06/University-of-Sussex-logo-transparent.png' width='150px' align='right'/><br><br><br><br><br>" +
-        "<h1>Informed Consent</h1>" +
-        // Overview
-        "<p align='left'><b>Invitation to Take Part</b><br>" +
-        "You are being invited to take part in a research study to further our understanding of Human psychology. Thank you for carefully reading this information sheet. This study is being conducted by Dr Dominique Makowski from the School of Psychology, University of Sussex, who is happy to be contacted (D.Makowski@sussex.ac.uk) if you have any questions.</p>" +
-        // Description
-        "<p align='left'><b>Why have I been invited and what will I do?</b><br>" +
-        "The goal is to study [TODO] The whole experiment will take you <b>about 35 min</b> to complete. Please make you sure that you are in a quiet environment, <b>on a computer</b> (the experiment is not mobile-friendly), and that you have time to complete it in one go.</p>" +
-        // Results and personal information
-        "<p align='left'><b>What will happen to the results and my personal information?</b><br>" +
-        "The results of this research may be written into a scientific publication. Your anonymity will be ensured in the way described in the consent information below. Please read this information carefully and then, if you wish to take part, please acknowledge that you have fully understood this sheet, and that you consent to take part in the study as it is described here.</p>" +
-        "<p align='left'><b>Consent</b><br></p>" +
-        // Bullet points
-        "<li align='left'>I understand that by signing below I am agreeing to take part in the University of Sussex research described here, and that I have read and understood this information sheet</li>" +
[…]

### 2025-02-28 experiment/questionnaires.js

@@ -1,176 +1,175 @@
+const questionnaires_instructions = {
+    type: jsPsychHtmlButtonResponse,
+    stimulus:
+        "<h2>Questionnaires</h2>" +
+        "<p>We will continue with a series of questionnaires.<br>It is important that you answer truthfully. Please read the statements carefully and answer according to what describe you the best.</p>",
+    choices: ["Continue"],
+    data: { screen: "questionnaire_instructions" },
+}
+
+// Convernience function to shuffle an object (used internally)
+function shuffleObject(obj) {
+    const entries = Object.entries(obj)
+    for (let i = entries.length - 1; i > 0; i--) {
+        const j = Math.floor(Math.random() * (i + 1))
+            ;[entries[i], entries[j]] = [entries[j], entries[i]]
+    }
+    return Object.fromEntries(entries)
+}
+
 // Items =================================================
 // Mini-IPIP6 questionnaire (Sibley, 2011)
-var ipip6_items = [
-    "I am the life of the party",
-    "I sympathise with others' feelings",
-    "I get chores done right away",
-    "I have frequent mood swings",
-    "I have a vivid imagination",
-    "I feel entitled to more of everything",
-    "I don't talk a lot",
-    "I am not interested in other people's problems",
-    "I have difficulty understanding abstract ideas",
-    "I like order",
-    "I make a mess of things",
-    "I deserve more things in life",
-    "I do not have a good imagination",
-    "I feel others' emotions",
-    "I am relaxed most of the time",
-    "I get upset easily",
-    "I seldom feel blue",
-    "I would like to be seen driving around in a really expensive car",
-    "I keep in the background",
-    "I am not really interested in others",
-    "I am not interested in abstract ideas",
-    "I often forget to put things back in their proper place",
-    "I talk to a lot of different people at parties",
-    "I would get a lot of pleasure from owning expensive luxury goods",
-]
-var ipip6_dimensions = [
-    "Extraversion_1",
-    "Agreeableness_2",
-    "Conscientiousness_3",
-    "Neuroticism_4",
-    "Openness_5",
-    "HonestyHumility_6_R",
-    "Extraversion_7_R",
-    "Agreeableness_8_R",
-    "Openness_9_R",
-    "Conscientiousness_10",
-    "Conscientiousness_11_R",
-    "HonestyHumility_12_R",
-    "Openness_13_R",
-    "Agreeableness_14",
[…]

### 2025-02-28 analysis/1_cleaning.qmd
started analyses files

---
title: "Illusion Game & PCS"
subtitle: "Preprocessing"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
```

### 2025-03-04 paper/R1/manuscript.qmd
Adding figures

@@
 ::: {#tbl-table1}
+```{r}
+question <- c("Question","Hypothesis","Sampling Plan","Analysis Plan","Rationale for Deciding the Sensitivity of the Test","Interpretation Given Different Outcomes","Theory That Could Be Shown Wrong by the Outcomes")
+answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the *BayesFactor::BFCorrelation()* function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+BF10 > 3 will be interpreted as evidence for a relationship between these two measures.", "The hypothesis that VI sensitivity is independent from PC.", "The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.")
 
-| **Question**                                                                                                                                 | **Hypothesis**                                                                                                                            | **Sampling Plan**                                                                                                               | **Analysis Plan**                                                                                                                                                                             | **Rationale for Deciding the Sensitivity of the Test**                                                                                           | **Interpretation Given Different Outcomes**                         | **Theory That Could Be Shown Wrong by the Outcomes**                                                                                                        |
-|----------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------|
-| Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?
-| Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS         | The goal is to recruit around 500 adult English speakers using Prolific.
-This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.
-| Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the `BayesFactor::BFCorrelation()` function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
-BF10 > 3 will be interpreted as evidence for a relationship between these two measures.
-| The hypothesis that VI sensitivity is independent from PC.
-| The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-|
+# join variables to create dataframe
+study_design_df <- data.frame(question, answer)
+
+## study design table
+study_design_table <- flextable::flextable(study_design_df)
+study_design_table <- set_table_properties(study_design_table, width = 1, layout = "autofit")
+study_design_table <- set_caption(study_design_table, caption = "Study Design Table")
+study_design_table <- delete_part(x = study_design_table, part = "header")
+study_design_table <- bold(study_design_table, j = "question", bold = TRUE, part = "body")
+study_design_table <- theme_apa(study_design_table)
+
+study_design_table
+```
 
@@
 
+::: {#tbl-figure1}
+
+![The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.webp)
+
+:::
+
+::: {#tbl-table3}
+
+```{r}
+illusion <- c("Ebbinghaus", "Müller-Lyer", "Vertical-Horizontal")
+example <- c("img/Ebbinghaus_Demo.png", "img/MullerLyer_Demo.png", "img/VerticalHorizontal_Demo.png")
+instructions <- c("Which red circle is bigger?", "Which red line is longer?", "Which red line is longer?")
+arrows <- c("img/answer_leftright_keyboard.png", "img/answer_updown_keyboard.png", "img/answer_leftright_keyboard.png")
+description <- c("Two circles surrounded by other circles.
+The circle surrounded by smaller circles appears larger than the one with the larger surrounding circles.", "Two parallel segments that end with inwards/outwards pointing arrow-like fins.
+The segment with inward-pointing fins is typically perceived to be longer", "Two lines segments, one horizontal and one angled.
+The angled line is usually perceived as longer.")
+
+illusion_example_df <- data.frame(illusion, example, arrows, instructions, description)
+
+illusion_example_table <- flextable::flextable(illusion_example_df, col_keys = c("illusion", "example", "instructions", "description")) |>
+    mk_par(
+    j = "instructions",
+    value = as_paragraph(
+      as_chunk(x = instructions),
+      as_image(src = arrows, width = .5, height = 0.5),
+      " "
+    )
+  )
+
+illusion_example_table <- illusion_example_table |>
+  colformat_image(j = "example", width = 1, height = 0.75)
+
+illusion_example_table <- set_table_properties(illusion_example_table, width = 1, layout = "autofit")
+illusion_example_table <- flextable::set_header_labels(illusion_example_table, illusion = "Illusion", example = "Example", instructions = "Task", description = "Description")
+illusion_example_table <- theme_apa(illusion_example_table)
+
+illusion_example_table
+```
+
+Illusion Task Example Stimuli
+:::
+
 ## Data Analysis

### 2025-03-04 paper/R1/manuscript.qmd
Minor changes
illsuion gme ppts/jpg
@EmmaLBenn let me know if this works

@@
     documentmode: man
+output: officedown::rdocx_document
 ---
@@
 ```{r}
+#| tbl-cap: "Study Design Table"
+
 question <- c("Question","Hypothesis","Sampling Plan","Analysis Plan","Rationale for Deciding the Sensitivity of the Test","Interpretation Given Different Outcomes","Theory That Could Be Shown Wrong by the Outcomes")
 answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
-This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the *BayesFactor::BFCorrelation()* function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
 BF10 > 3 will be interpreted as evidence for a relationship between these two measures.", "The hypothesis that VI sensitivity is independent from PC.", "The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.")
@@
 study_design_table <- set_table_properties(study_design_table, width = 1, layout = "autofit")
-study_design_table <- set_caption(study_design_table, caption = "Study Design Table")
+#study_design_table <- set_caption(study_design_table, caption = "Study Design Table")
 study_design_table <- delete_part(x = study_design_table, part = "header")
@@
 ![The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.webp)
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.png)
 
@@
 ```{r}
-illusion <- c("Ebbinghaus", "Müller-Lyer", "Vertical-Horizontal")
-example <- c("img/Ebbinghaus_Demo.png", "img/MullerLyer_Demo.png", "img/VerticalHorizontal_Demo.png")
-instructions <- c("Which red circle is bigger?", "Which red line is longer?", "Which red line is longer?")
-arrows <- c("img/answer_leftright_keyboard.png", "img/answer_updown_keyboard.png", "img/answer_leftright_keyboard.png")
-description <- c("Two circles surrounded by other circles.
+#| tbl-cap: "Illusion bla bla (look at the illusion game validation paper for some inspiration on the figure caption) "
+# illusion <- c("Ebbinghaus", "Müller-Lyer", "Vertical-Horizontal")
+# example <- c("img/Ebbinghaus_Demo.png", "img/MullerLyer_Demo.png", "img/VerticalHorizontal_Demo.png")
+# instructions <- c("Which red circle is bigger?", "Which red line is longer?", "Which red line is longer?")
+# arrows <- c("img/answer_leftright_keyboard.png", "img/answer_updown_keyboard.png", "img/answer_leftright_keyboard.png")
+# description <- c("Two circles surrounded by other circles.
 The circle surrounded by smaller circles appears larger than the one with the larger surrounding circles.", "Two parallel segments that end with inwards/outwards pointing arrow-like fins.
@@
 The angled line is usually perceived as longer.")
+# 
+# illusion_example_df <- data.frame(illusion, example, arrows, instructions, description)
+# 
+# illusion_example_table <- flextable::flextable(illusion_example_df, col_keys = c("illusion", "example", "instructions", "description")) |>
+#     mk_par(
+#     j = "instructions",
+#     value = as_paragraph(
+#       as_chunk(x = instructions),
+#       as_image(src = arrows, width = .5, height = 0.5),
+#       " "
+#     )
+#   )
+# 
+# illusion_example_table <- illusion_example_table |>
+#   colformat_image(j = "example", width = 1, height = 0.75)
+# 
+# illusion_example_table <- flextable::set_table_properties(illusion_example_table, width = 1, layout = "autofit")
+# illusion_example_table <- flextable::set_header_labels(illusion_example_table, illusion = "Illusion", example = "Example", instructions = "Task", description = "Description")
+# illusion_example_table <- flextable::theme_apa(illusion_example_table)
+# 
+# # illusion_example_table
+# 
+# save_as_pptx(illusion_example_table, path = "IllusionTable.pptx")
 
-illusion_example_df <- data.frame(illusion, example, arrows, instructions, description)
+knitr::include_graphics("img/IllusionTable.jpg")
 
-illusion_example_table <- flextable::flextable(illusion_example_df, col_keys = c("illusion", "example", "instructions", "description")) |>
-    mk_par(
-    j = "instructions",
-    value = as_paragraph(
-      as_chunk(x = instructions),
-      as_image(src = arrows, width = .5, height = 0.5),
-      " "
-    )
-  )
-
-illusion_example_table <- illusion_example_table |>
-  colformat_image(j = "example", width = 1, height = 0.75)
-
-illusion_example_table <- set_table_properties(illusion_example_table, width = 1, layout = "autofit")
-illusion_example_table <- flextable::set_header_labels(illusion_example_table, illusion = "Illusion", example = "Example", instructions = "Task", description = "Description")
-illusion_example_table <- theme_apa(illusion_example_table)
-
-illusion_example_table
 ```

### 2025-03-05 experiment/demographics.js
updated consent
- not finsihed

@@ -61,13 +61,14 @@ const ConsentForm = {
         text +=
             // Overview
             "<p align='left'><b>Invitation to Take Part</b><br>" +
-            "You are being invited to take part in a research study to further our understanding of Human psychology. Thank you for carefully reading this information sheet. This study is being conducted by Dr Dominique Makowski from the School of Psychology, University of Sussex, who is happy to be contacted (D.Makowski@sussex.ac.uk) if you have any questions.</p>" +
+            "You are being invited to take part in a research study to further our understanding of Human psychology." +
+            "Thank you for carefully reading this information sheet. This study is being conducted by Dr Dominique Makowski from the University of Sussex and his team (see contact information below)." +
             // Description
             "<p align='left'><b>Why have I been invited and what will I do?</b><br>" +
-            "The goal is to study [TODO] The whole experiment will take you <b>about 35 min</b> to complete. Please make you sure that you are in a quiet environment, <b>on a computer</b> (the experiment is not mobile-friendly), and that you have time to complete it in one go.</p>" +
+            "The goal is to study [TODO] The whole experiment will take you <b style='color:#FF5722;'>~30 min</b> to complete. Please make you sure that you are in a quiet environment, <b>on a computer</b> (the experiment is not mobile-friendly), and that you have time to complete it in one go.</p>" +
             // Results and personal information
             "<p align='left'><b>What will happen to the results and my personal information?</b><br>" +
-            "The results of this research may be written into a scientific publication. Your anonymity will be ensured in the way described in the consent information below. Please read this information carefully and then, if you wish to take part, please acknowledge that you have fully understood this sheet, and that you consent to take part in the study as it is described here.</p>" +
+            "The results of this research may be written into a scientific publication. Your anonymity will be ensured in the way described in the consent information below. <b>Please read this information carefully</b> and then, if you wish to take part, please acknowledge that you have fully understood this sheet, and that you consent to take part in the study as it is described here.</p>" +
             "<p align='left'><b>Consent</b><br></p>" +
             // Bullet points
             "<li align='left'>I understand that by signing below I am agreeing to take part in the University of Sussex research described here, and that I have read and understood this information sheet</li>" +
@@ -79,7 +80,7 @@ const ConsentForm = {
[…]

### 2025-03-10 experiment/IllusionGame.js
added illusion images

@@ -1,0 +1,449 @@
+// Global variables ===============================================================================
+var block_number = 1 // block indexing variable
+var trial_number = 1 // trial indexing variable
+
+// Instructions ==================================================================================
+// General instructions
+var IG_instructions = {
+    type: jsPsychHtmlButtonResponse,
+    choices: [ig_text_startpractice],
+    stimulus: ig_instructions,
+    data: { screen: "IG_instructions" },
+}
+
+function add_blocknumber(instructions, block) {
+    return "<p><b>" + ig_text_part + block + "/6" + "</b></p>" + instructions
+}
+
+// Math utilities =================================================================================
+function randomInteger(min = 1, max = 10) {
+    return Math.round(Math.random() * (max - min) + min)
+}
+
+function cumulative_probability(x, mean, sd) {
+    var z = (x - mean) / Math.sqrt(2 * sd * sd)
+    var t = 1 / (1 + 0.3275911 * Math.abs(z))
+    var a1 = 0.254829592
+    var a2 = -0.284496736
+    var a3 = 1.421413741
+    var a4 = -1.453152027
+    var a5 = 1.061405429
+    var erf = 1 - ((((a5 * t + a4) * t + a3) * t + a2) * t + a1) * t * Math.exp(-z * z)
+    var sign = 1
+    if (z < 0) {
+        sign = -1
+    }
+    return (1 / 2) * (1 + sign * erf)
+}
+
+function round_digits(x, digits = 2) {
+    return Number(Math.round(parseFloat(x + "e" + digits)) + "e-" + digits).toFixed(digits)
+}
+
+// EEG / Physio trigger
+// [x, y, width, height] in pixels. Set to [0, 0, 0, 0] to disable.
+const marker_position = [0, 0, 100, 100]
+
+function ig_create_marker(marker_position, color = "black") {
+    const html = `<div id="marker" style="position: absolute; background-color: ${color};\
+left:${marker_position[0]}px; top:${marker_position[1]}px; \
+width:${marker_position[2]}px; height:${marker_position[3]}px";></div>`
+    document.querySelector("body").insertAdjacentHTML("beforeend", html)
+}
+// Feedback and Debriefing ========================================================================
+function get_results(illusion_mean, illusion_sd, illusion_type) {
+    if (typeof illusion_type != "undefined") {
+        var trials = jsPsych.data.get().filter({ screen: "IG_Trial", type: illusion_type }) // results by block
+    } else {
[…]

### 2025-03-10 experiment/instructions.js

@@ -1,0 +1,61 @@
+const path_ig = "https://realitybending.github.io/IllusionGame/v3/"
+
+// Instructions ==================================================================================
+// General instructions
+var ig_instructions =
+    "<h1>Illusion Game</h1>" +
+    "<p>In this game of speed and reflex, you will need to make <b>visual judgments</b> (for instance, which circle is the biggest, or which line is the longest), as fast and as correctly as possible, while <b>resisting different visual illusions</b>.</p>" +
+    "<p>We are going to start with some examples of all the illusions you will encounter in the game.</p>"
+
+// Instructions for Illusion Trials
+var mullerlyer_instructions =
+    "<p>In this part, two horizontal red lines will appear one above the other.</p>" +
+    "<p>Your task is to select which <b>line is longer</b> in length as fast as you can, without making errors.</p>" +
+    "<p>Don't get distracted by the surrounding black arrows at the end of the red lines!</p>" +
+    "<p>Press <b>the UP or the DOWN arrow</b> to indicate where is the longer <b>red line</b>.</p>" +
+    `<div style='float: center'><img src='${path_ig}/images/MullerLyer_Demo.png' height='200'></img>` +
+    `<p><img src='${path_ig}/images/answer_updown_keyboard.PNG' height='100'></img></p>` +
+    "<p class='small'>In this example, the correct answer is the <b>UP arrow</b>.</p></div>" +
+    "<p>Are you ready? <b>Press ENTER to start</b></p>"
+
+var ebbinghaus_instructions =
+    "<p>In this part, two red circles will appear side by side on the screen.</p>" +
+    "<p>Your task is to select which <b>red circle is bigger</b> in size as fast as you can, without making errors.</p>" +
+    "<p>Don't get distracted by the surrounding black circles around the red circles!</p>" +
+    "<p>Press <b>the LEFT or the RIGHT arrow</b> to indicate which is the bigger <b>red circle</b>.</p>" +
+    `<div style='float: center'><img src='${path_ig}/images/Ebbinghaus_Demo.png' height='200'></img>` +
+    `<p><img src='${path_ig}/images/answer_leftright_keyboard.PNG' height='100'></img></p>` +
+    "<p class='small'>In this example, the correct answer is the <b>LEFT arrow</b>.</p></div>" +
+    "<p>Are you ready? <b>Press ENTER to start</b></p>"
+
+var verticalhorizontal_instructions =
[…]

### 2025-03-12 paper/R1/manuscript.qmd
Minor changes

--- 2025-03-04 paper/R1/manuscript.qmd
@@
 
-::: {#tbl-figure1}
+::: {#fig-figure1}
 
 ![The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.webp)
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.png)
 
+The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.
 :::
 
-::: {#tbl-table3}
+::: {#tbl-table2}
 
-```{r}
-illusion <- c("Ebbinghaus", "Müller-Lyer", "Vertical-Horizontal")
-example <- c("img/Ebbinghaus_Demo.png", "img/MullerLyer_Demo.png", "img/VerticalHorizontal_Demo.png")
-instructions <- c("Which red circle is bigger?", "Which red line is longer?", "Which red line is longer?")
-arrows <- c("img/answer_leftright_keyboard.png", "img/answer_updown_keyboard.png", "img/answer_leftright_keyboard.png")
-description <- c("Two circles surrounded by other circles.
-The circle surrounded by smaller circles appears larger than the one with the larger surrounding circles.", "Two parallel segments that end with inwards/outwards pointing arrow-like fins.
-The segment with inward-pointing fins is typically perceived to be longer", "Two lines segments, one horizontal and one angled.
-The angled line is usually perceived as longer.")
+![Illusions Task Example Stimuli.](img/illusion_table.png)
 
-illusion_example_df <- data.frame(illusion, example, arrows, instructions, description)
-
-illusion_example_table <- flextable::flextable(illusion_example_df, col_keys = c("illusion", "example", "instructions", "description")) |>
-    mk_par(
-    j = "instructions",
-    value = as_paragraph(
-      as_chunk(x = instructions),
-      as_image(src = arrows, width = .5, height = 0.5),
-      " "
-    )
-  )
-
-illusion_example_table <- illusion_example_table |>
-  colformat_image(j = "example", width = 1, height = 0.75)
-
-illusion_example_table <- set_table_properties(illusion_example_table, width = 1, layout = "autofit")
-illusion_example_table <- flextable::set_header_labels(illusion_example_table, illusion = "Illusion", example = "Example", instructions = "Task", description = "Description")
-illusion_example_table <- theme_apa(illusion_example_table)
-
-illusion_example_table
-```
-
-Illusion Task Example Stimuli
+Illusions Task Example Stimuli.
 :::
@@
 If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.
-Participants will be excluded if they fail at least one of these checks.
 Participants will be excluded if they fail at least one of the attention checks.**

### 2025-03-14 experiment/IllusionGame.js
fixed issues with images not showing up

@@ -165,7 +165,7 @@ function IG_create_trial(illusion_name = "Ponzo", type = "updown", marker = true
     var trial = {
         type: jsPsychImageKeyboardResponse,
         stimulus: function () {
-            return path_ig + jsPsych.timelineVariable("stimulus")
+            return jsPsych.timelineVariable("stimulus")
         },
         data: function () {
             return jsPsych.timelineVariable("data")
@@ -249,7 +249,7 @@ function IG_make_trials(stimuli, instructions, illusion_name, type, marker = tru
     // Preload images
     timeline.push({
         type: jsPsychPreload,
-        images: stim_list.map((a) => path_ig + a.stimulus),
+        images: stim_list.map((a) => a.stimulus),
         data: { screen: "IG_Preload" },
     })
 
@@ -275,7 +275,7 @@ function IG_make_trials(stimuli, instructions, illusion_name, type, marker = tru
         randomize_order: true,
         repetitions: 1,
         on_load: function () {
-            let path = stim_list.map(a => path_ig + a.stimulus)
+            let path = stim_list.map(a => a.stimulus)
             console.log(path) // Debugging to check if it works
         },
     })

### 2025-03-14 experiment/instructions.js

@@ -1,4 +1,3 @@
-const path_ig = "https://realitybending.github.io/IllusionGame/v3/"
 
 // Instructions ==================================================================================
 // General instructions
@@ -13,8 +12,8 @@ var mullerlyer_instructions =
     "<p>Your task is to select which <b>line is longer</b> in length as fast as you can, without making errors.</p>" +
     "<p>Don't get distracted by the surrounding black arrows at the end of the red lines!</p>" +
     "<p>Press <b>the UP or the DOWN arrow</b> to indicate where is the longer <b>red line</b>.</p>" +
-    `<div style='float: center'><img src='${path_ig}/images/MullerLyer_Demo.png' height='200'></img>` +
-    `<p><img src='${path_ig}/images/answer_updown_keyboard.PNG' height='100'></img></p>` +
+    `<div style='float: center'><img src='images/MullerLyer_Demo.png' height='200'></img>` +
+    `<p><img src='images/answer_updown_keyboard.PNG' height='100'></img></p>` +
     "<p class='small'>In this example, the correct answer is the <b>UP arrow</b>.</p></div>" +
     "<p>Are you ready? <b>Press ENTER to start</b></p>"
 
@@ -23,8 +22,8 @@ var ebbinghaus_instructions =
     "<p>Your task is to select which <b>red circle is bigger</b> in size as fast as you can, without making errors.</p>" +
     "<p>Don't get distracted by the surrounding black circles around the red circles!</p>" +
     "<p>Press <b>the LEFT or the RIGHT arrow</b> to indicate which is the bigger <b>red circle</b>.</p>" +
-    `<div style='float: center'><img src='${path_ig}/images/Ebbinghaus_Demo.png' height='200'></img>` +
-    `<p><img src='${path_ig}/images/answer_leftright_keyboard.PNG' height='100'></img></p>` +
+    `<div style='float: center'><img src='images/Ebbinghaus_Demo.png' height='200'></img>` +
+    `<p><img src='images/answer_leftright_keyboard.PNG' height='100'></img></p>` +
     "<p class='small'>In this example, the correct answer is the <b>LEFT arrow</b>.</p></div>" +
     "<p>Are you ready? <b>Press ENTER to start</b></p>"
 
@@ -33,8 +32,8 @@ var verticalhorizontal_instructions =
[…]

### 2025-03-18 paper/R1/manuscript.qmd
minor

@@
 ```{r}
-question <- c("Question","Hypothesis","Sampling Plan","Analysis Plan","Rationale for Deciding the Sensitivity of the Test","Interpretation Given Different Outcomes","Theory That Could Be Shown Wrong by the Outcomes")
+#| tbl-cap: "Study Design Table"
+
+question <- c("Question", "Hypothesis", "Sampling Plan", "Analysis Plan", "Rationale for Deciding the Sensitivity of the Test", "Interpretation Given Different Outcomes", "Theory That Could Be Shown Wrong by the Outcomes")
 answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
-This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the *BayesFactor::BFCorrelation()* function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
 BF10 > 3 will be interpreted as evidence for a relationship between these two measures.", "The hypothesis that VI sensitivity is independent from PC.", "The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.")
@@
 study_design_table <- set_table_properties(study_design_table, width = 1, layout = "autofit")
-study_design_table <- set_caption(study_design_table, caption = "Study Design Table")
+# study_design_table <- set_caption(study_design_table, caption = "Study Design Table")
 study_design_table <- delete_part(x = study_design_table, part = "header")
@@
 
+```{r}
+#| warning: false
+#| label: "fig-illusion_framework"
+#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
+#| out-width: "100%"
+#| fig-cap: "The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."
+
+knitr::include_graphics("img/example_stimuli.png")
+```
+
 For each illusion type, two continuous dimensions are orthogonally manipulated [see @makowski2021 for details on the rationale and execution], namely task difficulty and illusion strength, so that each trial corresponds to a unique combination.
@@
 
-::: {#fig-figure1}
+```{r}
+#| warning: false
+#| label: "fig-illusionexample"
+#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
+#| out-width: "100%"
+#| fig-cap: "The study involved three visual illusions, in which participants were instructed to respond as quickly as possible without making errors.
+Each illusion included two manipulated parameters: strength (e.g., the angle of the outward- or inward-pointing arrow-like fins in the Müller-Lyer illusion) and difficulty (e.g., the difference in line lengths in the Müller-Lyer illusion."
 
-![The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.](img/example_stimuli.png)
-
-The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength.
-:::
-
-::: {#tbl-table2}
-
-![Illusions Task Example Stimuli.](img/illusion_table.png)
-
-Illusions Task Example Stimuli.
-:::
+knitr::include_graphics("img/IllusionTable.jpg")
+```
 
@@
 If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.
+Participants will be excluded if they fail at least one of these checks.
 Participants will be excluded if they fail at least one of the attention checks.**

### 2025-03-18 paper/R1/manuscript.qmd
Added PCS stimuli table

--- 2025-03-12 paper/R1/manuscript.qmd
@@
 
+::: {tbl-table2}
+```{r}
+suggestion <- c("Hand Lowering", "Moving Hands Together", "Mosquito Hallucination", "Taste Hallucination", "Arm Rigidity", "Arm Immobilization", "Music Hallucination", "Negative Visual Illusion", "Amnesia", "Button Press")
+
+example <- c(
+  "Now the hand and arm feel heavy, as if the weight were pressing down...and as it feels heavier and heavier the hand and arm begin to move down...",
+  "You are thinking of your hands being pulled together, and they begin to move together...",
+"You may not have noticed a mosquito that has been buzzing, singing as mosquitoes do...",
+"Imagine that you have something sweet-tasting in your mouth, like a little sugar... ",
+"Your arm feels stiff as if tightly splinted...  Test how stiff and rigid it is ...",
+"Your hand is getting heavier, very heavy, as though it were being pressed against your lap.",
+"In a few moments, a recording of ‘Happy Birthday to You’ will be played for you.
+When the recording starts the volume will be turned way down and you will probably not be able to hear it, or you will hear it very faintly.",
+"You will see just two balls on the screen...  just two balls...",
+"You will remember nothing of what you did or felt from the time that you closed your eyes until I say to you: 'Now you can remember everything!'",
+"You will press the space bar six times but forget that I told you to do so, just as you will forget the other things, until I tell you, 'Now you can remember everything'."
+             )
+
+question <- c(
+  "How strongly did you feel your hand become heavy, where 0 means you felt your arm was no more heavy than normal and 5 means you felt your arm becoming as heavy as if you had a heavy object in your hand, pulling it down?",
+   "How strongly did you feel a force between your hands,  where 0 means you felt no force at all and 5 means you felt a force so strong it was as if your hands were real magnets?",
+   "How strongly did you feel the sensation of a mosquito being there, in either sound or touch, where 0 means you felt no sensation and 5 means you felt by any means as if there actually was a mosquito there?",
+  "How strongly did you taste a SWEET taste in your mouth, where 0 means you felt no taste at all and 5 means you felt a strong taste? 
+  How strongly did you taste a SOUR taste in your mouth, where 0 means you felt no taste at all and 5 means you felt a strong taste?",
+  "How stiff did your arm feel, where 0 means no more stiffness than normal and 5 means you could feel a stiffness so compelling no amount of effort would overcome it?",
+  "How strongly did you feel a heaviness in your hand, where 0 means you felt no heaviness at all and 5 means your hand felt so heavy it was as if a very heavy object was actually pressing it down?",
+  "Report how clearly you heard the music, where 0 means you did not hear any music at all and 5 means you heard it so clearly it was as though it was coming from the best sound system.",
+  "What color balls did you see in the screen?",
+  "How hard was it to remember events before you were told 'now you can remember everything', where 0 means you could remember events as easily as normal and 5 means you found it so difficult to remember it was as if there was an actual blank in your memory?",
+  "Report how strong an urge you felt to press the space bar, where 0 means you had no urge whatsoever and 5 means you had a clear urge to press the space bar repeatedly. 
+  Report how clearly you remembered being given the instruction to press the space bar six times, where 0 means you were able at that time to remember the instruction normally and 5 means you had no memory of the instruction at that time to press the space bar."
+  )
+
+# join variables to create dataframe
+pcs_df <- data.frame(suggestion, example, question)
+
+## study design table
+pcs_table <- flextable::flextable(pcs_df)
+pcs_table <- set_table_properties(pcs_table, width = 1, layout = "autofit")
+pcs_table <- set_caption(pcs_table, caption = "PCS Stimuli Table")
+# pcs_table <- bold(study_design_table, j = "question", bold = TRUE, part = "body")
+pcs_table <- theme_apa(pcs_table)
+
+pcs_table
+```
+
+:::
+
 ### Illusion Game

### 2025-03-19 paper/R1/manuscript.qmd
Minor changes to PCS table

@@
 
-::: {tbl-table2}
+::: {#tbl-table2}
 ```{r}
@@
 pcs_table <- set_caption(pcs_table, caption = "PCS Stimuli Table")
-# pcs_table <- bold(study_design_table, j = "question", bold = TRUE, part = "body")
+pcs_table <- set_header_labels(pcs_table, suggestion = "Suggestion", example = "Example Stimuli", question = "Task Item(s)")
 pcs_table <- theme_apa(pcs_table)
@@
 
+PCS Stimuli and Task Items
 :::
@@
 
-::: {#tbl-table2}
+::: {#tbl-table3}
 

### 2025-03-24 paper/R1/manuscript.qmd
minor changes to manuscript formatting

--- 2025-03-18 paper/R1/manuscript.qmd
@@
 
-Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by metacognitive abilities difference, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a top-town fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
+Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a **top-down** fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
 If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".
@@
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity). 
-**Although these variables are not directly analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.** Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
+**Although these variables are not directly analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.** 
+Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
 
@@
 This will be followed by 10 suggestions in a fixed order [see @lush2021], such as "now extend your arms ahead of you, with palms facing each other, hands about a foot apart" and "as you sit comfortably in your chair with your eyes closed, a picture of two balls will be displayed on the computer screen".
-**Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale (from 0-5).** Phenomenological control will be indexed by averaging the scores from the 10 scales.
+**Once the 10 suggestions are completed, participants will be asked to rate their subjective experiences and response to each suggestion on a 6-points Likert scale (from 0-5).** 
+Phenomenological control will be indexed by averaging the scores from the 10 scales.
 
@@
 Note that the illusion effect can be "incongruent" (biasing perception in the direction of the incorrect response) or "congruent" (facilitating, i.e., biasing perception in the direction of the correct response).
-Participants respond with a key arrow (left vs.
-right; or up vs.
-down), and their reaction time (RT) and accuracy are recorded.
+Participants respond with a key arrow (left vs. right; or up vs. down), and their reaction time (RT) and accuracy are recorded.
 
@@
 The PCS will contain several manipulation check indices to identify problematic participants. 
-**The phenomenological control task consists of various auditory and visual exercises.
-In one such exercise, participants receive the following instruction: “Open your eyes.
-You will see only two balls on the screen…just two balls”.
-However, three differently coloured balls are actually displayed.
-If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.
-In another exercise, participants are asked to press the spacebar six times.
-If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.
-Participants will be excluded if they fail at least one of these checks.
-Participants will be excluded if they fail at least one of the attention checks.**
+**The phenomenological control task consists of various auditory and visual exercises.**
+**In one such exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
+**However, three differently coloured balls are actually displayed.**
+**If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.**
+**In another exercise, participants are asked to press the spacebar six times.**
+**If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.**
+**Participants will be excluded if they fail at least one of these checks. Participants will be excluded if they fail at least one of the attention checks.**
 
 Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023]. 
-**If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).** In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
+**If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).**
+In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
 

### 2025-03-24 paper/R1/manuscript.qmd
minor

@@
     documentmode: man
-output: officedown::rdocx_document
 ---
@@
 question <- c("Question", "Hypothesis", "Sampling Plan", "Analysis Plan", "Rationale for Deciding the Sensitivity of the Test", "Interpretation Given Different Outcomes", "Theory That Could Be Shown Wrong by the Outcomes")
-answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
-This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings.
+answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence to support the absence of a relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings, for all three illusions.
 BF10 > 3 will be interpreted as evidence for a relationship between these two measures.", "The hypothesis that VI sensitivity is independent from PC.", "The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.")

### 2025-03-28 paper/R1/manuscript.qmd
changes to manuscript

@@
 Their robust and reliable effect makes them useful stimuli to explore how perception is constructed and shaped, and several theoretical models have been put forth to explain how they work. 
-**In particular, illusions have been reframed using a predictive coding account of perception [@notredame2014] in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].**
+**In particular, illusions have been reframed using a predictive coding account of perception [@notredame2014]** in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].
 
@@
 These accounts also provide a compelling framework to explain existing findings reporting interindividual variability in the sensitivity to illusions. 
-Indeed, several studies suggest a potential link with psychopathology, in particular schizophrenia [@costa2023] and autism [@gori2016], in which the reported lower sensitivity to illusions has been attributed to a diminished influence of top-down processes such as prior knowledge [@mitchell2010] and a greater emphasis on (i.e., precision of) sensory information [@palmer2017].
+Indeed, several studies suggest a potential link with psychopathology, in particular schizophrenia [@costa2023] and autism [@gori2016], in which the reported lower sensitivity to illusions has been attributed to a diminished influence of **perceptual** top-down processes such as prior knowledge [@mitchell2010] and a greater emphasis on (i.e., precision of) sensory information [@palmer2017].
 Evidence beyond psychopathology also suggests variability in the general population, potentially correlated with personality traits such as agreeableness and honest-humility [@makowski2023], as well as cognitive abilities [@shoshina2014].
@@
 The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
-This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its interference effect, hopefully bypassing some of the metacognitive processes at stake in other paradigms.
-
-Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a **top-down** fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
+This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its **interference effect —an effect that arises when the presence of an illusion disrupts an individual's ability to accurately discriminate between perceptual stimuli.**
+**This approach aims to bypass some of the metacognitive processes involved in other paradigms, offering a more direct and objective measure of how illusions influence perceptual judgment.**
+Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a **cognitive top-down** fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
 If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".
@@
 To encourage the empirical exploration of our ability and tendency to alter our phenomenological experience and further accelerate investigations away from the hypnotic context, @lush2021 adapted the Sussex-Waterloo Scale of Hypnotisability [SWASH, @lush2018] by removing all its references to hypnosis, to measure trait phenomenological control.
-**This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion on a 6-point Likert scale (from 0-5) and has demonstrated validity in online experiments [@lush2022].**
+**This newly developed phenomenological control scale (PCS) consists of 10 imaginative suggestions followed by subjective ratings for each suggestion and has demonstrated validity in online experiments [@lush2022].**
 
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
+**Note that both prior-knowledge and phenomenological control are considered top-down processes, but the cognitive impenetrability hypothesis suggests that the processes at stake for the illusions happen at a lower- encapsulated- level (in the form of *perceptual* priors)**.
 The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion paradigm that encompasses other illusion types (see @tbl-table1).
+**Additionally, we will explore the relationship between psychoticism, as a proxy for schizophrenia, and illusion sensitivity to assess the potential impact of lower-level effects—such as weak priors observed in individuals with schizophrenia [@costa2023]—on sensitivity to illusions.**
+**These analyses may offer evidence clarifying whether inter-individual variability in illusion sensitivity is driven by lower-level perceptual mechanisms or higher-level cognitive processes (see Table 1).**
 
@@
 
-The task is an adaptation of the one used in @makowski2023 to make it shorter and more reliable, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
+The task is an adaptation of the one used in @makowski2023 to make it shorter, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
 It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer, and Vertical-Horizontal.
+**In the original Illusion Game, 10 visual illusions were presented in two sets, following a practice trial, and separated by two short questionnaires.
+Participants completed a total of 1,340 trials, with the experiment lasting approximately 55 minutes.**
+**In the current procedure, only three illusions are used, selected based on the original study's findings that these illusions most strongly contribute to illusion sensitivity.**
 The procedure encompasses 2 sets of 80 trials for each illusion type. 
 Each set will include, in a random order, the 3 blocks of illusion types, in which trials are separated by a fixation cross, temporally (uniformly sampled duration of 500 - 1000s) and spatially jittered (around the centre of the screen in a radius of a 1 cm) to attenuate its potential usefulness as a reference point.
-After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities. 
+After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score; IES) as a gamification mechanism to increase motivation to perform to the best of one's abilities. 
 To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors.
@@
 
-For each illusion type, two continuous dimensions are orthogonally manipulated [see @makowski2021 for details on the rationale and execution], namely task difficulty and illusion strength, so that each trial corresponds to a unique combination.
+For each illusion type, two continuous dimensions are orthogonally manipulated namely task difficulty and illusion strength, so that each trial corresponds to a unique combination, **providing an objectively correct answer for each task.**
+**The use of these manipulations allows concise, standardised reporting of illusion parameters and ensures our stimuli are fully reproducible [see @makowski2021].**
+
 Task difficulty corresponds to the difficulty of the perceptual decision (e.g., if the task is to select the longest red line, task difficulty corresponds to how the lines are objectively different). 
 Illusion strength corresponds to the degree to which the illusion elements (e.g., the black arrow lines in Müller-Lyer) are interfering with the aforementioned task. 
-Note that the illusion effect can be "incongruent" (biasing perception in the direction of the incorrect response) or "congruent" (facilitating, i.e., biasing perception in the direction of the correct response).
+Note that the illusion effect can be either “incongruent”, **making the task more difficult by biasing perceptual decisions toward the incorrect response** or “congruent”, **making the task easier by biasing decisions toward the correct response (e.g., in the Muller-Lyer the longer line has outward-facing arrowheads making the difference between the lines becomes more apparent)**.
 Participants respond with a key arrow (left vs. right; or up vs. down), and their reaction time (RT) and accuracy are recorded.
 
-Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, separately for the 3 illusion types. 
+**To assess whether the illusions functioned as expected, stimuli will be categorized into three groups: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.** 
+**The Inverse Efficiency Score [IES; @townsend2014methods], which accounts for both speed and accuracy, will be calculated by dividing the average reaction time of correct responses by the proportion of correct responses.**
+**Significant differences in IES across these groups are expected and will serve as evidence that the illusion operated as intended.**
+
+Visual illusion sensitivity will be measured as the average error rate in the incongruent condition **(e.g., where the task of discriminating between lines was made more difficult due to the direction for the arrow-heads, in the Muller-Lyer Illusion)**, separately for the 3 illusion types.
 Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy trade off), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current **registered report**.
@@
 These questionnaires are included as a way of providing a break between the two cognitively taxing blocks and maintain paradigmatic consistency with previous studies [@makowski2023].
+**Secondly, the PID-5 will be used to examine the correlation between maladaptive traits and illusion sensitivity, evaluating the existence of the link proposed in previous studies [@costa2023].**
 
@@
 
-The PCS will contain several manipulation check indices to identify problematic participants. 
+**Reliability of the PCS will be assessed by computing the Omega coefficient [@revelle2019reliability] based on the polychoric correlation matrix.** 
+**Additionally,** the PCS will contain several manipulation check indices to identify problematic participants. 
 **The phenomenological control task consists of various auditory and visual exercises.**
-**In one such exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
+**At the start of the task, participants first hear someone say "hello." They are then asked to choose from several options, including "Hello," "Goodbye," "How are you," and "Thank you".**
+**Any participant who selects an option other than the correct one will be considered inattentive**
+**In another exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
 **However, three differently coloured balls are actually displayed.**
@@
 
+**The reliability of illusion sensitivity scores will be assessed using two indices.**
+**First, split-half reliability will be computed, to assess internal consistency, by correlating two equal subsets of individual scores, with high correlations expected - following [@cohen2013statistical] criteria.**
+**Correlations across the three illusions will be computed to assess inter-illusion reliability for the following groups: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.**
+**If the correlations are high between the Strong Illusion Strength and Mild Illusion Strength groups, they will be recomputed as a single score for the subsequent analysis.**
 Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023]. 
@@
 In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
+**Additionally, to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean; based on @makowski2023) will be excluded from the analysis.**
 
@@
 Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity. 
+**Additionally, Bayesian correlations will be computed using the BayesFactor package, employing a medium prior on the coefficient (r-scale parameter set to 1/3) to assess relationships between maladaptive trait facets and illusion sensitivity scores.**
+**Based on prior research (Makowski et al., 2023), we expect to find evidence (BF10 ≥ 3) supporting a relationship between the psychoticism facet of the PID-5 and illusion sensitivity.**
 Data analysis will be carried out using R, using *tidyverse* [@tidyverse] and *easystats* [@easystats; @parameters; @correlation; @bayestestR; @datawizard]. 

### 2025-04-17 analysis/preprocessing.R
finished pre-processing script
@DominiqueMakowski could you check this when you have time please

@@ -1,0 +1,233 @@
+library(jsonlite)
+library(progress)
+
+path <- "C:/Users/asf25/Box/IllusionPCS/"
+
+# JsPsych experiment ------------------------------------------------------
+
+files <- list.files(path, pattern = "*.csv")
+
+# Progress bar
+progbar <- progress_bar$new(total = length(files))
+
+alldata_sub <- data.frame()
+alldata_ig <- data.frame()
+
+for (file in files) {
+  progbar$tick()
+  
+  # file <- "tnr9kq0qja.csv"
+  rawdata <- read.csv(paste0(path, "/", file))
+  
+  # Initialize participant-level data
+  dat <- rawdata[rawdata$screen == "browser_info", ]
+  
+  data_ppt <- data.frame(
+    Participant = sub("\\.csv$", "", file),
+    Recruitment = dat$researcher,
+    # Condition = dat$condition,
+    Experiment_StartDate = as.POSIXct(paste(dat$date, dat$time), format = "%d/%m/%Y %H:%M:%S"),
+    # Experiment_Duration = rawdata[rawdata$screen == "demographics_debrief", "time_elapsed"] / 1000 / 60,
+    Browser_Version = paste(dat$browser, dat$browser_version),
+    Mobile = dat$mobile,
+    Platform = dat$os,
+    Screen_Width = dat$screen_width,
+    Screen_Height = dat$screen_height
+  )
+  
+  
+  if("prolific_id" %in% colnames(dat)){
+    data_ppt$Prolific_ID <- dat$prolific_id
+  }
+  
+  
+  data_ppt$Reward <- rawdata[rawdata$screen == "demographics_debrief", "Reward"]
+  
+  # Demographics
+  demog <- jsonlite::fromJSON(rawdata[rawdata$screen == "demographic_questions", ]$response)
+  
+  #Education
+  demog$Education <- ifelse(demog$Education == "other", demog$`Education-Comment`, demog$Education)
+  demog$`Education-Comment` <- NULL
+  
+  # #Discipline
+  demog$Discipline <- ifelse(!is.null(demog$Discipline), demog$Discipline, NA)
+  demog$Discipline <- ifelse(demog$Discipline == "other", demog$`Discipline-Comment`, demog$Discipline)
[…]

### 2025-04-17 analysis/1_cleaning.qmd
finished pre-processing script
@DominiqueMakowski could you check this when you have time please
finished data_cleaning script

@@
 
-## Data Preparation
+# Data Preparation
 
@@
 library(patchwork)
+
+data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
+
+data_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")|> 
+  group_by(Illusion_Type) |> 
+  mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+         Task_Difficulty = abs(Illusion_Difference),
+         Condition_Illusion = datawizard::categorize(
+           Illusion_Strength, split="quantile", n_groups=4,
+           labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
+         Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+  ungroup()
+
 ```
+
+The initial sample consisted of `r report::report_participants(data_ppt, age="Age", gender="Gender", education="Education")`, for a total trial number of `r nrow(data_ppt)`.
+
+```{r}
+
+```
+
+## Score Computation
+
+### IPIP6
+
+```{r}
+SD_per_dim <- function(x, dims="") {
+  m <- matrix(nrow=nrow(x), ncol=0)
+  for(s in dims) {
+    m <- cbind(m, sapply(as.data.frame(t(x[grepl(s, names(x))])), sd))
+  }
+  m
+}
+```
+
+```{r}
+#| code-fold: false
+
+ipip6 <- select(data_ppt, starts_with("ipip_"))
+ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]
+
+data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("ext", names(ipip6))])
+data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("cons", names(ipip6))])
+data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("neuro", names(ipip6))])
+data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("open", names(ipip6))])
+data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("honhum", names(ipip6))])
+data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("agr", names(ipip6))])
+data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))
+
+# data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Ext", names(ipip6))])
+# data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("Cons", names(ipip6))])
+# data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("Neuro", names(ipip6))])
+# data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("Ppen", names(ipip6))])
+# data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("HonHum", names(ipip6))])
+# data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("Agr", names(ipip6))])
+# data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))
+```
+
+### PID-5
+
+```{r}
+#| code-fold: false
+
+pid5 <- select(data_ppt, starts_with("pid_"))
+
+data_ppt$PID5_Disinhibition <- rowMeans(pid5[grepl("Dis", names(pid5))])
+data_ppt$PID5_Detachment <- rowMeans(pid5[grepl("Det", names(pid5))])
+data_ppt$PID5_NegativeAffect <- rowMeans(pid5[grepl("NegAff", names(pid5))])
+data_ppt$PID5_Antagonism <- rowMeans(pid5[grepl("Ant", names(pid5))])
+data_ppt$PID5_Psychoticism <- rowMeans(pid5[grepl("Psy", names(pid5))])
+data_ppt$PID5_SD <- rowMeans(SD_per_dim(pid5, c("Disinhibition", "Detachment", "NegativeAffect",
+                                             "Antagonism", "Psychoticism")))
+```
+
+### Experiment Duration
+
+The experiment's median duration is `r insight::format_value(median(data_ppt$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(data_ppt$Experiment_Duration, ci=0.5))`).
+
+```{r}
+data_ppt |>
+  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
+         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
+         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
+  ggplot(aes(y = Participant, x = Duration)) +
+  geom_point(aes(color = Category, shape = Category)) +
+  geom_vline(xintercept = median(data_ppt$Experiment_Duration), color = "red", linetype = "dashed") +
+  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
+  scale_color_manual(values = c("extra" = "red", ok = "black")) +
+  guides(color = "none", shape = "none") +
+  ggside::geom_xsidedensity(fill = "grey", color=NA) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
+  labs(
+    title = "Experiment Completion Time",
+    x = "Duration (in minutes)",
+    y = "Participants"
+  )  +
+  see::theme_modern() +
+  ggside::theme_ggside_void() +
+  theme(ggside.panel.scale = .3, 
+        axis.text.y = element_blank()) 
+```
+
+# Exclusion
+
+```{r}
+#| echo: true
+#| code-fold: false
+
+outliers <- list()
+outliers_half <- list()
+```
+
+## Reaction Time (per Block)
+
+```{r}
+#| message: false
+
+errorrate <- data_ig |>
+  group_by(Participant, Illusion_Type, Block) |>
+  summarize(ErrorRate_per_block = sum(Error) / n()) |>
+  ungroup() |>
+  arrange(desc(ErrorRate_per_block))
+
+d_all <- estimate_density(filter(data_ig, RT < 3)$RT)
+
+plot_distribution <- function(dat) {
+
+  data_error <- errorrate |>
+    filter(Participant %in% unique(dat$Participant)) |>
+    group_by(Participant, Block) |>
+    summarize(y = mean(ErrorRate_per_block), .groups="drop") |>
+    mutate(x = ifelse(Block == "A", 2.1, 2.3),
+           color = case_when(
+              Participant %in% outliers ~ "red",
+              Participant %in% outliers_half ~ "orange",
+              TRUE ~ "blue"
+            ))
+
+  dat |>
+    filter(RT < 3) |>
+    estimate_density(select = "RT", at = c("Participant", "Block")) |>
+    group_by(Participant) |>
+    normalize(select = "y") |>
+    ungroup() |>
+    mutate(
+      Participant = fct_relevel(Participant, sort(unique(dat$Participant))),
+      color = case_when(
+        Participant %in% outliers ~ "red",
+        Participant %in% outliers_half ~ "orange",
+        TRUE ~ "blue"
+      )
+    ) |>
+    ggplot(aes(x = x, y = y)) +
+    geom_bar(data = data_error, aes(fill = color), stat = "identity", width=0.19) +
+    geom_segment(aes(x = 2, xend = 2.4, y = 0.5, yend = 0.5), color = "black", linetype="dashed", linewidth = 0.5) +
+    geom_area(data = normalize(d_all, select = "y"), alpha = 0.2) +
+    geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block), linewidth = 0.8) +
+    # geom_vline(xintercept = 0.125, linetype = "dashed", color = "red", size = 0.5) +
+    scale_color_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
+    scale_fill_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
+    scale_x_continuous(expand = c(0, 0), breaks=c(0, 0.5, 1, 1.5, 2), labels=c("0", "0.5", "1", "1.5", "2")) +
+    scale_y_continuous(expand = c(0, 0)) +
+    coord_cartesian(xlim = c(0, 2.4)) +
+    theme_modern() +
+    theme(axis.text.y = element_blank(),
+          axis.line.y = element_blank(),
+          strip.text.x = element_text(size = rel(1.5)),
+          legend.position = "none") +
+    facet_wrap(~Participant, nrow=10) +
+    labs(y = "", x = "Reaction Time (s)")
+}
+```
+
+### 1-100
+
+```{r}
+#| fig-width: 15
+#| fig-height: 15
+#| eval: false
+
+plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:100],])
+```
+
+### 101-200
+
+```{r}
+#| fig-width: 15
+#| fig-height: 15
+#| eval: false
+
+plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:200],])
+
+```
+
+:::
+
+We discarded `r length(outliers)` participants (entirely) and `r length(outliers_half)` participant's second blocks.
+
+```{r}
+# data_ig <- data_ig |> 
+#   filter(!Participant %in% outliers) |> 
+#   filter(!((Block == "B") & (Participant %in% outliers_half)))
+```
+
+## Error Rate (per Block)
+
+For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).
+
+```{r p_outliers_errorblock}
+#| eval: false
+
+errorrate |>
+  estimate_density(at = c("Illusion_Type", "Block"), method = "KernSmooth") |>
+  ggplot(aes(x = x, y = y)) +
+  geom_line(aes(color = Illusion_Type, linetype = Block)) +
+  geom_vline(xintercept = 0.5, linetype = "dashed") +
+  scale_x_continuous(expand = c(0, 0), labels = scales::percent) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(values = c("Ebbinghaus" = "#2196F3", "MullerLyer" = "#4CAF50", "VerticalHorizontal" = "#FF5722")) +
+  labs(y = "Distribution", x = "Error Rate") +
+  theme_modern()
+
+remove_badblocks <- function(df) {
+  n <- nrow(df)
+  df <- df |>
+    group_by(Participant, Illusion_Type, Block) |>
+    mutate(ErrorRate_per_block = sum(Error) / n()) |>
+    ungroup() |>
+    filter(ErrorRate_per_block < 0.5) |>
+    select(-ErrorRate_per_block)
+
+  text <- paste0(
+    "We removed ",
+    n - nrow(df),
+    " (",
+    insight::format_value((n - nrow(df)) / n, as_percent = TRUE),
+    ") trials belonging to bad blocks."
+  )
+  list(data = df, text = text)
+}
+
+out <- remove_badblocks(data_ig)
+print(paste("Illusion (session 1):", out$text))
+data_ig <- out$data
+```
+
+## Reaction Time (per Block)
+
+```{r outliers_RTtrials}
+#| eval: false
+
+check_trials <- function(df) {
+  data <- df |>
+    mutate(Outlier = ifelse(RT >= 10, TRUE, FALSE)) |>
+    group_by(Participant) |>
+    mutate(Outlier = ifelse(RT < 0.150 | standardize(RT, robust = TRUE) > 4, TRUE, Outlier)) |>
+    ungroup()
+
+  p1 <- data |>
+    filter(RT < 10) |>
+    estimate_density(select = "RT", at = "Participant") |>
+    group_by(Participant) |>
+    normalize(select = "y") |>
+    ungroup() |>
+    merge(data |>
+      group_by(Participant) |>
+      mutate(Threshold = median(RT) + 4 * mad(RT)) |>
+      filter(Error == 0) |>
+      summarize(Threshold = mean(Threshold))) |>
+    mutate(Outlier = ifelse(x >= Threshold, TRUE, FALSE)) |>
+    ggplot(aes(x = x, y = y)) +
+    geom_area(data = normalize(estimate_density(filter(data, RT < 10), select = "RT"), select = "y"), alpha = 0.2) +
+    geom_line(aes(color = Participant, linetype = Outlier), alpha = 0.2) +
+    geom_vline(xintercept = c(125), linetype = "dashed", color = "red") +
+    scale_color_material_d("rainbow", guide = "none") +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    guides(linetype = "none") +
+    coord_cartesian(xlim = c(0, 5)) +
+    theme_modern() +
+    theme(axis.text.y = element_blank()) +
+    labs(y = "", x = "Reaction Time (s)")
+
+  p2 <- data |>
+    group_by(Participant) |>
+    summarize(Outlier = sum(Outlier) / n()) |>
+    mutate(Participant = fct_reorder(Participant, Outlier)) |>
+    ggplot(aes(x = Participant, y = Outlier)) +
+    geom_bar(stat = "identity", aes(fill = Participant)) +
+    scale_fill_material_d("rainbow", guide = "none") +
+    scale_x_discrete(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
+    see::theme_modern() +
+    theme(axis.text.x = element_blank()) +
+    labs(y = "Percentage of outlier trials")
+
+  text <- paste0(
+    "We removed ",
+    sum(data$Outlier),
+    " (",
+    insight::format_value(sum(data$Outlier) / nrow(data), as_percent = TRUE),
+    ") outlier trials (150 ms < RT < 4 MAD above median)."
+  )
+
+  data <- filter(data, Outlier == FALSE)
+  data$Outlier <- NULL
+
+  list(p = p1 / p2, data = data, text = text)
+}
+```
+
+```{r p_outliers_RTtrials}
+#| results: 'asis'
+#| eval: false
+
+out <- check_trials(data_ig)
+out$text
+out$p
+data_ig <- out$data
+```
+
+## Manipulation/Attention Checks
+
+```{r}
+
+outliers$Hello <- data_ppt |>
+  dplyr::filter(pcs_hello != "Hello")
+
+outliers$Hello <- outliers$Hello$Participant
+
+```
+
+We removed `r length(outliers$Hello)` (`r insight::format_percent(length(outliers$Hello) / nrow(data_ppt))`) participants that did not reported hearing "Hello" at the beggining of the experiment.
+
+```{r}
+# remove after more pilots
+outliers$Balls <- data_ppt |>
+  dplyr::filter(pcs_noballs == TRUE) # remove after more pilots
+
+# outliers$balls <- data_ppt |>
+#   dplyr::filter(pcs_noballs == "TRUE")
+
+outliers$Balls <- outliers$Balls$Participant
+
+```
+
+We removed `r length(outliers$Balls)` (`r insight::format_percent(length(outliers$Balls) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.
+
+```{r}
+
+outliers$Press <- data_ppt |>
+  dplyr::filter(pcs_press < 5)
+
+outliers$Press <- outliers$Press$Participant
+```
+
+We removed `r length(outliers$Press)` (`r insight::format_percent(length(outliers$Press) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.
+
+```{r}
+data_ppt <- filter(data_ppt, !Participant %in% outliers$Hello)
+data_ppt <- filter(data_ppt, !Participant %in% outliers$Balls)
+data_ppt <- filter(data_ppt, !Participant %in% outliers$Press)
+
+data_ig <- filter(data_ig, !Participant %in% outliers$Hello)
+data_ig <- filter(data_ig, !Participant %in% outliers$Balls)
+data_ig <- filter(data_ig, !Participant %in% outliers$Press)
+```
+
+# Final Sample
+
+```{r}
+# data_ppt<- filter(data_ppt, !Participant %in% outliers)
+# data_ig <- filter(data_ig, Participant %in% data_ppt$Participant)
+```
+
+### Age
+
+```{r}
+p_age <- estimate_density(data_ppt$Age) |>
+  normalize(select = y) |>
+  mutate(y = y * 86) |>  # To match the binwidth
+  ggplot(aes(x = x)) +
+  geom_histogram(data=dfsub, aes(x = Age, fill=Gender), bins=28) +
+  # geom_line(aes(y = y), color = "orange", linewidth=2) +
+  geom_vline(xintercept = mean(dfsub$Age), color = "red", linewidth=1.5) +
+  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
+  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
+  theme_modern(axis.title.space = 10) +
+  theme(
+    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
+    axis.text.y = element_text(size = rel(1.1)),
+    axis.text.x = element_text(size = rel(1.1)),
+    axis.title.x = element_blank()
+  )
+p_age
+```
+
+### Education
+
+```{r}
+p_edu <- data_ppt |>
+  mutate(Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |>
+  ggplot(aes(x = Education)) +
+  geom_bar(aes(fill = Education)) +
+  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
+  scale_fill_viridis_d(guide = "none") +
+  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
+  theme_modern(axis.title.space = 15) +
+  theme(
+    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
+    axis.text.y = element_text(size = rel(1.1)),
+    axis.text.x = element_text(size = rel(1.1)),
+    axis.title.x = element_blank()
+  )
+p_edu
+```
+
+### Ethnicity
+
+```{r p_ethnicity}
+ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
+  ggplot(aes(x, y, fill = group)) + 
+  ggwaffle::geom_waffle() + 
+  coord_equal() + 
+  scale_fill_flat_d() + 
+  ggwaffle::theme_waffle() +
+  labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
+  theme(
+    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+    plot.subtitle = element_text(size = rel(1.2)),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+```
+
+### Summary
+
+The final sample includes `r report::report_participants(dfsub, age="Age", sex="Sex", education="Education")`.
+
+## Save
+
+```{r}
+#| code-fold: false
+
+write.csv(data_ppt, "../data/data_participants.csv", row.names = FALSE)
+write.csv(data_ig, "../data/data_IllusionGame.csv", row.names = FALSE)
+```

### 2025-04-17 analysis/preprocessing.R

@@ -5,7 +5,8 @@ path <- "C:/Users/asf25/Box/IllusionPCS/"
 
 # JsPsych experiment ------------------------------------------------------
 
-files <- list.files(path, pattern = "*.csv")
+# files <- list.files(path, pattern = "*.csv")
+files <- "z9650zj48m.csv"
 
 # Progress bar
 progbar <- progress_bar$new(total = length(files))
@@ -15,8 +16,6 @@ alldata_ig <- data.frame()
 
 for (file in files) {
   progbar$tick()
-  
-  # file <- "tnr9kq0qja.csv"
   rawdata <- read.csv(paste0(path, "/", file))
   
   # Initialize participant-level data
@@ -27,7 +26,7 @@ for (file in files) {
     Recruitment = dat$researcher,
     # Condition = dat$condition,
     Experiment_StartDate = as.POSIXct(paste(dat$date, dat$time), format = "%d/%m/%Y %H:%M:%S"),
-    # Experiment_Duration = rawdata[rawdata$screen == "demographics_debrief", "time_elapsed"] / 1000 / 60,
+    Experiment_Duration = rawdata[rawdata$screen == "waitdatasaving", "time_elapsed"] / 1000 / 60,
     Browser_Version = paste(dat$browser, dat$browser_version),
     Mobile = dat$mobile,
     Platform = dat$os,
@@ -40,9 +39,6 @@ for (file in files) {
     data_ppt$Prolific_ID <- dat$prolific_id
   }
   
-  
-  data_ppt$Reward <- rawdata[rawdata$screen == "demographics_debrief", "Reward"]
-  
   # Demographics
   demog <- jsonlite::fromJSON(rawdata[rawdata$screen == "demographic_questions", ]$response)
   
@@ -50,11 +46,14 @@ for (file in files) {
   demog$Education <- ifelse(demog$Education == "other", demog$`Education-Comment`, demog$Education)
   demog$`Education-Comment` <- NULL
   
-  # #Discipline
+  #Discipline
   demog$Discipline <- ifelse(!is.null(demog$Discipline), demog$Discipline, NA)
[…]

### 2025-04-21 analysis/1_cleaning.qmd
minor
changes to analysis scripts

@@
 
-data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
+# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
+
+data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_participants.csv")
 
@@
                                              "Antagonism", "Psychoticism")))
+```
+
+### PCS
+
+```{r}
+
+pcs <-  select(data_ppt, starts_with("pcs_"))
+
+data_ppt$pcs_tastescore <- rowMeans(pcs[grepl("taste", names(pcs))])
+
+data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
+
+#delete with more participants (changes have been made in script)
+pcs <- pcs |>
+  rename(pcs_amnesia = pcs_rememeber)
+
+# data_ppt$pcs <- rowMeans(pcs[grepl("hands|music|mosquito|tastescore|arm|negative|postsession|amnesia(?!_w)", names(pcs))], na.rm = TRUE)
+
+data_ppt$pcs <- rowMeans(pcs[!grepl("pcs_amnesia_w", names(pcs)) & grepl("hands|music|mosquito|tastescore|arm|negative|postsession|amnesia", names(pcs))], na.rm = TRUE)
+
+data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
 ```

### 2025-04-21 analysis/2_analysis.qmd
changes to analysis scripts

---
title: "Illusion"
subtitle: "Illusion Task"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

logmod <- function(x) sign(x) * log(1 + abs(x))
sqrtmod <- function(x) sign(x) * sqrt(abs(x))
cbrtmod <- function(x) sign(x) * (abs(x)**(1 / 3))

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv") |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
  ) 
```

## PCS

### Compute subjective scale score

```{r}

```

### Reliability

## Illusion

### Reliability

### 2025-04-21 analysis/preprocessing.R

@@ -6,7 +6,7 @@ path <- "C:/Users/asf25/Box/IllusionPCS/"
 # JsPsych experiment ------------------------------------------------------
 
 # files <- list.files(path, pattern = "*.csv")
-files <- "z9650zj48m.csv"
+files <- "ryqwk2oej3.csv"
 
 # Progress bar
 progbar <- progress_bar$new(total = length(files))
@@ -17,6 +17,7 @@ alldata_ig <- data.frame()
 for (file in files) {
   progbar$tick()
   rawdata <- read.csv(paste0(path, "/", file))
+  
   
   # Initialize participant-level data
   dat <- rawdata[rawdata$screen == "browser_info", ]
@@ -69,21 +70,21 @@ for (file in files) {
   data_ppt$pcs_magnetichands <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_magnetichands_r", "response"])))
   data_ppt$pcs_mosquito <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_mosquito_r", "response"])))
   taste <- jsonlite::fromJSON(rawdata$response[rawdata$screen == "pcs_taste_r"])
-  data_ppt$tastesweet <- taste$TasteSweet_r
-  data_ppt$tastesour <- taste$TasteSour_r
+  data_ppt$pcs_tastesweet <- taste$TasteSweet_r
+  data_ppt$pcs_tastesour <- taste$TasteSour_r
   data_ppt$pcs_armrigidity <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_armrigidity_r", "response"])))
   data_ppt$pcs_armimmobile <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_armrimmobile_r", "response"])))
   data_ppt$pcs_music <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_music_r", "response"])))
   
-  # data_ppt$pcs_negativehallucination <- jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_negativehallucination_r", "response"])
-  data_ppt$pcs_negativehallucination <- as.integer(unlist(jsonlite::fromJSON(rawdata[rawdata$screen == "pcs_negativehallucination", "response"])))
[…]

### 2025-04-22 paper/R1/manuscript.qmd
changes made to manuscript

--- 2025-03-24 paper/R1/manuscript.qmd
@@
 Their robust and reliable effect makes them useful stimuli to explore how perception is constructed and shaped, and several theoretical models have been put forth to explain how they work. 
-**In particular, illusions have been reframed using a predictive coding account of perception [@notredame2014] in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].**
+**In particular, illusions have been reframed using a predictive coding account of perception [e.g., @notredame2014; @nour2015perception]** in which the brain optimally combines, using some flavour of Bayesian inference, perceptual inputs with prior knowledge to make sense of ambiguous environments [@friston2010].
 
@@
 
+```{r}
+#| warning: false
+#| label: fig-illusion_framework
+#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
+#| out-width: "100%"
+#| fig-cap: "The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion.
+Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."
+
+knitr::include_graphics("img/IG_framework.png")
+```
+
 The "Illusion Game" paradigm [@makowski2023] has been recently developed to measure illusion sensitivity to various illusion types through its behavioural impact (on response time and error rate) in a perceptual decision task (where participants have to respond as fast as possible; e.g., "which of the left or right circles is bigger").
-The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
-This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its interference effect, hopefully bypassing some of the metacognitive processes at stake in other paradigms.
+The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task (see @fig-illusion_framework).
+This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its **interference effect —an effect that disrupts an individual's ability to accurately discriminate between perceptual stimuli.**
+**This approach aims to bypass some of the metacognitive processes involved in other paradigms, offering a more direct and objective measure of how illusions influence perceptual judgment.**
 
-Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a **top-down** fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
+Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters).
 If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".
@@
 This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.
-The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion paradigm that encompasses other illusion types (see @tbl-table1).
+**Note that both prior-knowledge and phenomenological control are considered top-down processes, but the cognitive impenetrability hypothesis suggests that the processes at stake for the illusions happen at a lower- encapsulated- level (in the form of *perceptual* priors)**.
 
-::: {#tbl-table1}
+The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion paradigm that encompasses other illusion types.
+**Additionally, we will explore the relationship between psychoticism, as a proxy for schizophrenia, and illusion sensitivity to assess the potential impact of lower-level effects—such as weak priors observed in individuals with schizophrenia [@costa2023]—on sensitivity to illusions.**
+**These analyses may offer evidence clarifying whether inter-individual variability in illusion sensitivity is driven by lower-level perceptual mechanisms or higher-level cognitive processes (@tbl-DesignTable).**
+
 ```{r}
 #| tbl-cap: "Study Design Table"
+#| label: tbl-DesignTable
 
 question <- c("Question", "Hypothesis", "Sampling Plan", "Analysis Plan", "Rationale for Deciding the Sensitivity of the Test", "Interpretation Given Different Outcomes", "Theory That Could Be Shown Wrong by the Outcomes")
-answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?", "Replicating findings from Lush et al., 2022 paper, we expect evidence to support the absence of a relationship between VI and PCS", "The goal is to recruit around 500 adult English speakers using Prolific.
-This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to ‘medium’)", "Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings, for all three illusions.
-BF10 > 3 will be interpreted as evidence for a relationship between these two measures.", "The hypothesis that VI sensitivity is independent from PC.", "The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC.")
+answer <- c("Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?
+Additionally, is there a relationship between VI sensitivity and the psychoticism facet of the PID-5, as a proxy for schizophrenia-related traits?", "In line with Lush et al.
+(2022), we hypothesise that there will be evidence supporting the absence of a relationship between PC and VI sensitivity.
+Based on Makowski et al.
+(2023) and prior work on weak priors in schizophrenia, we hypothesise that higher psychoticism scores will be positively associated with VI sensitivity.", "The goal is to recruit around 500 adult English speakers using Prolific.
+This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate.", "Bayesian correlations will be conducted using the BayesFactor::correlationBF() function, with a medium prior (r-scale = 1/3), separately for: 1) PC scores and VI sensitivity (error rate and IES), across all three illusion types; and 2) Psychoticism facet scores from the PID-5 and VI sensitivity scores, across all three illusion types.", "For the PC–VI sensitivity relationship, we will interpret BF₁₀ ≤ 1/3 as evidence against a relationship, in line with Lush et al.
+(2022).
+For the psychoticism–VI sensitivity relationship, BF₁₀ > 3 will be interpreted as evidence supporting a relationship, following findings by Makowski et al.
+(2023).", "If there is no evidence for a PC–VI relationship across all three illusions, it would support the hypothesis that VI sensitivity is independent from PC.
+If a positive association is found between psychoticism and VI sensitivity, it may suggest a low-level perceptual basis for inter-individual differences in illusion sensitivity.", "The cognitive impenetrability of visual illusions, which posits that illusion sensitivity is driven solely by low-level processes and is not influenced by top-down mechanisms such as phenomenological control.
+Conversely, a lack of association with psychoticism would challenge the view that low-level perceptual alterations underlie illusion sensitivity in non-clinical populations.")
 
@@
 ```
-
-Study Design Table
-:::
 
@@
 **Although these variables are not directly analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.** 
-Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.
+**Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.**
 
@@
 
-The task is an adaptation of the one used in @makowski2023 to make it shorter and more reliable, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
-It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer, and Vertical-Horizontal.
-The procedure encompasses 2 sets of 80 trials for each illusion type. 
-Each set will include, in a random order, the 3 blocks of illusion types, in which trials are separated by a fixation cross, temporally (uniformly sampled duration of 500 - 1000s) and spatially jittered (around the centre of the screen in a radius of a 1 cm) to attenuate its potential usefulness as a reference point.
-After each illusion type block, a score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities. 
-To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors.
+The task is an adaptation of the one used in @makowski2023 to make it shorter, in which participants must make perceptual judgments (e.g., "which red line is the longer") as quickly and accurately as possible.
+It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer, and Vertical-Horizontal (see @fig-illusionexample).
 
@@
 #| warning: false
-#| label: "fig-illusion_framework"
+#| label: fig-illusionexample
 #| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
 #| out-width: "100%"
-#| fig-cap: "The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion (above).
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."
+#| fig-cap: "The study involved three visual illusions, in which participants were instructed to respond as quickly as possible without making errors.
+Each illusion included two manipulated parameters: strength (e.g., the angle of the outward- or inward-pointing arrow-like fins in the Müller-Lyer illusion) and difficulty (e.g., the difference in line lengths in the Müller-Lyer illusion)."
 
-knitr::include_graphics("img/example_stimuli.png")
+knitr::include_graphics("img/IllusionTable.jpg")
 ```
 
-For each illusion type, two continuous dimensions are orthogonally manipulated [see @makowski2021 for details on the rationale and execution], namely task difficulty and illusion strength, so that each trial corresponds to a unique combination.
+**In the original Illusion Game, 10 visual illusions were presented in two sets, following a practice trial, and separated by two short questionnaires.**
+**Participants completed a total of 1,340 trials, with the experiment lasting approximately 55 minutes.**
+**In the current procedure, only three illusions are used, selected based on the original study's findings that these illusions most strongly contribute to illusion sensitivity.**
+
+The procedure encompasses 2 sets of 80 trials for each illusion type, **preceded by a practice trial for each illusion**. 
+Each set will include, in a random order, the 3 blocks of illusion types, in which trials are separated by a fixation cross, temporally (uniformly sampled duration of 500 - 1000s) and spatially jittered (around the centre of the screen in a radius of a 1 cm) to attenuate its potential usefulness as a reference point.
+After each illusion type block, an arbitrary score is presented (computed as a scaled Inverse Efficiency Score) as a gamification mechanism to increase motivation to perform to the best of one's abilities.
+To mitigate for the potential variability in the speed/accuracy trade-off, the instructions emphasize with equal weight to be fast and to avoid errors.
+
+For each illusion type, two continuous dimensions are orthogonally manipulated namely task difficulty and illusion strength, so that each trial corresponds to a unique combination, **providing an objectively correct answer for each task.**
+**The use of these manipulations allows concise, standardised reporting of illusion parameters and ensures our stimuli are fully reproducible [see @makowski2021].**
+
 Task difficulty corresponds to the difficulty of the perceptual decision (e.g., if the task is to select the longest red line, task difficulty corresponds to how the lines are objectively different). 
 Illusion strength corresponds to the degree to which the illusion elements (e.g., the black arrow lines in Müller-Lyer) are interfering with the aforementioned task. 
-Note that the illusion effect can be "incongruent" (biasing perception in the direction of the incorrect response) or "congruent" (facilitating, i.e., biasing perception in the direction of the correct response).
+Note that the illusion effect can be either “incongruent”, **making the task more difficult by biasing perceptual decisions toward the incorrect response** or “congruent”, **making the task easier by biasing decisions toward the correct response (e.g., in the Müller-Lyer illusion, if the outwards-facing arrowheads are placed on the longer line, identifying which line is the longest becomes easier)**.
 Participants respond with a key arrow (left vs. right; or up vs. down), and their reaction time (RT) and accuracy are recorded.
 
-Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, separately for the 3 illusion types. 
+Visual illusion sensitivity will be measured as the average error rate in the incongruent condition, and separately for the 3 illusion types. 
 Although the error rate is arguably a crude score, which does not take into account the effect of varying illusion strength, the interaction with task difficulty and the possible adjustments in response strategy (speed-accuracy trade off), it is also the most simple and easy to reproduce, hence its usage as our primary outcome for the current **registered report**.
+**As a secondary exploratory outcome, the Inverse Efficiency Score [IES, @townsend2014methods] will also be computed.
+This metric incorporates both speed and accuracy by dividing the mean reaction time of correct responses by the proportion of correct responses, separately for each illusion.**
 
@@
 These questionnaires are included as a way of providing a break between the two cognitively taxing blocks and maintain paradigmatic consistency with previous studies [@makowski2023].
-
-```{r}
-#| warning: false
-#| label: "fig-illusionexample"
-#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "100%"
-#| fig-cap: "The study involved three visual illusions, in which participants were instructed to respond as quickly as possible without making errors.
-Each illusion included two manipulated parameters: strength (e.g., the angle of the outward- or inward-pointing arrow-like fins in the Müller-Lyer illusion) and difficulty (e.g., the difference in line lengths in the Müller-Lyer illusion."
-
-knitr::include_graphics("img/IllusionTable.jpg")
-```
+**Additionally, the psychoticism subscale of the PID-5 will be used to examine the correlation between maladaptive traits and illusion sensitivity, evaluating the existence of the link proposed in previous studies [@costa2023].**
 
@@
 **The phenomenological control task consists of various auditory and visual exercises.**
-**In one such exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
+**At the start of the task, participants first hear someone say "hello." They are then asked to choose from several options, including "Hello," "Goodbye," "How are you," and "Thank you".**
+**Any participant who selects an option other than the correct one will be considered inattentive**
+**In another exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
 **However, three differently coloured balls are actually displayed.**
 **If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.**
-**In another exercise, participants are asked to press the spacebar six times.**
+**Lastly, in another exercise, participants are asked to press the spacebar six times.**
 **If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.**
-**Participants will be excluded if they fail at least one of these checks. Participants will be excluded if they fail at least one of the attention checks.**
+**Participants will be excluded if they fail at least one of these checks.**
+**Lastly, reliability of the PCS will be assessed by computing Cronbach's alpha [@cronbach1951coefficient].**
+
+**To assess whether the illusions functioned as expected, stimuli will be categorized into three groups: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.**
+**Two Bayesian t-tests will be conducted to assess differences in IES between the Congruent and Mild conditions, and between the Mild and Strong conditions.**
+**Significant differences in these comparisons will provide evidence that the illusions functioned as intended.**
+
+**The two outcome measures—error rate and IES—will be computed for each illusion and for each illusion strength group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.**
+**Correlation will be computed between the mild and strong groups for each illusion and outcomes separately.**
+**If these correlations are high [*r* > .50, @cohen2013statistical], the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.**
+
+**Reliability analyses will then be conducted on all resulting indices.** **First, split-half reliability, to assess internal consistency, will be computed by correlating two equal subsets of individual scores, with high correlations expected (*r* > .5).**
+**Second, inter-illusion reliability will be evaluated using Cronbach’s alpha across the three illusions.**
 
@@
 In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
+**To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean, based on @makowski2023) will be excluded from the analysis.**
 
-After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the *BayesFactor* package; @BayesFactor].
+After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the *BayesFactor* package, @BayesFactor].
+This shifted beta prior, as recommended by @morey2018baysefactor, offers a balanced approach to estimating effect sizes, without placing undue weight on larger effect sizes or artificially inflating evidence for the null hypothesis.
 Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity. 
+**Additionally, Bayesian correlations will be computed using the BayesFactor package, employing a medium prior on the coefficient (r-scale parameter set to 1/3) to assess relationships between maladaptive trait facets and illusion sensitivity scores.**
+**Based on prior research (Makowski et al., 2023), we expect to find evidence (BF10 ≥ 3) supporting a relationship between the psychoticism facet of the PID-5 and illusion sensitivity.**
 Data analysis will be carried out using R, using *tidyverse* [@tidyverse] and *easystats* [@easystats; @parameters; @correlation; @bayestestR; @datawizard]. 
-The analysis script and additional information are available at *https://github.com/RealityBending/IllusionGamePhenomenologicalControl*.
+The analysis script and additional information are available at **https://osf.io/da3u6/?view_only=247d4efa1afe456aa07662732946d4e6** [Note this link will be replaced with the GitHub page of the current project upon completion of the review process to ensure continued anonymisation].
 
@@
 
-All the study materials, experiment, data, and analysis is available on GitHub at https://github.com/RealityBending/IllusionGamePhenomenologicalControl
+All the study materials, experiment, data, and analysis is available on GitHub.
+[For the review process the pcs materials, the illusion game, and the analyses scripts can be accessed here: **https://osf.io/da3u6/?view_only=247d4efa1afe456aa07662732946d4e6**.
+Note this link will be replaced with the GitHub page of the current project upon completion of the review process to ensure continued anonymisation].
 

### 2025-04-22 paper/R1/manuscript.tex
changes made to manuscript

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[bidi=default]{babel}
\babelprovide[main,import]{english}

% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}

\RequirePackage{longtable}
% \setlength\LTleft{0pt}
\RequirePackage{threeparttablex}

% %

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[nolongtablepatch]{lineno}
\linenumbers

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{Testing the Relationship between Phenomenological Control related
to Illusion Sensitivity}
\shorttitle{Illusion Sensitivity and Phenomenological Control}

\usepackage{etoolbox}

\authorsnames{Dominique Makowski,Ana Neves}

\affiliation{
{School of Psychology, University of Sussex}}

\leftheader{Makowski and Neves}

\abstract{Visual illusions highlight how easily our conscious experience
can be altered with respect to perceptual reality. Despite sharing
in-principle mechanisms with phenomenological control, i.e., the ability
to alter our perceptual experience to match task demands or
expectations, research tying the two remains scarce. This study aims to
replicate and expand Lush et al.~(2022) reporting an absence of
correlation between phenomenological control (measured using the
Phenomenological Control Scale) and illusion sensitivity to different
illusion types. \emph{{[}N participants were recruited in an online
study. Results will be added in the final version of the
manuscript{]}}.}
% 
\keywords{illusion sensitivity, visual illusions, phenomenological
control, suggestibility, hypnotizability}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}
\par{ }
\par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Dominique
Makowski:   Conceptualization, Data curation, formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft, Writing -- review \& editing; Ana Neves:   Project
administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique Makowski, Email: D.Makowski@sussex.ac.uk}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\resetlinenumber[1]

Visual Illusions are an interesting type of stimuli highlighting the
ease with which our phenomenological conscious experience can become
dissociated from physical reality. Their robust and reliable effect
makes them useful stimuli to explore how perception is constructed and
shaped, and several theoretical models have been put forth to explain
how they work. \textbf{In particular, illusions have been reframed using
a predictive coding account of perception (e.g., Notredame et al., 2014;
Nour \& Nour, 2015)} in which the brain optimally combines, using some
flavour of Bayesian inference, perceptual inputs with prior knowledge to
make sense of ambiguous environments (Friston, 2010).

Such computational model(s) propose to conceptualize illusions as
stimuli providing weak or conflicting sensory evidence (Gershman et al.,
2012; Sundareswara \& Schrater, 2008) that bias perception toward prior
knowledge. In other words, the weight of priors, in the form of
perceptual knowledge about the world (e.g., internalized rules of
perspective) is amplified when the sensory input is confusing. For
instance, in the Müller-Lyer illusion, we ``compute'' the two (actually
identical) lines as being of different lengths because the line flanked
with converging fins is misinterpreted as being further away (Notredame
et al., 2014). In this context, measuring sensitivity to illusion can be
operationalized as indexing the parameters of the Bayesian inference
process (e.g., prior precision).

These accounts also provide a compelling framework to explain existing
findings reporting interindividual variability in the sensitivity to
illusions. Indeed, several studies suggest a potential link with
psychopathology, in particular schizophrenia (Costa et al., 2023) and
autism (Gori et al., 2016), in which the reported lower sensitivity to
illusions has been attributed to a diminished influence of top-down
processes such as prior knowledge (Mitchell et al., 2010) and a greater
emphasis on (i.e., precision of) sensory information (Palmer et al.,
2017). Evidence beyond psychopathology also suggests variability in the
general population, potentially correlated with personality traits such
as agreeableness and honest-humility (Makowski et al., 2023), as well as
cognitive abilities (Shoshina \& Shelepin, 2014).

However, the exact nature of this interindividual variability and its
potential origin remains unclear. The somewhat mixed evidence in the
literature regarding its generalizability and strength could be related
to the variety of the paradigms used and the type of processes being
mobilised (Makowski et al., 2021). Indeed, traditional methods
frequently focus on participant's experience by prompting them to assess
the difference between two identical targets, estimate the target's
physical properties, or adjust the targets to match a reference stimulus
(Todorović, 2020). Relying on metacognitive judgments about one's
subjective experiences adds an additional layer to the measure that
might not be desired when attempting to measure illusion
\textbf{sensitivity}. Moreover, paradigms often face challenges in
diversifying the illusory effects (i.e., using multiple stimuli to
experimentally manipulate the strength of the illusion) and the illusion
types (i.e., using various illusions, such as Müller-Lyer, Ebbinghaus,
Delboeuf which might rely on a different admixture of mechanisms),
hindering the potential of obtaining a comprehensive, valid, and
reliable measure of illusion sensitivity.

\begin{figure}

\caption{\label{fig-illusion_framework}The parametric framework for
visual illusions (Makowski et al., 2021) applied to the Müller-Lyer
illusion. Below are examples of stimuli showcasing the manipulation of
two parameters, task difficulty and illusion strength.}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{img/IG_framework.png}

}

\end{figure}%

The ``Illusion Game'' paradigm (Makowski et al., 2023) has been recently
developed to measure illusion sensitivity to various illusion types
through its behavioural impact (on response time and error rate) in a
perceptual decision task (where participants have to respond as fast as
possible; e.g., ``which of the left or right circles is bigger''). The
stimuli for different classical illusions are created using the
\emph{Pyllusion} software (Makowski et al., 2021), which allows
researchers to modulate the strength of the illusion as a continuous
dimension, independently from the difficulty of the perceptual task (see
Figure~\ref{fig-illusion_framework}). This paradigm, inspired by
psychophysics, lends itself to the computational modelling of illusion
sensitivity through its \textbf{interference effect ---an effect that
disrupts an individual's ability to accurately discriminate between
perceptual stimuli.} \textbf{This approach aims to bypass some of the
metacognitive processes involved in other paradigms, offering a more
direct and objective measure of how illusions influence perceptual
judgment.}

Interestingly, the fact that inter-individual variability in illusion
sensitivity seems to persist in this task suggests that it is not solely
explained by \textbf{metacognitive ability differences}, and gives rise
to the following question: is the variability in illusion sensitivity
related to low-level perceptual processes (e.g., baseline precision of
perceptual priors), or rather to the ability to actively control and
``resist'' the illusion in order to achieve the task at hand
(higher-level modulation of the perceptual inference parameters). If the
latter is true, then illusion sensitivity measured in contexts with
strong task-demand characteristics, e.g., in paradigms where
participants' performance is explicitly or implicitly assessed (i.e.,
where there is an incentive to downplay the illusion effect) might
correlate with one's ability to alter one's subjective experience
following suggestions - a mechanism referred to as ``phenomenological
control''.

The idea that we are endowed with the potential to unconsciously alter
our subjective experience and distort reality - even momentarily - to
meet the goals at hand is not novel. While this phenomenon has been
historically often studied under the label of ``hypnotisability'' - the
tendency to alter our conscious experience to match external demands
(Lush et al., 2021), the term ``phenomenological control'' (PC) has been
recently introduced to disconnect this concept from the potentially
negative associations with hypnosis and the misconception that a
hypnotic context is necessary for responding to imaginative suggestions
(Dienes et al., 2022).

To encourage the empirical exploration of our ability and tendency to
alter our phenomenological experience and further accelerate
investigations away from the hypnotic context, Lush et al. (2021)
adapted the Sussex-Waterloo Scale of Hypnotisability (SWASH, Lush et
al., 2018) by removing all its references to hypnosis, to measure trait
phenomenological control. \textbf{This newly developed phenomenological
control scale (PCS) consists of 10 imaginative suggestions followed by
subjective ratings for each suggestion and has demonstrated validity in
online experiments (Lush et al., 2022).}

Interestingly, Lush et al. (2022) did test for a relationship between PC
and illusion sensitivity using the Müller-Lyer illusion (in which the
arrangement of the arrowheads flanking two lines makes them appear as
having different lengths), and reported evidence in favour of an absence
of correlation between the two measures. This finding was interpreted as
indicative of the cognitive impenetrability of illusions, implying that
the effect is driven by low-level processes and therefore not influenced
by top-down mechanisms such as PC. \textbf{Note that both
prior-knowledge and phenomenological control are considered top-down
processes, but the cognitive impenetrability hypothesis suggests that
the processes at stake for the illusions happen at a lower-
encapsulated- level (in the form of \emph{perceptual} priors)}.

The goal of this study is thus to replicate the results from Lush et al.
(2022) pointing to an absence of a relationship between phenomenological
control and illusion sensitivity, by generalising them to a different
illusion paradigm that encompasses other illusion types.
\textbf{Additionally, we will explore the relationship between
psychoticism, as a proxy for schizophrenia, and illusion sensitivity to
assess the potential impact of lower-level effects---such as weak priors
observed in individuals with schizophrenia (Costa et al., 2023)---on
sensitivity to illusions.} \textbf{These analyses may offer evidence
clarifying whether inter-individual variability in illusion sensitivity
is driven by lower-level perceptual mechanisms or higher-level cognitive
processes (Table~\ref{tbl-DesignTable}).}

\begin{table}

{\caption{{Study Design Table}{\label{tbl-DesignTable}}}
\vspace{-20pt}}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable*}[c]{cc}

\ascline{0.75pt}{000000}{1-2}

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Question}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{Is\ there\ a\ correlation\ between\ trait\ phenomenological\ control\ (PC)\ and\ visual\ illusion\ (VI)\ sensitivity?\ Additionally,\ is\ there\ a\ relationship\ between\ VI\ sensitivity\ and\ the\ psychoticism\ facet\ of\ the\ PID-5,\ as\ a\ proxy\ for\ schizophrenia-related\ traits?}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Hypothesis}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{In\ line\ with\ Lush\ et\ al.\ (2022),\ we\ hypothesise\ that\ there\ will\ be\ evidence\ supporting\ the\ absence\ of\ a\ relationship\ between\ PC\ and\ VI\ sensitivity.\ Based\ on\ Makowski\ et\ al.\ (2023)\ and\ prior\ work\ on\ weak\ priors\ in\ schizophrenia,\ we\ hypothesise\ that\ higher\ psychoticism\ scores\ will\ be\ positively\ associated\ with\ VI\ sensitivity.}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Sampling\ Plan}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{The\ goal\ is\ to\ recruit\ around\ 500\ adult\ English\ speakers\ using\ Prolific.\ This\ sample\ size\ is\ based\ on\ the\ ones\ used\ in\ Lush\ et\ al.,\ 2021\ and\ Lush\ et\ al.,\ 2022\ that\ we\ aim\ at\ replicate.}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Analysis\ Plan}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{BBayesian\ correlations\ will\ be\ conducted\ using\ the\ BayesFactor::correlationBF()\ function,\ with\ a\ medium\ prior\ (r-scale\ =\ 1/3),\ separately\ for:\ 1)\ PC\ scores\ and\ VI\ sensitivity\ (error\ rate\ and\ IES),\ across\ all\ three\ illusion\ types.\ 2)\ Psychoticism\ facet\ scores\ from\ the\ PID-5\ and\ VI\ sensitivity\ scores,\ across\ all\ three\ illusion\ types.}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Rationale\ for\ Deciding\ the\ Sensitivity\ of\ the\ Test}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{For\ the\ PC–VI\ sensitivity\ relationship,\ we\ will\ interpret\ BF₁₀\ ≤\ 1/3\ as\ evidence\ against\ a\ relationship,\ in\ line\ with\ Lush\ et\ al.\ (2022).\ For\ the\ psychoticism–VI\ sensitivity\ relationship,\ BF₁₀\ >\ 3\ will\ be\ interpreted\ as\ evidence\ supporting\ a\ relationship,\ following\ findings\ by\ Makowski\ et\ al.\ (2023).}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Interpretation\ Given\ Different\ Outcomes}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{If\ there\ is\ no\ evidence\ for\ a\ PC–VI\ relationship\ across\ all\ three\ illusions,\ it\ would\ support\ the\ hypothesis\ that\ VI\ sensitivity\ is\ independent\ from\ PC.\ If\ a\ positive\ association\ is\ found\ between\ psychoticism\ and\ VI\ sensitivity,\ it\ may\ suggest\ a\ low-level\ perceptual\ basis\ for\ inter-individual\ differences\ in\ illusion\ sensitivity.}}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{\textbf{Theory\ That\ Could\ Be\ Shown\ Wrong\ by\ the\ Outcomes}}}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{22}\selectfont{\global\setmainfont{Times New Roman}{The\ cognitive\ impenetrability\ of\ visual\ illusions,\ which\ posits\ that\ illusion\ sensitivity\ is\ driven\ solely\ by\ low-level\ processes\ and\ is\ not\ influenced\ by\ top-down\ mechanisms\ such\ as\ phenomenological\ control.\ Conversely,\ a\ lack\ of\ association\ with\ psychoticism\ would\ challenge\ the\ view\ that\ low-level\ perceptual\ alterations\ underlie\ illusion\ sensitivity\ in\ non-clinical\ populations.}}}} \\

\ascline{0.75pt}{000000}{1-2}

\end{longtable*}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\end{table}

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

We aim to recruit around 500 (in line with the sample sizes used in Lush
et al., 2021; Lush et al., 2022) adult English native speakers with a
desktop device using Prolific (www.prolific.co). Participants will be
first presented with an explanatory statement and the consent form, and
can proceed by pressing a button to confirm they have read and
understood the information. This study has been approved by the ethics
board of the School of Psychology of the University of Sussex
(ER/ASF25/5).

\subsection{Procedure}\label{procedure}

The experiment's setup follows of the born-open principle (De Leeuw,
2023). The online experiment, implemented entirely using JsPsych (De
Leeuw, 2015), has its code stored on GitHub and will leverage the power
of the platform to host the experiment for free. Participant's raw data
files (containing identifiers) \textbf{are} automatically stored in a
private OSF repository. The preprocessing and analysis scripts, as well
as the anonymized data, will be available directly on GitHub, ensuring
the transparency and reproducibility of all the analysis steps.

Participants will be presented with a consent form followed by
demographic questions (gender, education level, age, and ethnicity).
\textbf{Although these variables are not directly analyzed in the
current study, they will be used to provide to provide a detailed and
thorough description of the sample and maximizing data reusability.}
\textbf{Participants will then be administered the PCS and the Illusion
Game task (IG) in a counterbalanced order.}

\subsubsection{Phenomenological Control Scale
(PCS)}\label{phenomenological-control-scale-pcs}

Participants will be asked to put on their headphones and await further
auditory instructions. The PCS procedure starts with a recorded
introduction explaining that a series of tests will be applied to
evaluate how experiences can be created through imagination. This will
be followed by 10 suggestions in a fixed order (see Lush et al., 2021),
such as ``now extend your arms ahead of you, with palms facing each
other, hands about a foot apart'' and ``as you sit comfortably in your
chair with your eyes closed, a picture of two balls will be displayed on
the computer screen''. \textbf{Once the 10 suggestions are completed,
participants will be asked to rate their subjective experiences and
response to each suggestion on a 6-points Likert scale (from 0-5).}
Phenomenological control will be indexed by averaging the scores from
the 10 scales.

\subsubsection{Illusion Game}\label{illusion-game}

The task is an adaptation of the one used in Makowski et al. (2023) to
make it shorter, in which participants must make perceptual judgments
(e.g., ``which red line is the longer'') as quickly and accurately as
possible. It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer,
and Vertical-Horizontal (see Figure~\ref{fig-illusionexample}).

\begin{figure}

\caption{\label{fig-illusionexample}The study involved three visual
illusions, in which participants were instructed to respond as quickly
as possible without making errors. Each illusion included two
manipulated parameters: strength (e.g., the angle of the outward- or
inward-pointing arrow-like fins in the Müller-Lyer illusion) and
difficulty (e.g., the difference in line lengths in the Müller-Lyer
illusion).}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{img/IllusionTable.jpg}

}

\end{figure}%

\textbf{In the original Illusion Game, 10 visual illusions were
presented in two sets, following a practice trial, and separated by two
short questionnaires. Participants completed a total of 1,340 trials,
with the experiment lasting approximately 55 minutes.} \textbf{In the
current procedure, only three illusions are used, selected based on the
original study's findings that these illusions most strongly contribute
to illusion sensitivity.}

The procedure encompasses 2 sets of 80 trials for each illusion type,
\textbf{preceded by a practice trial for each illusion}. Each set will
include, in a random order, the 3 blocks of illusion types, in which
trials are separated by a fixation cross, temporally (uniformly sampled
duration of 500 - 1000s) and spatially jittered (around the centre of
the screen in a radius of a 1 cm) to attenuate its potential usefulness
as a reference point. After each illusion type block, an arbitrary score
is presented (computed as a scaled Inverse Efficiency Score) as a
gamification mechanism to increase motivation to perform to the best of
one's abilities. To mitigate for the potential variability in the
speed/accuracy trade-off, the instructions emphasize with equal weight
to be fast and to avoid errors.

For each illusion type, two continuous dimensions are orthogonally
manipulated namely task difficulty and illusion strength, so that each
trial corresponds to a unique combination, \textbf{providing an
objectively correct answer for each task.} \textbf{The use of these
manipulations allows concise, standardised reporting of illusion
parameters and ensures our stimuli are fully reproducible (see Makowski
et al., 2021).}

Task difficulty corresponds to the difficulty of the perceptual decision
(e.g., if the task is to select the longest red line, task difficulty
corresponds to how the lines are objectively different). Illusion
strength corresponds to the degree to which the illusion elements (e.g.,
the black arrow lines in Müller-Lyer) are interfering with the
aforementioned task. Note that the illusion effect can be either
``incongruent'', \textbf{making the task more difficult by biasing
perceptual decisions toward the incorrect response} or ``congruent'',
\textbf{making the task easier by biasing decisions toward the correct
response (e.g., in the Müller-Lyer illusion, if the outwards-facing
arrowheads are placed on the longer line, identifying which line is the
longest becomes easier)}. Participants respond with a key arrow (left
vs.~right; or up vs.~down), and their reaction time (RT) and accuracy
are recorded.

Visual illusion sensitivity will be measured as the average error rate
in the incongruent condition, and separately for the 3 illusion types.
Although the error rate is arguably a crude score, which does not take
into account the effect of varying illusion strength, the interaction
with task difficulty and the possible adjustments in response strategy
(speed-accuracy trade off), it is also the most simple and easy to
reproduce, hence its usage as our primary outcome for the current
\textbf{registered report}. \textbf{As a secondary exploratory outcome,
the Inverse Efficiency Score (IES, Townsend \& Ashby, 2014) will also be
computed. This metric incorporates both speed and accuracy by dividing
the mean reaction time of correct responses by the proportion of correct
responses, separately for each illusion.}

The two sets of 3 illusion blocks will be separated by 2 short
questionnaires acting as a break, namely the IPIP-6 (Sibley et al.,
2011), measuring 6 personality traits with 24 analogue scales items, and
the PID-5 (Krueger et al., 2011), measuring 5 maladaptive personality
traits with 25 Likert scales items. These questionnaires are included as
a way of providing a break between the two cognitively taxing blocks and
maintain paradigmatic consistency with previous studies (Makowski et
al., 2023). \textbf{Additionally, the psychoticism subscale of the PID-5
will be used to examine the correlation between maladaptive traits and
illusion sensitivity, evaluating the existence of the link proposed in
previous studies (Costa et al., 2023).}

\subsection{Data Analysis}\label{data-analysis}

The PCS will contain several manipulation check indices to identify
problematic participants. \textbf{The phenomenological control task
consists of various auditory and visual exercises.} \textbf{At the start
of the task, participants first hear someone say ``hello.'' They are
then asked to choose from several options, including ``Hello,''
``Goodbye,'' ``How are you,'' and ``Thank you''.} \textbf{Any
participant who selects an option other than the correct one will be
considered inattentive} \textbf{In another exercise, participants
receive the following instruction: ``Open your eyes. You will see only
two balls on the screen\ldots just two balls''.} \textbf{However, three
differently coloured balls are actually displayed.} \textbf{If
participants select the option ``no balls were shown'', it indicates
they failed to pay attention to both the auditory instructions and the
visual stimuli.} \textbf{Lastly, in another exercise, participants are
asked to press the spacebar six times.} \textbf{If they press it fewer
than five times within the allotted time, it suggests a lack of
attentiveness to the auditory instructions.} \textbf{Participants will
be excluded if they fail at least one of these checks.} \textbf{Lastly,
reliability of the PCS will be assessed by computing Cronbach's alpha
(Cronbach, 1951).}

\textbf{To assess whether the illusions functioned as expected, stimuli
will be categorized into three groups: Strong Illusion Strength \&
Incongruent, Mild Illusion Strength \& Incongruent, and Congruent.}
\textbf{Two Bayesian t-tests will be conducted to assess differences in
IES between the Congruent and Mild conditions, and between the Mild and
Strong conditions.} \textbf{Significant differences in these comparisons
will provide evidence that the illusions functioned as intended.}

\textbf{The two outcome measures---error rate and IES---will be computed
for each illusion and for each illusion strength group: Strong Illusion
Strength \& Incongruent, Mild Illusion Strength \& Incongruent, and
Congruent.} \textbf{Correlation will be computed between the mild and
strong groups for each illusion and outcomes separately.} \textbf{If
these correlations are high (\emph{r} \textgreater{} .50, Cohen, 2013),
the mild and strong illusion strength groups will be collapsed and the
outcomes will be recomputed across all trials, otherwise they will be
treated as separate in subsequent analyses.}

\textbf{Reliability analyses will then be conducted on all resulting
indices.} \textbf{First, split-half reliability, to assess internal
consistency, will be computed by correlating two equal subsets of
individual scores, with high correlations expected (\emph{r}
\textgreater{} .5).} \textbf{Second, inter-illusion reliability will be
evaluated using Cronbach's alpha across the three illusions.}

Illusion Game outliers will be flagged based on their RT distributions,
following the same procedure as in (Makowski et al., 2023). \textbf{If
the RT is collapsed to the left (i.e., has \textgreater{} 1/3 of
ultra-fast responses - typically \textless{} 200 ms) in the first set,
the entire participant will be discarded (suggesting that they did not
properly do the task), but if only the second set is bad, then only the
second set will be discarded (as the illusion sensitivity can still be
estimated, albeit with less precision).} In addition, the removal of
individual trials will also be performed {[}RT \textless{} 200 ms or
\textgreater{} 3 SD; following Thériault et al. (2024){]}. \textbf{To
mitigate the risk of confounding effects driven by extreme speed or
accuracy strategies, participants whose RTs are significantly slower
than the group average (RT \textgreater{} 4 SD above the mean, based on
Makowski et al. (2023)) will be excluded from the analysis.}

After removing problematic participants and trials, the outcome measures
(PC and VI sensitivity scores) will be computed and the Bayesian
correlation (with medium prior on the coefficient, i.e., r-scale
parameter set to 1/3) will be computed (using the \emph{BayesFactor}
package, Morey \& Rouder, 2024). This shifted beta prior, as recommended
by Morey and Rouder (2018), offers a balanced approach to estimating
effect sizes, without placing undue weight on larger effect sizes or
artificially inflating evidence for the null hypothesis. Following Lush
et al. (2022), we expect to collect evidence against (BF10 \textless=
1/3) a relationship between PCS and VI sensitivity.
\textbf{Additionally, Bayesian correlations will be computed using the
BayesFactor package, employing a medium prior on the coefficient
(r-scale parameter set to 1/3) to assess relationships between
maladaptive trait facets and illusion sensitivity scores.} \textbf{Based
on prior research (Makowski et al., 2023), we expect to find evidence
(BF10 ≥ 3) supporting a relationship between the psychoticism facet of
the PID-5 and illusion sensitivity.} Data analysis will be carried out
using R, using \emph{tidyverse} (Wickham et al., 2019) and
\emph{easystats} (Lüdecke et al., 2020, 2022; Makowski et al., 2019,
2022; Patil et al., 2022). The analysis script and additional
information are available at \textbf{osf link} {[}Note this link will be
replaced with the GitHub page of the current project upon completion of
the review process to ensure continued anonymisation{]}.

\section{Results}\label{results}

\emph{This section will be completed after data is collected.}

\section{Discussion}\label{discussion}

\emph{This section will be completed after data is collected.}

\section{Data Availability}\label{data-availability}

All the study materials, experiment, data, and analysis is available on
GitHub. {[}For the review process these materials can be accessed here:
\textbf{OSF read-only link}. Note this link will be replaced with the
GitHub page of the current project upon completion of the review process
to ensure continued anonymisation{]}.

\section{Acknowledgments}\label{acknowledgments}

We would like to thank An Shu Te for her help in setting up the project,
Ryan Scott for his help in implementing the phenomenological control
scale, and Zoltan Dienes for his input, feedback and guidance.

\pagebreak

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-cohen2013statistical}
Cohen, J. (2013). \emph{Statistical power analysis for the behavioral
sciences}. routledge.

\bibitem[\citeproctext]{ref-costa2023}
Costa, A. L. L., Costa, D. L., Pessoa, V. F., Caixeta, F. V., \& Maior,
R. S. (2023). Systematic review of visual illusions in schizophrenia.
\emph{Schizophrenia Research}, \emph{252}, 13--22.
\url{https://doi.org/10.1016/j.schres.2022.12.030}

\bibitem[\citeproctext]{ref-cronbach1951coefficient}
Cronbach, L. J. (1951). Coefficient alpha and the internal structure of
tests. \emph{Psychometrika}, \emph{16}(3), 297--334.

\bibitem[\citeproctext]{ref-de2015jspsych}
De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating
behavioral experiments in a web browser. \emph{Behavior Research
Methods}, \emph{47}, 1--12.

\bibitem[\citeproctext]{ref-deleeuw2023}
De Leeuw, J. R. (2023). DataPipe: Born-open data collection for online
experiments. \emph{Behavior Research Methods}, \emph{56}(3), 2499--2506.
\url{https://doi.org/10.3758/s13428-023-02161-x}

\bibitem[\citeproctext]{ref-dienes2022}
Dienes, Z., Lush, P., Palfi, B., Roseboom, W., Scott, R., Parris, B.,
Seth, A., \& Lovell, M. (2022). Phenomenological control as cold
control. \emph{Psychology of Consciousness: Theory, Research, and
Practice}, \emph{9}(2), 101--116.
\url{https://doi.org/10.1037/cns0000230}

\bibitem[\citeproctext]{ref-friston2010}
Friston, K. (2010). The free-energy principle: a unified brain theory?
\emph{Nature Reviews Neuroscience}, \emph{11}(2), 127--138.
\url{https://doi.org/10.1038/nrn2787}

\bibitem[\citeproctext]{ref-gershman2012}
Gershman, S. J., Vul, E., \& Tenenbaum, J. B. (2012). Multistability and
Perceptual Inference. \emph{Neural Computation}, \emph{24}(1), 1--24.
\url{https://doi.org/10.1162/neco_a_00226}

\bibitem[\citeproctext]{ref-gori2016}
Gori, S., Molteni, M., \& Facoetti, A. (2016). Visual illusions: An
interesting tool to investigate developmental dyslexia and autism
spectrum disorder. \emph{Frontiers in Human Neuroscience}, \emph{10}.
\url{https://doi.org/10.3389/fnhum.2016.00175}

\bibitem[\citeproctext]{ref-krueger2011}
Krueger, R. F., Eaton, N. R., Derringer, J., Markon, K. E., Watson, D.,
\& Skodol, A. E. (2011). Personality
in{\emph{DSM{\textendash}5:}}Helping Delineate Personality Disorder
Content and Framing the Metastructure. \emph{Journal of Personality
Assessment}, \emph{93}(4), 325--331.
\url{https://doi.org/10.1080/00223891.2011.577478}

\bibitem[\citeproctext]{ref-parameters}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., \& Makowski, D. (2020).
\emph{Extracting, computing and exploring the parameters of statistical
models using {\textbraceleft}r{\textbraceright}.} \emph{5}, 2445.
\url{https://doi.org/10.21105/joss.02445}

\bibitem[\citeproctext]{ref-easystats}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., Wiernik, B. M., Bacher, E.,
Thériault, R., \& Makowski, D. (2022). \emph{Easystats: Framework for
easy statistical modeling, visualization, and reporting}.
\url{https://easystats.github.io/easystats/}

\bibitem[\citeproctext]{ref-lush2018}
Lush, P., Moga, G., McLatchie, N., \& Dienes, Z. (2018). The
Sussex-Waterloo Scale of Hypnotizability (SWASH): measuring capacity for
altering conscious experience. \emph{Neuroscience of Consciousness},
\emph{2018}(1). \url{https://doi.org/10.1093/nc/niy006}

\bibitem[\citeproctext]{ref-lush2021}
Lush, P., Scott, R. B., Seth, A. K., \& Dienes, Z. (2021). The
Phenomenological Control Scale: Measuring the Capacity for Creating
Illusory Nonvolition, Hallucination and Delusion. \emph{Collabra:
Psychology}, \emph{7}(1). \url{https://doi.org/10.1525/collabra.29542}

\bibitem[\citeproctext]{ref-lush2022}
Lush, P., Seth, A., Dienes, Z., \& Scott, R. B. (2022). \emph{Trait
phenomenological control in top-down and bottom-up effects: ASMR,
visually evoked auditory response and the müller-lyer illusion}.
\url{http://dx.doi.org/10.31234/osf.io/hw4y9}

\bibitem[\citeproctext]{ref-bayestestR}
Makowski, D., Ben-Shachar, M. S., \& Lüdecke, D. (2019).
\emph{bayestestR: Describing effects and their uncertainty, existence
and significance within the bayesian framework.} \emph{4}, 1541.
\url{https://doi.org/10.21105/joss.01541}

\bibitem[\citeproctext]{ref-makowski2021}
Makowski, D., Lau, Z. J., Pham, T., Paul Boyce, W., \& Annabel Chen, S.
H. (2021). A Parametric Framework to Generate Visual Illusions Using
Python. \emph{Perception}, \emph{50}(11), 950--965.
\url{https://doi.org/10.1177/03010066211057347}

\bibitem[\citeproctext]{ref-makowski2023}
Makowski, D., Te, A. S., Kirk, S., Liang, N. Z., \& Chen, S. H. A.
(2023). A novel visual illusion paradigm provides evidence for a general
factor of illusion sensitivity and personality correlates.
\emph{Scientific Reports}, \emph{13}(1).
\url{https://doi.org/10.1038/s41598-023-33148-5}

\bibitem[\citeproctext]{ref-correlation}
Makowski, D., Wiernik, B. M., Patil, I., Lüdecke, D., \& Ben-Shachar, M.
S. (2022).
\emph{{\textbraceleft}{\textbraceleft}Correlation{\textbraceright}{\textbraceright}:
Methods for correlation analysis}.
\url{https://CRAN.R-project.org/package=correlation}

\bibitem[\citeproctext]{ref-mitchell2010}
Mitchell, P., Mottron, L., Soulières, I., \& Ropar, D. (2010).
Susceptibility to the Shepard illusion in participants with autism:
reduced top{-}down influences within perception? \emph{Autism Research},
\emph{3}(3), 113--119. \url{https://doi.org/10.1002/aur.130}

\bibitem[\citeproctext]{ref-morey2018baysefactor}
Morey, R. D., \& Rouder, J. N. (2018). \emph{BayseFactor: Computation of
bayes factors for common designs}.

\bibitem[\citeproctext]{ref-BayesFactor}
Morey, R. D., \& Rouder, J. N. (2024). \emph{BayesFactor: Computation of
bayes factors for common designs}.
\url{https://CRAN.R-project.org/package=BayesFactor}

\bibitem[\citeproctext]{ref-notredame2014}
Notredame, C.-E., Pins, D., Deneve, S., \& Jardri, R. (2014). What
visual illusions teach us about schizophrenia. \emph{Frontiers in
Integrative Neuroscience}, \emph{8}.
\url{https://doi.org/10.3389/fnint.2014.00063}

\bibitem[\citeproctext]{ref-nour2015perception}
Nour, M. M., \& Nour, J. M. (2015). Perception, illusions and bayesian
inference. \emph{Psychopathology}, \emph{48}(4), 217--221.

\bibitem[\citeproctext]{ref-palmer2017}
Palmer, C. J., Lawson, R. P., \& Hohwy, J. (2017). Bayesian approaches
to autism: Towards volatility, action, and behavior. \emph{Psychological
Bulletin}, \emph{143}(5), 521--542.
\url{https://doi.org/10.1037/bul0000097}

\bibitem[\citeproctext]{ref-datawizard}
Patil, I., Makowski, D., Ben-Shachar, M. S., Wiernik, B. M., Bacher, E.,
\& Lüdecke, D. (2022).
\emph{{\textbraceleft}Datawizard{\textbraceright}: An
{\textbraceleft}r{\textbraceright} package for easy data preparation and
statistical transformations}. \emph{7}, 4684.
\url{https://doi.org/10.21105/joss.04684}

\bibitem[\citeproctext]{ref-shoshina2014}
Shoshina, I. I., \& Shelepin, Yu. E. (2014). Effectiveness of
Discrimination of the Sizes of Line Segments by Humans with Different
Cognitive Style Parameters. \emph{Neuroscience and Behavioral
Physiology}, \emph{44}(7), 748--753.
\url{https://doi.org/10.1007/s11055-014-9978-2}

\bibitem[\citeproctext]{ref-sibley2011mini}
Sibley, C. G., Luyten, N., Purnomo, M., Mobberley, A., Wootton, L. W.,
Hammond, M. D., Sengupta, N., Perry, R., West-Newman, T., Wilson, M. S.,
et al. (2011). The mini-IPIP6: Validation and extension of a short
measure of the big-six factors of personality in new zealand. \emph{New
Zealand Journal of Psychology}, \emph{40}(3).

\bibitem[\citeproctext]{ref-sundareswara2008}
Sundareswara, R., \& Schrater, P. R. (2008). Perceptual multistability
predicted by search model for Bayesian decisions. \emph{Journal of
Vision}, \emph{8}(5), 12. \url{https://doi.org/10.1167/8.5.12}

\bibitem[\citeproctext]{ref-theriault2024}
Thériault, R., Ben-Shachar, M. S., Patil, I., Lüdecke, D., Wiernik, B.
M., \& Makowski, D. (2024). Check your outliers﻿! An introduction to
identifying statistical outliers in R with easystats. \emph{Behavior
Research Methods}. \url{https://doi.org/10.3758/s13428-024-02356-w}

\bibitem[\citeproctext]{ref-todorovic2020}
Todorović, D. (2020). What Are Visual Illusions? \emph{Perception},
\emph{49}(11), 1128--1199.
\url{https://doi.org/10.1177/0301006620962279}

\bibitem[\citeproctext]{ref-townsend2014methods}
Townsend, J. T., \& Ashby, F. G. (2014). Methods of modeling capacity in
simple processing systems. In \emph{Cognitive theory} (pp. 199--239).
Psychology Press.

\bibitem[\citeproctext]{ref-tidyverse}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
\emph{Welcome to the {\textbraceleft}tidyverse{\textbraceright}}.
\emph{4}, 1686. \url{https://doi.org/10.21105/joss.01686}

\end{CSLReferences}

\end{document}

### 2025-04-28 analysis/1_cleaning.qmd
changes to analysis script

@@
 
-data_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")|> 
+# data_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")|> 
+#   group_by(Illusion_Type) |> 
+#   mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+#          Task_Difficulty = abs(Illusion_Difference),
+#          Condition_Illusion = datawizard::categorize(
+#            Illusion_Strength, split="quantile", n_groups=4,
+#            labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
+#          Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+#   ungroup()
+
+data_ig <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_illusion.csv")|> 
   group_by(Illusion_Type) |> 
@@
 The initial sample consisted of `r report::report_participants(data_ppt, age="Age", gender="Gender", education="Education")`, for a total trial number of `r nrow(data_ppt)`.
-
-```{r}
-
-```
 
@@
 
-data_ppt$pcs_tastescore <- rowMeans(pcs[grepl("taste", names(pcs))])
+pcs$pcs_tastescore <- rowMeans(pcs[grepl("taste", names(pcs))])
 
-data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
+pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
 
-#delete with more participants (changes have been made in script)
-pcs <- pcs |>
-  rename(pcs_amnesia = pcs_rememeber)
-
-# data_ppt$pcs <- rowMeans(pcs[grepl("hands|music|mosquito|tastescore|arm|negative|postsession|amnesia(?!_w)", names(pcs))], na.rm = TRUE)
+pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
 
@@
 
-outliers$Hello <- data_ppt |>
-  dplyr::filter(pcs_hello != "Hello")
-
-outliers$Hello <- outliers$Hello$Participant
+# outliers$Hello <- data_ppt |>
+#   dplyr::filter(pcs_hello != "Hello")
+# 
+# outliers$Hello <- outliers$Hello$Participant
 
@@
 # remove after more pilots
-outliers$Balls <- data_ppt |>
-  dplyr::filter(pcs_noballs == TRUE) # remove after more pilots
-
-# outliers$balls <- data_ppt |>
-#   dplyr::filter(pcs_noballs == "TRUE")
-
-outliers$Balls <- outliers$Balls$Participant
-
+# outliers$Balls <- data_ppt |>
+#   dplyr::filter(pcs_noballs == TRUE) # remove after more pilots
+# 
+# # outliers$balls <- data_ppt |>
+# #   dplyr::filter(pcs_noballs == "TRUE")
+# 
+# outliers$Balls <- outliers$Balls$Participant
 ```
@@
 
-outliers$Press <- data_ppt |>
-  dplyr::filter(pcs_press < 5)
-
-outliers$Press <- outliers$Press$Participant
+# outliers$Press <- data_ppt |>
+#   dplyr::filter(pcs_press < 5)
+# 
+# outliers$Press <- outliers$Press$Participant
 ```
@@
   ggplot(aes(x = x)) +
-  geom_histogram(data=dfsub, aes(x = Age, fill=Gender), bins=28) +
+  geom_histogram(data=data_ppt, aes(x = Age, fill=Gender), bins=28) +
   # geom_line(aes(y = y), color = "orange", linewidth=2) +
-  geom_vline(xintercept = mean(dfsub$Age), color = "red", linewidth=1.5) +
+  geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth=1.5) +
   # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
@@
 p_edu <- data_ppt |>
-  mutate(Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |>
+  mutate(Education = fct_relevel(Education, "Other", "Elementary school", "High school", "Bachelor", "Master", "Doctorate")) |>
   ggplot(aes(x = Education)) +
@@
 
-The final sample includes `r report::report_participants(dfsub, age="Age", sex="Sex", education="Education")`.
+The final sample includes `r report::report_participants(data_ppt, age="Age", sex="Sex", education="Education")`.
 

### 2025-04-28 analysis/2_analysis.qmd
changes to analysis script

@@
 ---
-title: "Illusion"
+title: "Illusion Game & PCS"
 subtitle: "Illusion Task"

### 2025-04-28 analysis/3_Illusion.qmd
changes to analysis script

---
title: "Illusion Game & PCS"
subtitle: "Illusion Task"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: false
---

```{r}
#| echo: false

options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE
)

fig.width <- 6 * sqrt(2) # see::golden_ratio(7)
fig.height <- 6

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 200,
  fig.width = fig.width,
  fig.height = fig.height
)
```

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

logmod <- function(x) sign(x) * log(1 + abs(x))
sqrtmod <- function(x) sign(x) * sqrt(abs(x))
cbrtmod <- function(x) sign(x) * (abs(x)**(1 / 3))

#taken from illusion game international
df <- read.csv("../data/data_IllusionGame.csv") |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Side = sign(Illusion_Difference), 
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
         Task_Difficulty = abs(Illusion_Difference),
         Condition_Illusion = datawizard::categorize(
           Illusion_Strength, split="quantile", n_groups=4,
           labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
         Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup() |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
  ) 
```

## GAMs Fitting

```{r make_GAMs}
#| eval: false

if (!dir.exists("models")) {
  dir.create("models")
}

  for (ill in c(

  "Ebbinghaus",
  "VerticalHorizontal",
  "MullerLyer"
)) {
  # print(ill)
  model <- brms::brm(
    Error ~ t2(Illusion_Difference, Illusion_Strength, bs = "tp") + (1 | Participant),
    data = filter(df, Illusion_Type == ill),
    family = "bernoulli",
    algorithm = "sampling",
    refresh = 400
  )

  name <- paste0("gam_", tolower(ill), "_err")
  assign(name, model) # Rename with string
  save(list = name, file = paste0("models/", name, ".Rdata"))

  model <- brms::brm(
    brms::bf(
      RT ~ t2(Illusion_Difference, Illusion_Strength, bs = "tp") + poly(ISI, 2) + (1 | Participant),
      sigma ~ t2(Illusion_Difference, Illusion_Strength, bs = "tp") + poly(ISI, 2) + (1 | Participant),
      beta ~ t2(Illusion_Difference, Illusion_Strength, bs = "tp") + poly(ISI, 2) + (1 | Participant)
    ),
    data = filter(df, Illusion_Type == ill, Error == 0),
    family = "exgaussian",
    init = 0,
    algorithm = "sampling",
    refresh = 400
  )

  name <- paste0("gam_", tolower(ill), "_rt")
  assign(name, model) # Rename with string
  save(list = name, file = paste0("models/", name, ".Rdata"))
}
```

## Ebbinghaus

### Model Selection

```{r ebbinghaus_testmodels}
test_models <- function(data) {
  # TODO: add random effect
  models_err <- list()
  models_rt <- list()
  for (diff in c(
    "Illusion_Difference",
    "logmod(Illusion_Difference)",
    "sqrtmod(Illusion_Difference)",
    "cbrtmod(Illusion_Difference)"
  )) {
    for (ill in c(
      "abs(Illusion_Strength)",
      "logmod(abs(Illusion_Strength))",
      "sqrtmod(abs(Illusion_Strength))",
      "cbrtmod(abs(Illusion_Strength))"
    )) {
      print(paste(diff, "x", ill))
      err <- glmmTMB::glmmTMB(
        as.formula(
          paste0(
            "Error ~ Illusion_Effect / (",
            diff,
            " * ",
            ill,
            ") + (1|Participant)"
          )
        ),
        data = data,
        family = "binomial"
      )
      models_err[[paste(diff, "*", ill)]] <- err

      rt <- glmmTMB::glmmTMB(
        as.formula(
          paste0(
            "RT ~ Illusion_Effect / (",
            diff,
            " * ",
            ill,
            ") + poly(ISI, 2) + (1|Participant)"
          )
        ),
        data = data
      )
      models_rt[[paste(diff, "*", ill)]] <- rt
    }
  }
  mutate(performance::test_performance(models_err), Outcome = "Error") |>
    rbind(mutate(performance::test_performance(models_rt), Outcome = "RT")) |>
    select(-Model, -log_BF) |>
    datawizard::convert_na_to(select = "BF", replacement = 1) |>
    arrange(Outcome, desc(BF)) |>
    display(footer = "Each model is compared to 'Illusion_Difference'")
}

test_models(filter(df, Illusion_Type == "Ebbinghaus"))
```

### Error Rate

```{r}
data <- filter(df, Illusion_Type == "Ebbinghaus")
```

#### Descriptive

```{r}
plot_desc_errors_obs <- function(data, n_groups=5) {
  data |>
    mutate(
      Illusion_Difference = datawizard::categorize(Illusion_Difference,
        n_groups = n_groups,
        split = "equal_length",
        label = "median"
      ),
      bins = datawizard::categorize(Illusion_Strength, split = "equal_length", n_groups = 20)
    ) |>
    group_by(Illusion_Difference, bins) |>
    summarize(
      Illusion_Strength = mean(Illusion_Strength),
      Error = sum(Error) / n()
    ) |>
    ggplot(aes(x = Illusion_Strength)) +
    geom_bar(aes(y = Error, fill = Illusion_Difference),
      alpha = 1 / 3, position = "dodge", stat = "identity", width = diff(range(data$Illusion_Strength)) / 20
    )
}

plot_desc_errors <- function(data) {

  dat <- data |>
    mutate(
      Illusion_Difference =
        datawizard::categorize(Illusion_Difference,
          n_groups = 5,
          split = "equal_length",
          label = "median"
        )
    )

  col <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(dat$Illusion_Difference)))

  plot_desc_errors_obs(data) +
    geom_smooth(data=dat,
                aes(y = Error, color = Illusion_Difference, fill = Illusion_Difference),
      method = "gam",
      formula = y ~ s(x, bs = "cr", k=8),
      method.args = list(family = "binomial"),
      alpha = 1 / 3
    ) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
    scale_color_manual(values = col) +
    scale_fill_manual(values = col) +
    coord_cartesian(ylim = c(0, 1), xlim = range(dat$Illusion_Strength)) +
    labs(x = "Illusion Strength", y = "Probability of Error") +
    theme_modern() +
    facet_grid(~Illusion_Side, labeller = "label_both")
}

plot_desc_errors(data)
```

#### Model Specification

```{r m_err_ebbinghaus}
#| eval: false

formula <- brms::bf(
  Error ~ Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) +
    (1 + Illusion_Effect / (logmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
  family = "bernoulli",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_ebbinghaus_err <- brms::brm(formula,
  data = data,
  refresh = 100,
  normalize = FALSE
)

save(illusion_ebbinghaus_err, file = "models/illusion_ebbinghaus_err.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_ebbinghaus_err.Rdata")
load("models/gam_ebbinghaus_err.Rdata")
```

```{r plot_err_ebbinghaus}

plot_model_errors <- function(data, model, gam) {
  pred <- estimate_relation(model, length = c(3, 100))

  dat <- data |>
    mutate(
      Illusion_Difference = datawizard::categorize(
        Illusion_Difference,
        n_groups = 3,
        split = "equal_length",
        label = "median"
      ),
      Illusion_Strength = as.character(Illusion_Strength)
    ) |>
    group_by(Illusion_Strength, Illusion_Difference) |>
    summarize(
      Error = sum(Error),
      n = n()
    ) |>
    group_by(Illusion_Strength) |>
    mutate(
      n = sum(n),
      Error = Error / n,
      Illusion_Strength = as.numeric(Illusion_Strength)
    ) |>
    ungroup()

  plot_desc_errors_obs(data, n_groups = 3) +
    scale_fill_manual(values = c("#F44336", "#FFC107", "#4CAF50"), guide = "none") +
    ggnewscale::new_scale_fill() +
    geom_ribbon(data=pred, aes(
      ymin = CI_low, ymax = CI_high,
      fill = Illusion_Difference, group = Illusion_Difference
    ),
    alpha = 0.2
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.5), linetype = "dotted", alpha = 0.5) +
    geom_line(
      data = estimate_relation(gam, length = c(3, 100)),
      aes(y=Predicted, color = Illusion_Difference, group = Illusion_Difference),
      linetype = "dotted"
    )  +
    geom_line(data=pred, aes(y=Predicted, color = Illusion_Difference, group = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = c("#4CAF50", "#FFC107", "#F44336"), trans = "reverse") +
    scale_fill_gradientn(colours = c("#4CAF50", "#FFC107", "#F44336"), trans = "reverse") +
    theme_modern(axis.title.space = 5) +
    guides(color = "none") +
    labs(
      color = "Difficulty",
      fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

p_ebbinghaus_err <- plot_model_errors(data, illusion_ebbinghaus_err, gam_ebbinghaus_err)
p_ebbinghaus_err
```

#### Model Performance

```{r}
performance::performance(illusion_ebbinghaus_err, metrics = c("R2", "ICC")) |>
  display()
```

### Response Time

```{r}
data <- filter(df, Illusion_Type == "Ebbinghaus", Error == 0)
```

#### Descriptive

```{r}
plot_desc_rt <- function(data) {
  dat <- data |>
    mutate(
      Illusion_Difference =
        datawizard::categorize(Illusion_Difference,
          n_groups = 5,
          split = "equal_length",
          label = "median"
        )
    )

  col <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(dat$Illusion_Difference)))

  dat |>
    ggplot(aes(x = Illusion_Strength, y = RT)) +
    geom_jitter2(aes(color = Illusion_Difference),
      width = diff(range(dat$Illusion_Strength)) / 50,
      alpha = 1 / 3
    ) +
    geom_smooth(aes(y = RT, color = Illusion_Difference, fill = Illusion_Difference),
      method = "gam",
      formula = y ~ s(x, bs = "cr", k=8),
      alpha = 1 / 10
    ) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_color_manual(values = col) +
    scale_fill_manual(values = col) +
    coord_cartesian(ylim = c(0, 2)) +
    labs(x = "Illusion Strength", y = "Response Time") +
    theme_modern() +
    ggside::geom_ysidedensity(aes(fill = Illusion_Difference), color = NA, alpha = 0.3) +
    ggside::theme_ggside_void() +
    ggside::scale_ysidex_continuous(expand = c(0, 0)) +
    ggside::ggside() +
    facet_grid(~Illusion_Side, labeller = "label_both")
}

plot_desc_rt(data)
```

#### Model Specification

```{r m_rt_ebbinghaus}
#| eval: false

formula <- brms::bf(
  RT ~ Illusion_Effect / (Illusion_Difference * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 + Illusion_Effect / (Illusion_Difference * abs(Illusion_Strength)) | Participant),
  sigma ~ Illusion_Effect / (Illusion_Difference * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 | Participant),
  beta ~ Illusion_Effect / (Illusion_Difference * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 | Participant),
  family = "exgaussian",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_ebbinghaus_rt <- brms::brm(formula,
  data = data,
  refresh = 100,
  init = 0,
  normalize = FALSE
)

save(illusion_ebbinghaus_rt, file = "models/illusion_ebbinghaus_rt.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_ebbinghaus_rt.Rdata")
load("models/gam_ebbinghaus_rt.Rdata")
```

```{r}
plot_model_rt <- function(data, model, gam) {
 pred <- estimate_relation(
  model,
  at = c("Illusion_Effect", "Illusion_Difference", "Illusion_Strength"),
  length = c(3,100)
)

  dat <- data |>
    mutate(Illusion_Difference = datawizard::categorize(
      Illusion_Difference,
      n_groups = 3,
      split = "equal_length",
      label = "median"
    ))

  col <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(dat$Illusion_Difference)))

  pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
    ggdist::stat_gradientinterval(
      data = dat,
      aes(y = RT, fill = Illusion_Difference),
      geom = "slab",
      scale = 1,
      width = diff(range(dat$Illusion_Strength)) / 20,
      fill_type = "gradient",
      position = "dodge"
    ) +
    scale_fill_manual(values = col, guide = "none") +
    ggnewscale::new_scale_fill() +
    geom_ribbon(aes(
      ymin = CI_low, ymax = CI_high,
      fill = Illusion_Difference, group = Illusion_Difference
    ),
    alpha = 0.2
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
    geom_line(
      data = estimate_relation(gam, at=c("Illusion_Difference", "Illusion_Strength"), length = c(3, 100)),
      aes(color = Illusion_Difference, group = Illusion_Difference),
      linetype = "dotted"
    ) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_gradientn(colours = rev(col), trans = "reverse") +
    scale_fill_gradientn(colours = rev(col), trans = "reverse") +
    coord_cartesian(ylim = c(0.25, 1.25)) +
    labs(
      x = "Illusion Strength",
      y = "Response Time (s)",
      color = "Difficulty",
      fill = "Difficulty"
    ) +
    theme_modern(axis.title.space = 5)
  # ggnewscale::new_scale_fill() +
  # scale_fill_manual(values = c("#F44336", "#FFC107", "#4CAF50"), guide = "none") +
  # ggside::geom_ysidedensity(data=dat,
  #                           aes(fill = Illusion_Difference, y = RT), color = NA, alpha = 0.3) +
  # ggside::theme_ggside_void() +
  # ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  # ggside::ggside()
}

p_ebbinghaus_rt <- plot_model_rt(data, illusion_ebbinghaus_rt, gam_ebbinghaus_rt)
p_ebbinghaus_rt
```

#### Model Performance

```{r}
performance::performance(illusion_ebbinghaus_rt, metrics = c("R2", "ICC")) |>
  display()
```

```{r}
plot_ppcheck <- function(model, gam) {

  pred <- modelbased::estimate_prediction(model, keep_iterations = 50) |>
    bayestestR::reshape_iterations() |>
    mutate(iter_group = as.factor(iter_group)) |>
    estimate_density(select = "iter_value", at = "iter_group")

  predgam <- modelbased::estimate_prediction(gam, keep_iterations = 50) |>
    bayestestR::reshape_iterations() |>
    mutate(iter_group = as.factor(iter_group)) |>
    estimate_density(select = "iter_value", at = "iter_group")

  estimate_density(insight::get_data(model)$RT) |>
    ggplot(aes(x = x, y = y)) +
    geom_area(fill = "#9E9E9E") +
    geom_line(
      data = predgam,
      aes(group = iter_group), color = "#2196F3", size = 0.1, alpha = 0.5
    ) +
    geom_line(
      data = pred,
      aes(group = iter_group), color = "#FF5722", size = 0.1, alpha = 0.5
    ) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    coord_cartesian(xlim = c(0, 2)) +
    theme_modern() +
    labs(x = "Reaction Time (ms)", y = "", title = "Posterior Predictive Check") +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.y = element_blank()
    )
}

plot_ppcheck(illusion_ebbinghaus_rt, gam_ebbinghaus_rt)
```

### Visualization

```{r}
plot_all <- function(data, p_err, p_rt, question="") {
  illname <- unique(data$Illusion_Type)
  # Get stimuli
  dat <- data |>
    filter(Error == 0) |>
    filter(
      Illusion_Type == illname,
      Response %in% c("left", "up")
    ) |>
    select(File, Illusion_Strength, Illusion_Difference)

  dat <- rbind(
    filter(dat, Illusion_Strength == min(Illusion_Strength)) |>
      filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
    filter(dat, Illusion_Strength == max(Illusion_Strength)) |>
      filter(Illusion_Difference %in% c(min(Illusion_Difference), max(Illusion_Difference))),
    filter(dat, Illusion_Difference == min(Illusion_Difference)) |>
      filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength))),
    filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
      filter(Illusion_Strength %in% c(min(Illusion_Strength), max(Illusion_Strength)))
  )

  img_leftdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
    filter(Illusion_Strength == min(Illusion_Strength)) |>
    pull(File) |>
    unique()
  img_rightdown <- filter(dat, Illusion_Difference == max(Illusion_Difference)) |>
    filter(Illusion_Strength == max(Illusion_Strength)) |>
    pull(File) |>
    unique()
  img_leftup <- filter(dat, Illusion_Strength == min(Illusion_Strength)) |>
    filter(Illusion_Difference == min(Illusion_Difference)) |>
    pull(File) |>
    unique()
  img_rightup <- filter(dat, Illusion_Strength == max(Illusion_Strength)) |>
    filter(Illusion_Difference == min(Illusion_Difference)) |>
    pull(File) |>
    unique()

  img_leftdown <- paste0("../experiment/stimuli/", img_leftdown, ".png") |>
    png::readPNG() |>
    grid::rasterGrob(interpolate = TRUE) |>
    patchwork::wrap_elements()
  img_rightdown <- paste0("../experiment/stimuli/", img_rightdown, ".png") |>
    png::readPNG() |>
    grid::rasterGrob(interpolate = TRUE) |>
    patchwork::wrap_elements()
  img_leftup <- paste0("../experiment/stimuli/", img_leftup, ".png") |>
    png::readPNG() |>
    grid::rasterGrob(interpolate = TRUE) |>
    patchwork::wrap_elements()
  img_rightup <- paste0("../experiment/stimuli/", img_rightup, ".png") |>
    png::readPNG() |>
    grid::rasterGrob(interpolate = TRUE) |>
    patchwork::wrap_elements()

  p <- ((p_err + theme(
    axis.title.x = element_blank(),
    plot.title = element_blank()
  )) /
    (p_rt + theme(plot.title = element_blank(), legend.position = "none"))) +
    patchwork::plot_layout(guides = "collect")

  wrap_elements(((img_leftup | patchwork::plot_spacer() | img_rightup) / p / (img_leftdown | patchwork::plot_spacer() | img_rightdown) +
    patchwork::plot_layout(heights = c(0.5, 1.5, 0.5)) +
    patchwork::plot_annotation(
      title = case_when(
        illname == "MullerLyer" ~ "Müller-Lyer",
        illname == "VerticalHorizontal" ~ "Vertical-Horizontal",
        TRUE ~ "Ebbinghaus"
      ),
      subtitle = question,
      theme = theme(plot.title = element_text(
        size=rel(1.75), face="bold", hjust=0.5),
        plot.subtitle = element_text(size=rel(1), face="italic", hjust=0.5, margin=margin(0,0,-30,0))))))
}

p_ebbinghaus <- plot_all(data, p_ebbinghaus_err, p_ebbinghaus_rt, "Which red circle is bigger?")
p_ebbinghaus
```

## Müller-Lyer

### Model Selection

```{r mullerlyer_testmodels}
test_models(filter(df, Illusion_Type == "MullerLyer"))
```

### Error Rate

```{r}
data <- filter(df, Illusion_Type == "MullerLyer")
```

#### Descriptive

```{r}
plot_desc_errors(data)
```

#### Model Specification

```{r m_err_mullerlyer}
#| eval: false

formula <- brms::bf(
  Error ~ Illusion_Effect / (logmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) +
    (1 + Illusion_Effect / (logmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) | Participant),
  family = "bernoulli",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_mullerlyer_err <- brms::brm(formula,
  data = data,
  refresh = 100,
  normalize = FALSE
)

save(illusion_mullerlyer_err, file = "models/illusion_mullerlyer_err.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_mullerlyer_err.Rdata")
load("models/gam_mullerlyer_err.Rdata")
```

```{r}
p_mullerlyer_err <- plot_model_errors(data, illusion_mullerlyer_err, gam_mullerlyer_err)
p_mullerlyer_err
```

#### Model Performance

```{r}
performance::performance(illusion_mullerlyer_err, metrics = c("R2", "ICC")) |>
  display()
```

### Response Time

```{r}
data <- filter(df, Illusion_Type == "MullerLyer", Error == 0)
```

#### Descriptive

```{r}
plot_desc_rt(data)
```

#### Model Specification

```{r m_rt_mullerlyer}
#| eval: false

formula <- brms::bf(
  RT ~ Illusion_Effect / (sqrtmod(Illusion_Difference) * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 + Illusion_Effect / (sqrtmod(Illusion_Difference) * abs(Illusion_Strength)) | Participant),
  sigma ~ Illusion_Effect / (sqrtmod(Illusion_Difference) * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 | Participant),
  beta ~ Illusion_Effect / (sqrtmod(Illusion_Difference) * abs(Illusion_Strength)) + poly(ISI, 2) +
    (1 | Participant),
  family = "exgaussian",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_mullerlyer_rt <- brms::brm(formula,
  data = data,
  refresh = 100,
  init = 0,
  normalize = FALSE
)

save(illusion_mullerlyer_rt, file = "models/illusion_mullerlyer_rt.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_mullerlyer_rt.Rdata")
load("models/gam_mullerlyer_rt.Rdata")
```

```{r}
p_mullerlyer_rt <- plot_model_rt(data, illusion_mullerlyer_rt, gam_mullerlyer_rt)
p_mullerlyer_rt
```

#### Model Performance

```{r}
performance::performance(illusion_mullerlyer_rt, metrics = c("R2", "ICC")) |>
  display()
```

```{r}
plot_ppcheck(illusion_mullerlyer_rt, gam_mullerlyer_rt)
```

### Visualization

```{r}
p_mullerlyer <- plot_all(data, p_mullerlyer_err, p_mullerlyer_rt, "Which red line is longer?")
p_mullerlyer
```

## Vertical-Horizontal

### Model Selection

```{r verticalhorizontal_testmodels}
test_models(filter(df, Illusion_Type == "VerticalHorizontal"))
```

### Error Rate

```{r}
data <- filter(df, Illusion_Type == "VerticalHorizontal")
```

#### Descriptive

```{r}
plot_desc_errors(data)
```

#### Model Specification

```{r m_err_verticalhorizontal}
#| eval: false

formula <- brms::bf(
  Error ~ Illusion_Effect / (sqrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) +
    (1 + Illusion_Effect / (sqrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) | Participant),
  family = "bernoulli",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_verticalhorizontal_err <- brms::brm(formula,
  data = data,
  refresh = 100,
  normalize = FALSE
)

save(illusion_verticalhorizontal_err, file = "models/illusion_verticalhorizontal_err.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_verticalhorizontal_err.Rdata")
load("models/gam_verticalhorizontal_err.Rdata")
```

```{r}
p_verticalhorizontal_err <- plot_model_errors(
  data,
  illusion_verticalhorizontal_err,
  gam_verticalhorizontal_err
)
p_verticalhorizontal_err
```

#### Model Performance

```{r}
performance::performance(illusion_verticalhorizontal_err, metrics = c("R2", "ICC")) |>
  display()
```

### Response Time

```{r}
data <- filter(df, Illusion_Type == "VerticalHorizontal", Error == 0)
```

#### Descriptive

```{r}
plot_desc_rt(data)
```

#### Model Specification

```{r m_rt_verticalhorizontal}
#| eval: false

formula <- brms::bf(
  RT ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) + poly(ISI, 2) +
    (1 + Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) | Participant),
  sigma ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) + poly(ISI, 2) +
    (1 | Participant),
  beta ~ Illusion_Effect / (cbrtmod(Illusion_Difference) * sqrtmod(abs(Illusion_Strength))) + poly(ISI, 2) +
    (1 | Participant),
  family = "exgaussian",
  decomp = "QR"
)

# brms::get_prior(formula, data = data)
# brms::validate_prior(formula)

illusion_verticalhorizontal_rt <- brms::brm(formula,
  data = data,
  refresh = 100,
  init = 0,
  normalize = FALSE
)

save(illusion_verticalhorizontal_rt, file = "models/illusion_verticalhorizontal_rt.Rdata")
```

#### Model Inspection

```{r}
load("models/illusion_verticalhorizontal_rt.Rdata")
load("models/gam_verticalhorizontal_rt.Rdata")
```

```{r}
p_verticalhorizontal_rt <- plot_model_rt(
  data,
  illusion_verticalhorizontal_rt,
  gam_verticalhorizontal_rt
)
p_verticalhorizontal_rt
```

#### Model Performance

```{r}
performance::performance(illusion_verticalhorizontal_rt, metrics = c("R2", "ICC")) |>
  display()
```

```{r}
plot_ppcheck(illusion_verticalhorizontal_rt, gam_verticalhorizontal_rt)
```

### Visualization

```{r}
p_verticalhorizontal <- plot_all(data, p_verticalhorizontal_err, p_verticalhorizontal_rt, "Which red line is longer?")
p_verticalhorizontal
```

```{r mixedmodels}
#| fig-width: 16
#| fig-height: 10

p <- p_ebbinghaus | p_mullerlyer | p_verticalhorizontal
p
ggsave("figures/figure1.png", p, dpi=300, width=16, height=10)
```

## Individual Scores

### Extract Scores

```{r}
#| eval: false

for(m in c("ebbinghaus", "mullerlyer", "verticalhorizontal")) {
  for(i in c("err", "rt")) {
    load(paste0("models/illusion_", m, "_", i, ".Rdata"))
  }
}
```

```{r get_scores}
get_scores <- function(model, illusion="Ebbinghaus") {
  family <- insight::find_response(model)
  scores <- modelbased::estimate_grouplevel(model, dispersion=TRUE, component="all", test=NULL, diagnostic=NULL) |>
    data_filter(str_detect(Level, "Participant")) |>
    data_filter(str_detect(Parameter, "Illusion_EffectIncongruent")) |>
    mutate(
      Level = str_remove_all(Level, "Participant."),
      Level = str_remove_all(Level, "_sigma."),
      Level = str_remove_all(Level, "_beta."),
      Parameter = str_remove_all(Parameter, "Illusion_EffectIncongruent:"),
      Parameter = str_remove_all(Parameter, "cbrtmod"),
      Parameter = str_remove_all(Parameter, "sqrtmod"),
      Parameter = str_remove_all(Parameter, "logmod"),
      Parameter = str_remove_all(Parameter, "abs"),
      Parameter = case_when(str_detect(Parameter, "Illusion_Difference:Illusion_Strength") ~ "Interaction",
                            str_detect(Parameter, "Illusion_Strength") ~ "Strength",
                            str_detect(Parameter, "Illusion_Difference") ~ "Diff",
                            str_detect(Parameter, "Illusion_EffectIncongruent") ~ "Incongruent", # Temporary
                            TRUE ~ Parameter),
      Parameter = case_when(str_detect(Group, "_sigma") ~ paste0(Parameter, "Sigma"),
                            str_detect(Group, "_beta") ~ paste0(Parameter, "Beta"),
                            TRUE ~ Parameter), 
      Model = family,
      Illusion=illusion
    ) |> 
    data_filter(Parameter != "Incongruent")

  scores
}

ebbinghaus_err <- get_scores(illusion_ebbinghaus_err, "Ebbinghaus")
ebbinghaus_rt <- get_scores(illusion_ebbinghaus_rt, "Ebbinghaus")
mullerlyer_err <- get_scores(illusion_mullerlyer_err, "MullerLyer")
mullerlyer_rt <- get_scores(illusion_mullerlyer_rt, "MullerLyer")
verticalhorizontal_err <- get_scores(illusion_verticalhorizontal_err, "VerticalHorizontal")
verticalhorizontal_rt <- get_scores(illusion_verticalhorizontal_rt, "VerticalHorizontal")
```

### Reliability

```{r}
#| warning: false

get_reliability <- function(scores) {
  scores |> 
    summarise(SD_Median = sd(Median, na.rm=TRUE),
              Mean_MAD = mean(MAD, na.rm=TRUE),
              Model = first(Model),
              .by=c("Illusion", "Parameter")) |> 
    mutate(Reliability = SD_Median / Mean_MAD, 
           Parameter = paste0(Model, "_", Parameter)) |> 
    select(-c(SD_Median, Mean_MAD, Model)) 
}

rel <- rbind(
  get_reliability(ebbinghaus_err),
  get_reliability(ebbinghaus_rt),
  get_reliability(mullerlyer_err),
  get_reliability(mullerlyer_rt),
  get_reliability(verticalhorizontal_err),
  get_reliability(verticalhorizontal_rt)
) |> 
  pivot_wider(names_from = "Illusion", values_from = "Reliability")

rel |> 
  gt::gt() |> 
  gt::fmt_number() |> 
  gt::data_color(columns=-1, method="numeric", 
                 palette=c("red", "orange", "yellow", "#CDDC39", "#8BC34A", "#4CAF50"), 
                 domain=c(0, 3), na_color="#2E7D32") |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) |> 
  gt::tab_header("Reliability")
```

### Figure

```{r}
#| fig-width: 17
#| fig-height: 18

get_plot <- function(scores) {
  colors <- c("Diff" = "#2196F3", "Strength" = "#9C27B0", "Interaction" = "#FF5722",
              "IncongruentBeta" = "#f9cb35", "IncongruentSigma" = "#f98e09",  # Temporary
              "DiffBeta" = "#57106e", "InterceptBeta" = "#f98e09",
              "DiffSigma" = "#8a226a", "InterceptSigma" = "#f9cb35")

  scores |>
    ggplot(aes(x = Median, y = Level)) +
    geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color = Parameter), linewidth=0.2, size=0.25) +
    scale_color_manual(values = colors, guide = "none") +
    scale_fill_manual(values = colors, guide = "none") +
    labs(y = "Participants") +
    theme_modern() +
    theme(
      strip.placement = "oustide",
      axis.title.x = element_blank(),
      axis.text.y = element_blank()
    ) +
    ggside::geom_xsidedensity(aes(fill = Parameter, y = after_stat(scaled)), color = NA, alpha = 0.7) +
    ggside::theme_ggside_void() +
    ggside::scale_xsidey_continuous(expand = c(0, 0)) +
    ggside::ggside() +
    facet_grid(~Parameter, switch = "both", scales = "free") +
    ggtitle(paste(scores$Illusion[1], "-", scores$Model[1]))
}

p <- (get_plot(ebbinghaus_err) + get_plot(ebbinghaus_rt)) /
  (get_plot(mullerlyer_err) + get_plot(mullerlyer_rt)) /
  (get_plot(verticalhorizontal_err) + get_plot(verticalhorizontal_rt)) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Inter- and Intra- Variability of Illusion Scores", theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
p
```

### Save Scores

```{r}
tidy_scores <- function(scores) {
  illusion <- scores$Illusion[1]
  family <- scores$Model[1]
  scores |>
    select(Parameter, Participant = Level, Median) |>
    pivot_wider(names_from = "Parameter", values_from = "Median") |>
    data_addprefix(paste0(illusion, "_"), select = -1) |>
    data_addsuffix(paste0("_", family), select = -1)
}

scores <- tidy_scores(ebbinghaus_err) |> 
  merge(tidy_scores(ebbinghaus_rt), by="Participant") |> 
  merge(tidy_scores(mullerlyer_err), by="Participant") |>
  merge(tidy_scores(mullerlyer_rt), by="Participant") |>
  merge(tidy_scores(verticalhorizontal_err), by="Participant") |>
  merge(tidy_scores(verticalhorizontal_rt), by="Participant")

write.csv(scores, "../data/scores_illusio.csv", row.names = FALSE)
```

### 2025-04-28 paper/R1/manuscript.qmd
changes to analysis script

@@
 
-```{r}
-#| warning: false
-#| label: fig-illusion_framework
-#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "100%"
-#| fig-cap: "The parametric framework for visual illusions (Makowski et al., 2021) applied to the Müller-Lyer illusion.
-Below are examples of stimuli showcasing the manipulation of two parameters, task difficulty and illusion strength."
-
-knitr::include_graphics("img/IG_framework.png")
-```
-
 The "Illusion Game" paradigm [@makowski2023] has been recently developed to measure illusion sensitivity to various illusion types through its behavioural impact (on response time and error rate) in a perceptual decision task (where participants have to respond as fast as possible; e.g., "which of the left or right circles is bigger").
-The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task (see @fig-illusion_framework).
+The stimuli for different classical illusions are created using the *Pyllusion* software [@makowski2021], which allows researchers to modulate the strength of the illusion as a continuous dimension, independently from the difficulty of the perceptual task.
 This paradigm, inspired by psychophysics, lends itself to the computational modelling of illusion sensitivity through its **interference effect —an effect that disrupts an individual's ability to accurately discriminate between perceptual stimuli.**
@@
 It includes 3 illusion types, namely Ebbinghaus, Müller-Lyer, and Vertical-Horizontal (see @fig-illusionexample).
+**In the original Illusion Game, 10 visual illusions were presented in two sets, following a practice trial, and separated by two short questionnaires.**
+**Participants completed a total of 1,340 trials, with the experiment lasting approximately 55 minutes.**
+**In the current procedure, only three illusions are used, selected based on the original study's findings that these illusions most strongly contribute to illusion sensitivity.**
 
@@
 
-**In the original Illusion Game, 10 visual illusions were presented in two sets, following a practice trial, and separated by two short questionnaires.**
-**Participants completed a total of 1,340 trials, with the experiment lasting approximately 55 minutes.**
-**In the current procedure, only three illusions are used, selected based on the original study's findings that these illusions most strongly contribute to illusion sensitivity.**
-
 The procedure encompasses 2 sets of 80 trials for each illusion type, **preceded by a practice trial for each illusion**. 
@@
 
-For each illusion type, two continuous dimensions are orthogonally manipulated namely task difficulty and illusion strength, so that each trial corresponds to a unique combination, **providing an objectively correct answer for each task.**
+For each illusion type, two continuous dimensions are orthogonally manipulated namely task difficulty and illusion strength, so that each trial corresponds to a unique combination, **providing an objectively correct answer for each trial.**
 **The use of these manipulations allows concise, standardised reporting of illusion parameters and ensures our stimuli are fully reproducible [see @makowski2021].**
@@
 In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
-**To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean, based on @makowski2023) will be excluded from the analysis.**
+**To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average [RT > 4 SD above the mean, based on @makowski2023] will be excluded from the analysis.**
 
 After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the *BayesFactor* package, @BayesFactor].
-This shifted beta prior, as recommended by @morey2018baysefactor, offers a balanced approach to estimating effect sizes, without placing undue weight on larger effect sizes or artificially inflating evidence for the null hypothesis.
+**This shifted beta prior, as recommended by @morey2018baysefactor, offers a balanced approach to estimating effect sizes, without placing undue weight on larger effect sizes or artificially inflating evidence for the null hypothesis.**
 Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity. 

### 2025-04-28 analysis/preprocessing.R

@@ -5,8 +5,7 @@ path <- "C:/Users/asf25/Box/IllusionPCS/"
 
 # JsPsych experiment ------------------------------------------------------
 
-# files <- list.files(path, pattern = "*.csv")
-files <- "ryqwk2oej3.csv"
+files <- list.files(path, pattern = "*.csv")
 
 # Progress bar
 progbar <- progress_bar$new(total = length(files))
@@ -190,7 +189,7 @@ for (file in files) {
   
   df_ig <- ig[, c("Illusion_Type", "Illusion_Difference", "Illusion_Strength")]
   df_ig$participant <- sub("\\.csv$", "", file)
-  df_ig$File <- gsub("https://realitybending.github.io/IllusionGame/v3/stimuli/", "", ig$stimulus)
+  df_ig$File <- gsub(".png", "", ig$file)
   df_ig$Block <- ig$block
   df_ig$Trial <- ig$trial_number
   df_ig$ISI <- ig$isi

### 2025-04-28 analysis/1_cleaning.qmd
changes to anlaysis scripts

@@
 
-### PCS
-
-```{r}
-
-pcs <-  select(data_ppt, starts_with("pcs_"))
-
-pcs$pcs_tastescore <- rowMeans(pcs[grepl("taste", names(pcs))])
-
-pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
-
-pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
-
-data_ppt$pcs <- rowMeans(pcs[!grepl("pcs_amnesia_w", names(pcs)) & grepl("hands|music|mosquito|tastescore|arm|negative|postsession|amnesia", names(pcs))], na.rm = TRUE)
-
-data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
-```
-
 ### Experiment Duration
@@
 ## Reaction Time (per Block)
+
+**Note**.
+If the RT is collapsed to the left (i.e., has > 1/3 of ultra-fast responses - typically < 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
 
@@
 ## Reaction Time (per Block)
+
+**Note**.
+The removal of individual trials will also be performed [RT < 200 ms or > 3 SD; following Thériault et al.
+(2024)].
+To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean, based on Makowski et al., 2023) will be excluded from the analysis.
 

### 2025-04-28 analysis/2_analysis.qmd
changes to anlaysis scripts

@@
 library(patchwork)
-library(brms)
+library(psych)
 
@@
 
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv") |>
+# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
+
+data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_participants.csv")
+
+# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv")  |>
+  # group_by(Illusion_Type) |> 
+  # mutate(Illusion_Side = sign(Illusion_Difference), 
+  #   Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+  #   Task_Difficulty = abs(Illusion_Difference),
+  #        Condition_Illusion = datawizard::categorize(
+  #          Illusion_Strength, split="quantile", n_groups=4,
+  #          labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
+  #        Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+  # ungroup() |>
+  # mutate(
+  #   Block = as.factor(Block),
+  #   Illusion_Side = as.factor(Illusion_Side),
+  #   Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
+  # ) 
+  #
+
+#taken from illusion game international
+data_ig <- read.csv("../data/data_IllusionGame.csv") |>
+  group_by(Illusion_Type) |> 
+  mutate(Illusion_Side = sign(Illusion_Difference), 
+    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+    Task_Difficulty = abs(Illusion_Difference),
+      Condition_Illusion = case_when(
+        Illusion_Effect == "Congruent" ~ "Congruent",
+        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
+        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
+    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+  ungroup() |>
   mutate(
@@
     Illusion_Side = as.factor(Illusion_Side),
-    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
+    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
+    Response_Correct = as.factor(Response_Correct)
   ) 
@@
 
+*Note:* Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score.
+The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
+The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.
+
 ```{r}
+data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
+data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
+data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
+data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
+
+```
+
+### Reliability
+```{r}
+
+data_ppt |>
+  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
+  psych::alpha()
+
+```
+
+## Illusion
+
+### Error rate
+
+**Note.** Error rate computed by dividing  the number of errors by the total number of trials.
+
+```{r}
+
+Error_rate <- data_ig |>
+  group_by(Participant, Illusion_Type, Condition_Illusion) |>
+  summarise(
+    incorrect_responses = sum(Response_Correct == "false", na.rm = TRUE),
+    total_responses = n(),
+    error_rate = (incorrect_responses / total_responses) * 100
+  )
+
+Error_rate
+
+```
+
+#### Difference in Error rate between each condition and illusion type
+
+::: {.panel-tabset}
+
+##### Ebbinghaus
+
+```{r}
+
+ER_ebbinghaus <- Error_rate |> filter(Illusion_Type == "Ebbinghaus")
+
+ER_CongruentMild_ebbinghaus_BF <- BayesFactor::ttestBF(
+  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Congruent"],
+  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+ER_StrongMild_ebbinghaus_BF <- BayesFactor::ttestBF(
+  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
+  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+##### MullerLyer
+
+```{r}
+
+ER_mullerlyer <- Error_rate |> filter(Illusion_Type == "MullerLyer")
+
+ER_CongruentMild_mullerlye_BF <- BayesFactor::ttestBF(
+  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Congruent"],
+  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+ER_StrongMild_mullerlye_BF <- BayesFactor::ttestBF(
+  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
+  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+##### Vertical-Horizontal
+
+```{r}
+
+ER_verticalhorizontal  <- Error_rate |> filter(Illusion_Type == "VerticalHorizontal")
+
+ER_CongruentMild_vh_BF <- BayesFactor::ttestBF(
+  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Congruent"],
+  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+ER_StrongMild_vh_BF <- BayesFactor::ttestBF(
+  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
+  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+:::
+
+### IES
+
+**Note.** The IES is calculated by dividing the average correct RT by the proportion of correct responses, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent
+
+```{r}
+
+IES <- data_ig |>
+  group_by(Participant, Illusion_Type, Condition_Illusion) |>
+  summarise(
+    mean_RT = mean(RT[Response_Correct == "true"], na.rm = TRUE),  # Only calculate mean RT for correct responses
+    correct_proportion = mean(Response_Correct == "true", na.rm = TRUE),  # Proportion of correct responses
+    IES = mean_RT / correct_proportion  # Compute IES
+  )
+
+IES
+
+```
+
+#### Difference in IES between each condition and illusion type
+
+::: {.panel-tabset}
+
+##### Ebbinghaus
+
+```{r}
+
+IES_ebbinghaus <- IES |> filter(Illusion_Type == "Ebbinghaus")
+
+BF_Congruent_vs_Mild_ebbinghaus <- BayesFactor::ttestBF(
+  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Congruent"],
+  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+BF_Mild_vs_Strong_ebbinghaus <- BayesFactor::ttestBF(
+  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
+  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+##### MullerLyer
+
+```{r}
+
+IES_mullerlyer <- IES |> filter(Illusion_Type == "MullerLyer")
+
+BF_Congruent_vs_Mild_mullerlyer <- BayesFactor::ttestBF(
+  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Congruent"],
+  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+BF_Mild_vs_Strong_mullerlyer <- BayesFactor::ttestBF(
+  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
+  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+##### Vertical-Horizontal
+
+```{r}
+
+IES_verticalhorizontal  <- IES |> filter(Illusion_Type == "VerticalHorizontal")
+
+BF_Congruent_vs_Mild_verticalhorizontal <- BayesFactor::ttestBF(
+  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Congruent"],
+  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
+  paired = TRUE
+)
+
+BF_Mild_vs_Strong_verticalhorizontal <- BayesFactor::ttestBF(
+  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
+  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Strong"],
+  paired = TRUE
+)
+
+```
+
+:::
+
+### Correlations
+
+**Note**.
+If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.
+
+#### Error Rate
+
+```{r}
+
+ER_Incongruent_eb <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
+BayesFactor::correlationBF(
+  y = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Mild"], 
+  x = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Strong"])
+
+ER_Incongruent_ml <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
+BayesFactor::correlationBF(
+  y = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Mild"], 
+  x = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Strong"])
+
+ER_Incongruent_vh <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
+BayesFactor::correlationBF(
+  y = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Mild"], 
+  x = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Strong"])
+
+```
+
+#### IES
+
+```{r}
+
+IES_Incongruent_eb <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
+BayesFactor::correlationBF(
+  y = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Mild"], 
+  x = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Strong"])
+
+IES_Incongruent_ml <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
+BayesFactor::correlationBF(
+  y = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Mild"], 
+  x = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Strong"])
+
+IES_Incongruent_vh <- IES|> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
+BayesFactor::correlationBF(
+  y = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Mild"], 
+  x = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Strong"])
 
@@
 
-## Illusion
+*Note*. Reliability analyses are conducted on all resulting indices.
 
-### Reliability
+#### Split-half
+
+```{r}
+
+```
+
+#### Inter-illusion reliability
+
+```{r}
+
+```

### 2025-04-29 analysis/1_cleaning.qmd
analyis script completed
currently making the analysis script (i.e., delete_after) cleaner and more efficient (i.e., 2_analysis).

@@
 
-data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_participants.csv")
+data_ppt <- read.csv("../data/rawdata_participants.csv")
 
-# data_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")|> 
-#   group_by(Illusion_Type) |> 
-#   mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
-#          Task_Difficulty = abs(Illusion_Difference),
-#          Condition_Illusion = datawizard::categorize(
-#            Illusion_Strength, split="quantile", n_groups=4,
-#            labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
-#          Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
-#   ungroup()
+# path_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
+path_ig <- read.csv("../data/rawdata_illusion.csv")
 
-data_ig <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_illusion.csv")|> 
+data_ig <- path_ig |>
   group_by(Illusion_Type) |> 
@@
 
-out <- check_trials(data_ig)
-out$text
-out$p
-data_ig <- out$data
+# out <- check_trials(data_ig)
+# out$text
+# out$p
+# data_ig <- out$data
 ```

### 2025-04-29 analysis/2_analysis.qmd
analyis script completed
currently making the analysis script (i.e., delete_after) cleaner and more efficient (i.e., 2_analysis).

--- 2025-04-21 analysis/2_analysis.qmd
@@
 ---
-title: "Illusion"
-subtitle: "Illusion Task"
+title: "Illusion Game & PCS"
+subtitle: "Analysis"
 date: "`r Sys.Date()`"
@@
     toc: true
+    warning: false 
+    message: false 
 execute: 
@@
 ```{r}
-#| message: false
-#| warning: false
-
 library(tidyverse)
@@
 library(patchwork)
-library(brms)
+library(psych)
 
-logmod <- function(x) sign(x) * log(1 + abs(x))
-sqrtmod <- function(x) sign(x) * sqrt(abs(x))
-cbrtmod <- function(x) sign(x) * (abs(x)**(1 / 3))
+# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
 
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv") |>
+data_ppt <- read.csv("../data/data_participants.csv")
+
+data_ig <- read.csv("../data/data_IllusionGame.csv") |>
+  group_by(Illusion_Type) |> 
+  mutate(Illusion_Side = sign(Illusion_Difference), 
+    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+    Task_Difficulty = abs(Illusion_Difference),
+      Illusion_Group = case_when(
+        Illusion_Effect == "Congruent" ~ "Congruent",
+        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
+        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
+    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+  ungroup() |>
   mutate(
@@
     Illusion_Side = as.factor(Illusion_Side),
-    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
+    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
+    Response_Correct = as.factor(Response_Correct)
   )
+
 ```
@@
 
+**Note:** Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score.
+The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
+The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.
+
 ```{r}
+data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
+data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
+data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
+data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
 
@@
 
-## Illusion
+```{r}
 
-### Reliability
+pcs_rel <- data_ppt |>
+  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
+  psych::alpha()
+
+```
+
+## Illusion Sensitivity Scores
+
+### Error Rate
+
+```{r}
+errorrate <- data_ig |>
+  group_by(Participant, Illusion_Type, Illusion_Group) |>
+  summarize(ErrorRate = sum(Error) / n()) |>
+  ungroup() |>
+  as.data.frame()
+```
+
+### IES
+
+**Note.** The IES is calculated by dividing the *average correct* RT by the *proportion of correct responses*, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent
+
+```{r}
+
+IES <- data_ig |>
+  group_by(Participant, Illusion_Type, Illusion_Group) |>
+  summarise(IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true"))) |>
+  ungroup() |>
+  as.data.frame()
+```
+
+```{r}
+
+data_vs <- left_join(IES, errorrate, by= c("Participant", "Illusion_Type", "Illusion_Group"))
+```
+
+### Correlations between Strong and Mild
+
+**Note**.
+If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.
+
+```{r}
+
+compute_correlation <- function(data, illusion_type) {
+  illusion_data <- data |> dplyr::filter(Illusion_Group != "Congruent", Illusion_Type == illusion_type) |>
+    pivot_wider(id_cols = Participant,
+      names_from = Illusion_Group,
+      values_from = c(ErrorRate, IES),
+      names_sep = "_")
+
+  ER_Incongruent <- BayesFactor::correlationBF(
+    y = illusion_data$ErrorRate_Mild, 
+    x = illusion_data$ErrorRate_Strong
+  )
+
+  IES_Incongruent <- BayesFactor::correlationBF(
+    y = illusion_data$IES_Mild, 
+    x = illusion_data$IES_Strong
+  )
+
+  return(list(
+    ER_Incongruent = ER_Incongruent,
+    IES_Incongruent = IES_Incongruent
+  ))
+}
+
+```
+
+```{r}
+
+cor_ebbinghaus <- compute_correlation(data_vs, "Ebbinghaus")
+cor_mullerlyer <- compute_correlation(data_vs, "MullerLyer")
+cor_verticalhorizontal <- compute_correlation(data_vs, "VerticalHorizontal")
+```
+
+```{r}
+
+er_ebbinghaus_post<- describe_posterior(cor_ebbinghaus$ER_Incongruent)
+er_mullerlyer_post <- describe_posterior(cor_mullerlyer$ER_Incongruent)
+er_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$ER_Incongruent)
+```
+
+```{r}
+
+ies_ebbinghaus_post <- describe_posterior(cor_ebbinghaus$IES_Incongruent)
+ies_mullerlyer_post <- describe_posterior(cor_mullerlyer$IES_Incongruent)
+ies_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$IES_Incongruent)
+
+```
+
+### Colapsed scores
+
+```{r}
+
+```
+
+### Illusion Manipulation Checks
+
+**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended
+
+```{r}
+compute_ttest <- function(data, illusion_type) {
+  illusion_data <- data |> dplyr::filter(Illusion_Type == illusion_type)
+
+  ERcongruent_vs_mild <- BayesFactor::ttestBF(
+    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
+    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
+    paired = TRUE
+  )
+
+  IEScongruent_vs_mild <- BayesFactor::ttestBF(
+    x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
+    y = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
+    paired = TRUE
+  )
+
+  ERmild_vs_strong <- BayesFactor::ttestBF(
+    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
+    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Strong"],
+    paired = TRUE
+  )
+
+  IESmild_vs_strong <- BayesFactor::ttestBF(
+    x = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
+    y = illusion_data$IES[illusion_data$Illusion_Group == "Strong"],
+    paired = TRUE
+  )
+
+  return(list(
+    ERcongruent_vs_mild = ERcongruent_vs_mild,
+    ERmild_vs_strong = ERmild_vs_strong,
+    IEScongruent_vs_mild = IEScongruent_vs_mild,
+    IESmild_vs_strong = IESmild_vs_strong
+  ))
+}
+
+````
+
+```{r}
+
+ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
+ttest <- compute_ttest(data_vs, "MullerLyer")
+ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
+```
+
+```{r}
+
+er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
+er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
+
+er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
+bf_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
+
+er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
+er_verticalhorizontal_post_ms <- describe_posterior(ttestcor_verticalhorizontal$ERmild_vs_strong)
+```
+
+```{r}
+
+IES_ebbinghaus_post_cm <- describe_posterior(cor_ebbinghaus$IEScongruent_vs_mild)
+IES_ebbinghaus_post_ms <- describe_posterior(cor_ebbinghaus$IESmild_vs_strong)
+
+IES_mullerlyer_post_cm <- describe_posterior(cor_mullerlyer$IEScongruent_vs_mild)
+IES_mullerlyer_post_ms <- describe_posterior(cor_mullerlyer$IESmild_vs_strong)
+
+IES_verticalhorizontal_post_cm <- describe_posterior(cor_verticalhorizontal$IEScongruent_vs_mild)
+IES_verticalhorizontal_post_ms <- describe_posterior(cor_verticalhorizontal$IESmild_vs_strong)
+```

### 2025-04-29 analysis/.vdoc.43f97950-26a5-418d-962b-9fea91f00032.r

@@ -1,0 +1,628 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| message: false
+#| warning: false
+
+library(tidyverse)
+library(ggdist)
+library(ggside)
+library(easystats)
+library(patchwork)
+library(psych)
+
+
+# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
+
+data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/data_participants.csv")
+
+# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv")  |>
+  # group_by(Illusion_Type) |> 
+  # mutate(Illusion_Side = sign(Illusion_Difference), 
+  #   Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+  #   Task_Difficulty = abs(Illusion_Difference),
+  #        Condition_Illusion = datawizard::categorize(
+  #          Illusion_Strength, split="quantile", n_groups=4,
+  #          labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
+  #        Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
+  # ungroup() |>
+  # mutate(
+  #   Block = as.factor(Block),
+  #   Illusion_Side = as.factor(Illusion_Side),
+  #   Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
+  # ) 
+  #  
+    
+#taken from illusion game international
+data_ig <- read.csv("../data/data_IllusionGame.csv") |>
+  group_by(Illusion_Type) |> 
+  mutate(Illusion_Side = sign(Illusion_Difference), 
[…]

### 2025-04-30 analysis/1_cleaning.qmd
updated analysis script

@@
 ```{r}
-#| code-fold: false
-
 ipip6 <- select(data_ppt, starts_with("ipip_"))
@@
 ```{r}
-#| code-fold: false
-
 pid5 <- select(data_ppt, starts_with("pid_"))
@@
 ```{r p_ethnicity}
-ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
-  ggplot(aes(x, y, fill = group)) + 
-  ggwaffle::geom_waffle() + 
-  coord_equal() + 
-  scale_fill_flat_d() + 
-  ggwaffle::theme_waffle() +
-  labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
-  theme(
-    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
-    plot.subtitle = element_text(size = rel(1.2)),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
+# ggwaffle::waffle_iron(data_ppt, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
+#   ggplot(aes(x, y, fill = group)) + 
+#   ggwaffle::geom_waffle() + 
+#   coord_equal() + 
+#   scale_fill_flat_d() + 
+#   ggwaffle::theme_waffle() +
+#   labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
+#   theme(
+#     plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+#     plot.subtitle = element_text(size = rel(1.2)),
+#     axis.title.x = element_blank(),
+#     axis.title.y = element_blank()
+#   )
 ```

### 2025-04-30 analysis/2_analysis.qmd
updated analysis script

@@
 
+# collapsed_groups_data <- data_vs |>
+#  filter(Illusion_Group %in% c("Mild", "Strong")) |>
+#   group_by(Participant, Illusion_Type) |>
+#   summarise(
+#     Illusion_Group = "Incongruent",
+#     ErrorRate = mean(ErrorRate, na.rm = TRUE),
+#     IES = mean(IES, na.rm = TRUE),
+#     .groups = "drop"
+#   )
+
+# data_vs <- bind_rows(data_vs, collapsed_groups_data) |>
+#   arrange(Participant, Illusion_Type, factor(Illusion_Group, levels = c("Mild", "Strong", "Congruent", "Incongruent"))) |>
+#   filter(Illusion_Group == c( "Congruent", "Incongruent"))
+
+```
+
+```{r}
+data_vs_wide <- data_vs|>
+  select(Participant, Illusion_Type, Illusion_Group, ErrorRate, IES) |>
+  unite("group", Illusion_Type, Illusion_Group, sep = "_") |>
+  pivot_wider(names_from = group, values_from = c(ErrorRate, IES)) |>
+  as.data.frame()
 ```
@@
 
-**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended
+**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended.
 
@@
 
+  # Congruent vs Incongruent if correlation high between mild and strong
+
+  # ERcong_vs_incon < BayesFactor::ttestBF(
+  #   x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
+  #   y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Incongruent"],
+  #   paired = TRUE
+  # )
+  # 
+  # IEScong_vs_incon <- BayesFactor::ttestBF(
+  #   x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
+  #   y = illusion_data$IES[illusion_data$Illusion_Group == "Incongruent"],
+  #   paired = TRUE
+  # )
+
   return(list(
@@
     IESmild_vs_strong = IESmild_vs_strong
+    # ERcong_vs_incon = ERcong_vs_incon,
+    # IEScong_vs_incon = IEScong_vs_incon
   ))
@@
 
-````
+```
 
@@
 ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
-ttest <- compute_ttest(data_vs, "MullerLyer")
+ttest_mullerlyer <- compute_ttest(data_vs, "MullerLyer")
 ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
@@
 er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
+# er_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$ERcong_vs_incon)
 
 er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
-bf_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
+er_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
+# er_mullerlyer_post_ci <- describe_posterior(ttest_mullerlyer$ERcong_vs_incon)
 
 er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
-er_verticalhorizontal_post_ms <- describe_posterior(ttestcor_verticalhorizontal$ERmild_vs_strong)
+er_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$ERmild_vs_strong)
+# er_verticalhorizontal_post_ci <- describe_posterior(ttestcor_verticalhorizontal$ERcong_vs_incon)
+
 ```
@@
 
-IES_ebbinghaus_post_cm <- describe_posterior(cor_ebbinghaus$IEScongruent_vs_mild)
-IES_ebbinghaus_post_ms <- describe_posterior(cor_ebbinghaus$IESmild_vs_strong)
+IES_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$IEScongruent_vs_mild)
+IES_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$IESmild_vs_strong)
+# IES_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)
 
-IES_mullerlyer_post_cm <- describe_posterior(cor_mullerlyer$IEScongruent_vs_mild)
-IES_mullerlyer_post_ms <- describe_posterior(cor_mullerlyer$IESmild_vs_strong)
+IES_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$IEScongruent_vs_mild)
+IES_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$IESmild_vs_strong)
+# IES_mullerlyer_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)
 
-IES_verticalhorizontal_post_cm <- describe_posterior(cor_verticalhorizontal$IEScongruent_vs_mild)
-IES_verticalhorizontal_post_ms <- describe_posterior(cor_verticalhorizontal$IESmild_vs_strong)
+IES_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$IEScongruent_vs_mild)
+IES_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$IESmild_vs_strong)
+# IES_verticalhorizontal_post_ci <- describe_posterior(ttest_verticalhorizontal$IEScong_vs_incon)
 ```
+
+### Reliability
+
+**Note**. Reliability analyses are conducted on all resulting indices.
+
+#### Split-half
+
+```{r}
+
+compute_all_split_half <- function(df, score_vars = c("ErrorRate", "IES")) {
+
+  conditions <- c("Congruent", "Strong", "Mild")
+  # conditions <- c("Congruent", "Incongruent")
+  illusions <- c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")
+  pairs <- list(
+    c("Ebbinghaus", "MullerLyer"),
+    c("VerticalHorizontal", "MullerLyer"),
+    c("VerticalHorizontal", "Ebbinghaus")
+  )
+
+  # Loop through score vars and compute
+  results <- list()
+
+  for (score in score_vars) {
+    cat("\n###", score, "\n")
+
+    for (cond in conditions) {
+      cat("\n##", cond, "\n")
+
+      for (pair in pairs) {
+        var1 <- paste(score, pair[1], cond, sep = "_")
+        var2 <- paste(score, pair[2], cond, sep = "_")
+
+        if (all(c(var1, var2) %in% colnames(df))) {
+          cat("Pair:", var1, "vs", var2, "\n")
+          res <- performance::item_split_half(df[, c(var1, var2)])
+          results[[paste(score, cond, paste(pair, collapse = "_"), sep = "_")]] <- res
+        } else {
+          warning(paste("Missing columns:", var1, "or", var2))
+        }
+      }
+    }
+  }
+
+  return()
+}
+
+split_half_results  <- compute_all_split_half(data_vs_wide)
+
+```
+
+#### Inter-illusion Reliability
+
+Error Rate
+
+```{r}
+
+con_rel_er <- data_vs_wide |>
+  select(matches("ErrorRate.*_Congruent")) |>
+  psych::alpha()
+
+mild_rel_er <- data_vs_wide |>
+  select(matches("ErrorRate.*_mild")) |>
+  psych::alpha()
+
+strong_rel_er <- data_vs_wide |>
+  select(matches("ErrorRate.*_strong")) |>
+  psych::alpha()
+
+# inc_rel_er <- data_vs_wide |>
+#   select(matches("ErrorRate.*_Inongruent")) |>
+#   psych::alpha()
+
+```
+
+```{r}
+
+con_rel_IES <- data_vs_wide |>
+  select(matches("IES.*_Congruent")) |>
+  psych::alpha()
+
+mild_rel_IES <- data_vs_wide |>
+  select(matches("IES.*_mild")) |>
+  psych::alpha()
+
+strong_rel_IES <- data_vs_wide |>
+  select(matches("IES.*_strong")) |>
+  psych::alpha()
+
+# inc_rel_IES <- data_vs_wide |>
+#   select(matches("IES.*_Inongruent")) |>
+#   psych::alpha()
+
+```
+
+## PCS and Illusion Sensitivity
+
+### Error Rate
+
+#### Ebbinghaus 
+```{r}
+
+cor_pcs_er_ebb_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_Ebbinghaus_Congruent)
+
+er_pcs_ebb_con_post <- describe_posterior(cor_pcs_er_ebb_con)
+
+cor_pcs_er_ebb_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_Ebbinghaus_Mild)
+
+er_pcs_ebb_mild_post <- describe_posterior(cor_pcs_er_ebb_m)
+
+cor_pcs_er_ebb_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_Ebbinghaus_Strong)
+
+er_pcs_ebb_str_post <- describe_posterior(cor_pcs_er_ebb_s)
+
+# Incongruent
+# cor_pcs_er_ebb_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
+# 
+# er_pcs_ebb_inc_post <- describe_posterior(cor_pcs_er_ebb_inc)
+
+```
+
+#### Muller-Lyer
+```{r}
+
+cor_pcs_er_mul_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_MullerLyer_Congruent)
+
+er_pcs_mul_con_post <- describe_posterior(cor_pcs_er_mul_con)
+
+cor_pcs_er_mul_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_MullerLyer_Mild)
+
+er_pcs_mul_mild_post <- describe_posterior(cor_pcs_er_mul_m)
+
+cor_pcs_er_mul_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_MullerLyer_Strong)
+
+er_pcs_mul_str_post <- describe_posterior(cor_pcs_er_mul_s)
+
+# Incongruent
+# cor_pcs_er_mul_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
+# 
+# er_pcs_mul_inc_post <- describe_posterior(cor_pcs_er_mul_inc)
+
+```
+
+#### VerticalHorizontal
+```{r}
+
+cor_pcs_er_vh_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_VerticalHorizontal_Congruent)
+
+er_pcs_vh_con_post <- describe_posterior(cor_pcs_er_vh_con)
+
+cor_pcs_er_vh_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_VerticalHorizontal_Mild)
+
+er_pcs_vh_mild_post <- describe_posterior(cor_pcs_er_vh_m)
+
+cor_pcs_er_vh_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$ErrorRate_VerticalHorizontal_Strong)
+
+er_pcs_vh_str_post <- describe_posterior(cor_pcs_er_vh_s)
+
+# Incongruent
+# cor_pcs_er_vh_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
+# 
+# er_pcs_vh_inc_post <- describe_posterior(cor_pcs_er_vh_inc)
+```
+
+### IES
+
+#### Ebbinghaus 
+```{r}
+
+cor_pcs_IES_ebb_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$IES_Ebbinghaus_Congruent)
+
+IES_pcs_ebb_con_post <- describe_posterior(cor_pcs_IES_ebb_con)
+
+cor_pcs_IES_ebb_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$IES_Ebbinghaus_Mild)
+
+IES_pcs_ebb_mild_post <- describe_posterior(cor_pcs_IES_ebb_m)
+
+cor_pcs_IES_ebb_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs,
+  x = data_vs_wide$IES_Ebbinghaus_Strong)
+
+IES_pcs_ebb_str_post <- describe_posterior(cor_pcs_IES_ebb_s)
+
+# Incongruent
+# cor_pcs_IES_ebb_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
+# 
+# IES_pcs_ebb_inc_post <- describe_posterior(cor_pcs_IES_ebb_inc)
+
+```
+
+#### Muller-Lyer
+```{r}
+
+cor_pcs_IES_mul_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_MullerLyer_Congruent)
+
+IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_mul_con)
+
+cor_pcs_IES_mul_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_MullerLyer_Mild)
+
+IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_mul_m)
+
+cor_pcs_IES_mul_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_MullerLyer_Strong)
+
+IES_pcs_mul_str_post <- describe_posterior(cor_pcs_IES_mul_s)
+
+# Incongruent
+# cor_pcs_IES_mul_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
+# 
+# IES_pcs_mul_inc_post <- describe_posterior(cor_pcs_IES_mul_inc)
+```
+
+#### Vertical Horizontal
+```{r}
+
+cor_pcs_IES_vh_con <- BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_VerticalHorizontal_Congruent)
+
+IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_vh_con)
+
+cor_pcs_IES_vh_m <- BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_VerticalHorizontal_Mild)
+
+IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_vh_m)
+
+cor_pcs_IES_vh_s <-  BayesFactor::correlationBF(
+  y = data_ppt$pcs, 
+  x = data_vs_wide$IES_VerticalHorizontal_Strong)
+
+IES_pcs_vh_str_post <- describe_posterior(cor_pcs_IES_vh_s)
+
+# Incongruent
+# cor_pcs_IES_vh_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
+# 
+# IES_pcs_vh_inc_post <- describe_posterior(cor_pcs_IES_vh_inc)
+```
+
+## Correlations between Illusion scores and Psychoticism
+
+#### Ebbinghaus 
+```{r}
+
+cor_pcs_psy_ebb_con <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_Ebbinghaus_Congruent)
+
+IES_psy_ebb_con_post <- describe_posterior(cor_pcs_psy_ebb_con)
+
+cor_pcs_psy_ebb_m <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_Ebbinghaus_Mild)
+
+IES_psy_ebb_mild_post <- describe_posterior(cor_pcs_psy_ebb_m)
+
+cor_pcs_psy_ebb_s <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_Ebbinghaus_Strong)
+
+IES_psy_ebb_str_post <- describe_posterior(cor_pcs_psy_ebb_s)
+
+# Incongruent
+# cor_pcs_psy_ebb_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$IES_Ebbinghaus_Incongruent)
+# 
+# IES_psy_ebb_inc_post <- describe_posterior(cor_pcs_psy_ebb_inc)
+```
+
+#### Muller-Lyer 
+```{r}
+
+cor_pcs_psy_mul_con <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_MullerLyer_Congruent)
+
+IES_psy_mul_con_post <- describe_posterior(cor_pcs_psy_mul_con)
+
+cor_pcs_psy_mul_m <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_MullerLyer_Mild)
+
+IES_psy_mul_mild_post <- describe_posterior(cor_pcs_psy_mul_m)
+
+cor_pcs_psy_mul_s <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_MullerLyer_Strong)
+
+IES_psy_mul_str_post <- describe_posterior(cor_pcs_psy_mul_s)
+
+# Incongruent
+# cor_pcs_psy_mul_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$IES_MullerLyer_Incongruent)
+# 
+# IES_psy_mul_inc_post <- describe_posterior(cor_pcs_psy_mul_inc)
+
+```
+
+#### Vertical-Horizontal
+```{r}
+
+cor_pcs_psy_vh_con <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_VerticalHorizontal_Congruent)
+
+IES_psy_vh_con_post <- describe_posterior(cor_pcs_psy_vh_con)
+
+cor_pcs_psy_vh_m <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_VerticalHorizontal_Mild)
+
+IES_psy_vh_mild_post <- describe_posterior(cor_pcs_psy_vh_m)
+
+cor_pcs_psy_vh_s <- BayesFactor::correlationBF(
+  y = data_ppt$PID5_Psychoticism,
+  x = data_vs_wide$IES_VerticalHorizontal_Strong)
+
+IES_psy_vh_str_post <- describe_posterior(cor_pcs_psy_vh_s)
+
+# Incongruent
+# cor_pcs_psy_vh_inc <-  BayesFactor::correlationBF(
+#   y = data_ppt$pcs,
+#   x = data_vs_wide$IES_VerticalHorizontal_Incongruent)
+# 
+# IES_psy_vh_inc_post <- describe_posterior(cor_pcs_psy_vh_inc)
+
+```

### 2025-05-01 analysis/delete_after.qmd
minor changes to manuscript

---
title: "Illusion Game & PCS"
subtitle: "Illusion Task"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)

# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")

data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/data_participants.csv")

# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv")  |>
  # group_by(Illusion_Type) |> 
  # mutate(Illusion_Side = sign(Illusion_Difference), 
  #   Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
  #   Task_Difficulty = abs(Illusion_Difference),
  #        Condition_Illusion = datawizard::categorize(
  #          Illusion_Strength, split="quantile", n_groups=4,
  #          labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
  #        Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  # ungroup() |>
  # mutate(
  #   Block = as.factor(Block),
  #   Illusion_Side = as.factor(Illusion_Side),
  #   Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
  # ) 
  #

#taken from illusion game international
data_ig <- read.csv("../data/data_IllusionGame.csv") |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Side = sign(Illusion_Difference), 
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
    Task_Difficulty = abs(Illusion_Difference),
      Condition_Illusion = case_when(
        Illusion_Effect == "Congruent" ~ "Congruent",
        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup() |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
    Response_Correct = as.factor(Response_Correct)
  ) 
```

## PCS

### Compute subjective scale score

*Note:* Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.

```{r}
data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)

```

### Reliability
```{r}

data_ppt |>
  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
  psych::alpha()

```

## Illusion

### Error rate

**Note.** Error rate computed by dividing  the number of errors by the total number of trials.

```{r}

Error_rate <- data_ig |>
  group_by(Participant, Illusion_Type, Condition_Illusion) |>
  summarise(
    incorrect_responses = sum(Response_Correct == "false", na.rm = TRUE),
    total_responses = n(),
    error_rate = (incorrect_responses / total_responses) * 100
  )

Error_rate

```

#### Difference in Error rate between each condition and illusion type

::: {.panel-tabset}

##### Ebbinghaus

```{r}

ER_ebbinghaus <- Error_rate |> filter(Illusion_Type == "Ebbinghaus")

ER_CongruentMild_ebbinghaus_BF <- BayesFactor::ttestBF(
  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Congruent"],
  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_ebbinghaus_BF <- BayesFactor::ttestBF(
  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

##### MullerLyer

```{r}

ER_mullerlyer <- Error_rate |> filter(Illusion_Type == "MullerLyer")

ER_CongruentMild_mullerlye_BF <- BayesFactor::ttestBF(
  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Congruent"],
  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_mullerlye_BF <- BayesFactor::ttestBF(
  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

##### Vertical-Horizontal

```{r}

ER_verticalhorizontal  <- Error_rate |> filter(Illusion_Type == "VerticalHorizontal")

ER_CongruentMild_vh_BF <- BayesFactor::ttestBF(
  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Congruent"],
  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_vh_BF <- BayesFactor::ttestBF(
  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

:::

### IES

**Note.** The IES is calculated by dividing the average correct RT by the proportion of correct responses, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent

```{r}

IES <- data_ig |>
  group_by(Participant, Illusion_Type, Condition_Illusion) |>
  summarise(
    mean_RT = mean(RT[Response_Correct == "true"], na.rm = TRUE),  # Only calculate mean RT for correct responses
    correct_proportion = mean(Response_Correct == "true", na.rm = TRUE),  # Proportion of correct responses
    IES = mean_RT / correct_proportion  # Compute IES
  )

IES

```

#### Difference in IES between each condition and illusion type

::: {.panel-tabset}

##### Ebbinghaus

```{r}

IES_ebbinghaus <- IES |> filter(Illusion_Type == "Ebbinghaus")

BF_Congruent_vs_Mild_ebbinghaus <- BayesFactor::ttestBF(
  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Congruent"],
  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_ebbinghaus <- BayesFactor::ttestBF(
  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

##### MullerLyer

```{r}

IES_mullerlyer <- IES |> filter(Illusion_Type == "MullerLyer")

BF_Congruent_vs_Mild_mullerlyer <- BayesFactor::ttestBF(
  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Congruent"],
  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_mullerlyer <- BayesFactor::ttestBF(
  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

##### Vertical-Horizontal

```{r}

IES_verticalhorizontal  <- IES |> filter(Illusion_Type == "VerticalHorizontal")

BF_Congruent_vs_Mild_verticalhorizontal <- BayesFactor::ttestBF(
  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Congruent"],
  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_verticalhorizontal <- BayesFactor::ttestBF(
  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Strong"],
  paired = TRUE
)

```

:::

### Correlations

**Note**. If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.

#### Error Rate

::: {.panel-tabset}

##### Ebbinghaus

```{r}

ER_Incongruent_eb <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
ER_Incongruent_eb_BF <- BayesFactor::correlationBF(
  y = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Strong"])

describe_posterior(ER_Incongruent_eb_BF)
ER_Incongruent_eb_BF
```

##### MullerLyer

```{r}

ER_Incongruent_ml <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
ER_Incongruent_ml_BF <- BayesFactor::correlationBF(
  y = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Strong"])

describe_posterior(ER_Incongruent_ml_BF)
ER_Incongruent_ml_BF
```

##### Vertical-Horizontal

```{r}
ER_Incongruent_vh <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
ER_Incongruent_vh_BF <- BayesFactor::correlationBF(
  y = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Strong"])

describe_posterior(ER_Incongruent_vh_BF)
ER_Incongruent_vh_BF
```

:::

#### IES

::: {.panel-tabset}

##### Ebbinghaus

```{r}

IES_Incongruent_eb <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
IES_Incongruent_eb_BF <- BayesFactor::correlationBF(
  y = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Strong"])

describe_posterior(IES_Incongruent_eb_BF)
IES_Incongruent_eb_BF
```

##### MullerLyer

```{r}
IES_Incongruent_ml <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
IES_Incongruent_ml_BF <- BayesFactor::correlationBF(
  y = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Strong"])

describe_posterior(IES_Incongruent_ml_BF)
IES_Incongruent_ml_BF
```

##### Vertical-Horizontal

```{r}
IES_Incongruent_vh <- IES|> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
IES_Incongruent_vh_BF <- BayesFactor::correlationBF(
  y = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Strong"])

describe_posterior(IES_Incongruent_vh_BF)
IES_Incongruent_vh_BF
```
:::

### Reliability

*Note*. Reliability analyses are conducted on all resulting indices.

#### Split-half

##### Error Rate

```{r}

ER_wide <- Error_rate|>
  select(Participant, Illusion_Type, Condition_Illusion, error_rate) |>
  unite("condition", Illusion_Type, Condition_Illusion, sep = "_") |>
  pivot_wider(names_from = condition, values_from = error_rate) |>
  as.data.frame()

## Congruent 
performance::item_split_half(ER_wide[, c("Ebbinghaus_Congruent", "MullerLyer_Congruent")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Congruent", "MullerLyer_Congruent")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Congruent", "Ebbinghaus_Congruent")])

## Strong
performance::item_split_half(ER_wide[, c("Ebbinghaus_Strong", "MullerLyer_Strong")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Strong", "MullerLyer_Strong")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Strong", "Ebbinghaus_Strong")])

## Mild
performance::item_split_half(ER_wide[, c("Ebbinghaus_Mild", "MullerLyer_Mild")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Mild", "MullerLyer_Mild")])
performance::item_split_half(ER_wide[, c("VerticalHorizontal_Mild", "Ebbinghaus_Mild")])
```

##### IES

```{r}
IES_wide <- IES |>
  select(Participant, Illusion_Type, Condition_Illusion, IES) |>
  unite("condition", Illusion_Type, Condition_Illusion, sep = "_") |>
  pivot_wider(names_from = condition, values_from = IES) |>
  as.data.frame()

## Congruent 
performance::item_split_half(IES_wide[, c("Ebbinghaus_Congruent", "MullerLyer_Congruent")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Congruent", "MullerLyer_Congruent")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Congruent", "Ebbinghaus_Congruent")])

## Strong
performance::item_split_half(IES_wide[, c("Ebbinghaus_Strong", "MullerLyer_Strong")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Strong", "MullerLyer_Strong")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Strong", "Ebbinghaus_Strong")])

## Mild
performance::item_split_half(IES_wide[, c("Ebbinghaus_Mild", "MullerLyer_Mild")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Mild", "MullerLyer_Mild")])
performance::item_split_half(IES_wide[, c("VerticalHorizontal_Mild", "Ebbinghaus_Mild")])

```

#### Inter-illusion reliability

##### IES

```{r}

IES_wide |>
  select("Ebbinghaus_Congruent", "MullerLyer_Congruent", "VerticalHorizontal_Congruent") |>
  psych::alpha()

IES_wide |>
  select("Ebbinghaus_Mild", "MullerLyer_Mild", "VerticalHorizontal_Mild") |>
  psych::alpha()

IES_wide |>
  select("Ebbinghaus_Strong", "MullerLyer_Strong", "VerticalHorizontal_Strong") |>
  psych::alpha()

```

##### Error Rate

```{r}

ER_wide |>
  select("Ebbinghaus_Congruent", "MullerLyer_Congruent", "VerticalHorizontal_Congruent") |>
  psych::alpha()

ER_wide |>
  select("Ebbinghaus_Mild", "MullerLyer_Mild", "VerticalHorizontal_Mild") |>
  psych::alpha()

ER_wide |>
  select("Ebbinghaus_Strong", "MullerLyer_Strong", "VerticalHorizontal_Strong") |>
  psych::alpha()

```

## Correlations between PCS and VI error rate and IES

::: {.panel-tabset}

### PCS and Error Rate

#### Ebbinghaus 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$Ebbinghaus_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$Ebbinghaus_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$Ebbinghaus_Strong)
```

#### Muller-Lyer 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$MullerLyer_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$MullerLyer_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$MullerLyer_Strong)

```

#### Vertical- Horizontal
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$VerticalHorizontal_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$VerticalHorizontal_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = ER_wide$VerticalHorizontal_Strong)

```

### PCS and IES

#### Ebbinghaus 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$Ebbinghaus_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$Ebbinghaus_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$Ebbinghaus_Strong)

```

#### Muller-Lyer 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$MullerLyer_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$MullerLyer_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$MullerLyer_Strong)

```

#### Vertical- Horizontal
```{r}

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$VerticalHorizontal_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$VerticalHorizontal_Mild)

BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = IES_wide$VerticalHorizontal_Strong)

```

## Correlations between Illusion scores and Psychoticism

::: {.panel-tabset}

#### Ebbinghaus 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$Ebbinghaus_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$Ebbinghaus_Mild)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$Ebbinghaus_Strong)

```

#### Muller-Lyer 
```{r}

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$MullerLyer_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$MullerLyer_Mild)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$MullerLyer_Strong)

```

#### Vertical-Horizontal
```{r}

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$VerticalHorizontal_Congruent)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$VerticalHorizontal_Mild)

BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = IES_wide$VerticalHorizontal_Strong)

```

### 2025-05-01 paper/R1/manuscript.qmd
minor changes to manuscript

@@
 Participants will be presented with a consent form followed by demographic questions (gender, education level, age, and ethnicity). 
-**Although these variables are not directly analyzed in the current study, they will be used to provide to provide a detailed and thorough description of the sample and maximizing data reusability.** 
+**Although these variables are not directly analyzed in the current study, they will be used to provide a detailed and thorough description of the sample and maximizing data reusability.** 
 **Participants will then be administered the PCS and the Illusion Game task (IG) in a counterbalanced order.**
@@
 
-The PCS will contain several manipulation check indices to identify problematic participants. 
-**The phenomenological control task consists of various auditory and visual exercises.**
-**At the start of the task, participants first hear someone say "hello." They are then asked to choose from several options, including "Hello," "Goodbye," "How are you," and "Thank you".**
-**Any participant who selects an option other than the correct one will be considered inattentive**
-**In another exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
-**However, three differently coloured balls are actually displayed.**
-**If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.**
-**Lastly, in another exercise, participants are asked to press the spacebar six times.**
-**If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.**
-**Participants will be excluded if they fail at least one of these checks.**
-**Lastly, reliability of the PCS will be assessed by computing Cronbach's alpha [@cronbach1951coefficient].**
+**The phenomenological control task will include several manipulation check indices to identify problematic participants.** 
+**The task consists of various auditory and visual exercises; at the outset, participants hear a voice say “hello” and are asked to select the corresponding phrase from multiple options (e.g., “Hello,” “Goodbye,” “How are you,” “Thank you”).**
+**Selecting an incorrect response indicates inattention to auditory stimuli.** 
+**In a subsequent exercise, participants are instructed: “Open your eyes. You will see only two balls on the screen… just two balls”.** 
+**However, three differently coloured balls are displayed.** 
+**If a participant selects the response “no balls were shown,” it suggests they failed to attend to both the auditory instruction and the visual stimuli.** 
+**In another task, participants are instructed to press the spacebar six times.** 
+**Pressing it fewer than five times within the allotted time indicates a failure to follow the auditory instructions.** 
+**Participants who fail any one of these manipulation checks will be excluded from further analysis.** 
+**Finally, reliability of the PCS will be assessed by computing Cronbach's alpha [@cronbach1951coefficient].**
 
@@
 **Two Bayesian t-tests will be conducted to assess differences in IES between the Congruent and Mild conditions, and between the Mild and Strong conditions.**
+**The IES is calculated by dividing the average correct RTs by the proportion of correct responses.**
 **Significant differences in these comparisons will provide evidence that the illusions functioned as intended.**

### 2025-05-20 analysis/1_cleaning.qmd
simplify: remove SD per dim

@@
 ---
-title: "Illusion Game & PCS"
+title: "Illusion Sensitivity and Phenomenological Control"
 subtitle: "Preprocessing"
@@
 ```{r}
-SD_per_dim <- function(x, dims="") {
-  m <- matrix(nrow=nrow(x), ncol=0)
-  for(s in dims) {
-    m <- cbind(m, sapply(as.data.frame(t(x[grepl(s, names(x))])), sd))
-  }
-  m
-}
-```
-
-```{r}
 ipip6 <- select(data_ppt, starts_with("ipip_"))
@@
 data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("agr", names(ipip6))])
-data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))
-
-# data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Ext", names(ipip6))])
-# data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("Cons", names(ipip6))])
-# data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("Neuro", names(ipip6))])
-# data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("Ppen", names(ipip6))])
-# data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("HonHum", names(ipip6))])
-# data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("Agr", names(ipip6))])
-# data_ppt$IPIP6_SD <- rowMeans(SD_per_dim(ipip6, c("Extraversion", "Conscientiousness", "Neuroticism", "Openness", "HonestyHumility", "Agreeableness")))
 ```
@@
 data_ppt$PID5_Psychoticism <- rowMeans(pid5[grepl("Psy", names(pid5))])
-data_ppt$PID5_SD <- rowMeans(SD_per_dim(pid5, c("Disinhibition", "Detachment", "NegativeAffect",
-                                             "Antagonism", "Psychoticism")))
 ```

### 2025-05-20 analysis/2_analysis.qmd
simplify: remove SD per dim

@@
 ---
-title: "Illusion Game & PCS"
+title: "Illusion Sensitivity and Phenomenological Control"
 subtitle: "Analysis"

### 2025-05-20 analysis/1_cleaning.qmd
simplify outlier handling in IG
@AnafNeves

@@
 
-```{r}
-#| echo: true
-#| code-fold: false
+## Illusion Game Participants
 
-outliers <- list()
-outliers_half <- list()
-```
-
-## Reaction Time (per Block)
-
-**Note**.
-If the RT is collapsed to the left (i.e., has > 1/3 of ultra-fast responses - typically < 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).
+We removed participants with an error rate > 45% (suggesting chance-level, i.e., random, responding) for any of the illusion types.
+Outliers detected in the first block will be discarded entirely (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (suggesting that they got "bored" when having to re-do the task), as the illusion sensitivity can still be estimated, albeit with less precision.
+Additionally, participants whose RTs are significantly slower than the group average (possibly indicative of a different strategy) were also excluded from the analysis (RT > 4 SD above the mean).
 
@@
 
-errorrate <- data_ig |>
-  group_by(Participant, Illusion_Type, Block) |>
-  summarize(ErrorRate_per_block = sum(Error) / n()) |>
-  ungroup() |>
-  arrange(desc(ErrorRate_per_block))
+data_outliers <- data_ig |>
+  summarize(p_Error = sum(Error) / n(),
+            z_RT = mean(RT),
+            .by = c("Participant", "Illusion_Type", "Block")) |> 
+  mutate(z_RT = as.numeric(standardize(z_RT)), .by = c("Illusion_Type", "Block")) |>
+  arrange(desc(p_Error))
 
-d_all <- estimate_density(filter(data_ig, RT < 3)$RT)
+outliers <- filter(data_outliers, p_Error > 0.45 | z_RT > 4)
+outliers_full <- unique(filter(outliers, Block == "A")$Participant)
+outliers_half <- unique(filter(outliers, Block == "B")$Participant)
+outliers_half <- outliers_half[!(outliers_half %in% outliers_full)]
 
-plot_distribution <- function(dat) {
+p1 <- data_outliers |> 
+  ggplot(aes(x = p_Error)) +
+  geom_density(aes(color=Illusion_Type)) +
+  geom_vline(xintercept = 0.45, linetype = "dashed", color="red")
 
-  data_error <- errorrate |>
-    filter(Participant %in% unique(dat$Participant)) |>
-    group_by(Participant, Block) |>
-    summarize(y = mean(ErrorRate_per_block), .groups="drop") |>
-    mutate(x = ifelse(Block == "A", 2.1, 2.3),
-           color = case_when(
-              Participant %in% outliers ~ "red",
-              Participant %in% outliers_half ~ "orange",
-              TRUE ~ "blue"
-            ))
+p2 <- data_outliers |> 
+  ggplot(aes(x = z_RT)) +
+  geom_density(aes(color=Illusion_Type)) +
+  geom_vline(xintercept = 4, linetype = "dashed", color="red")
 
-  dat |>
-    filter(RT < 3) |>
-    estimate_density(select = "RT", at = c("Participant", "Block")) |>
-    group_by(Participant) |>
-    normalize(select = "y") |>
-    ungroup() |>
-    mutate(
-      Participant = fct_relevel(Participant, sort(unique(dat$Participant))),
-      color = case_when(
-        Participant %in% outliers ~ "red",
-        Participant %in% outliers_half ~ "orange",
-        TRUE ~ "blue"
-      )
-    ) |>
-    ggplot(aes(x = x, y = y)) +
-    geom_bar(data = data_error, aes(fill = color), stat = "identity", width=0.19) +
-    geom_segment(aes(x = 2, xend = 2.4, y = 0.5, yend = 0.5), color = "black", linetype="dashed", linewidth = 0.5) +
-    geom_area(data = normalize(d_all, select = "y"), alpha = 0.2) +
-    geom_line(aes(color = color, group = interaction(Participant, Block), linetype = Block), linewidth = 0.8) +
-    # geom_vline(xintercept = 0.125, linetype = "dashed", color = "red", size = 0.5) +
-    scale_color_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
-    scale_fill_manual(values = c("red" = "#F44336", "orange" = "#FF9800", "blue" = "blue"), guide = "none") +
-    scale_x_continuous(expand = c(0, 0), breaks=c(0, 0.5, 1, 1.5, 2), labels=c("0", "0.5", "1", "1.5", "2")) +
-    scale_y_continuous(expand = c(0, 0)) +
-    coord_cartesian(xlim = c(0, 2.4)) +
-    theme_modern() +
-    theme(axis.text.y = element_blank(),
-          axis.line.y = element_blank(),
-          strip.text.x = element_text(size = rel(1.5)),
-          legend.position = "none") +
-    facet_wrap(~Participant, nrow=10) +
-    labs(y = "", x = "Reaction Time (s)")
-}
+p1 / p2
 ```
 
-### 1-100
-
-```{r}
-#| fig-width: 15
-#| fig-height: 15
-#| eval: false
-
-plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:100],])
-```
-
-### 101-200
-
-```{r}
-#| fig-width: 15
-#| fig-height: 15
-#| eval: false
-
-plot_distribution(data_ig[data_ig$Participant %in% data_ppt$Participant[1:200],])
-
-```
-
-:::
-
-We discarded `r length(outliers)` participants (entirely) and `r length(outliers_half)` participant's second blocks.
+We discarded `r length(outliers_full)` participants (entirely) and `r length(outliers_half)` participant's second blocks based on error rate.
 
@@
 
-## Error Rate (per Block)
-
-For each block, we computed the error rate and, if more than 50%, we discarded the whole block (as it likely indicates that instructions got mixed up, for instance participants were selecting the smaller instead of the bigger circle).
-
-```{r p_outliers_errorblock}
-#| eval: false
-
-errorrate |>
-  estimate_density(at = c("Illusion_Type", "Block"), method = "KernSmooth") |>
-  ggplot(aes(x = x, y = y)) +
-  geom_line(aes(color = Illusion_Type, linetype = Block)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed") +
-  scale_x_continuous(expand = c(0, 0), labels = scales::percent) +
-  scale_y_continuous(expand = c(0, 0)) +
-  scale_color_manual(values = c("Ebbinghaus" = "#2196F3", "MullerLyer" = "#4CAF50", "VerticalHorizontal" = "#FF5722")) +
-  labs(y = "Distribution", x = "Error Rate") +
-  theme_modern()
-
-remove_badblocks <- function(df) {
-  n <- nrow(df)
-  df <- df |>
-    group_by(Participant, Illusion_Type, Block) |>
-    mutate(ErrorRate_per_block = sum(Error) / n()) |>
-    ungroup() |>
-    filter(ErrorRate_per_block < 0.5) |>
-    select(-ErrorRate_per_block)
-
-  text <- paste0(
-    "We removed ",
-    n - nrow(df),
-    " (",
-    insight::format_value((n - nrow(df)) / n, as_percent = TRUE),
-    ") trials belonging to bad blocks."
-  )
-  list(data = df, text = text)
-}
-
-out <- remove_badblocks(data_ig)
-print(paste("Illusion (session 1):", out$text))
-data_ig <- out$data
-```
-
-## Reaction Time (per Block)
-
-**Note**.
-The removal of individual trials will also be performed [RT < 200 ms or > 3 SD; following Thériault et al.
-(2024)].
-To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean, based on Makowski et al., 2023) will be excluded from the analysis.
-
-```{r outliers_RTtrials}
-#| eval: false
-
-check_trials <- function(df) {
-  data <- df |>
+```{r}
+data_outliers <- df |>
   mutate(Outlier = ifelse(RT >= 10, TRUE, FALSE)) |>
@@
   ungroup()
-
-  p1 <- data |>
-    filter(RT < 10) |>
-    estimate_density(select = "RT", at = "Participant") |>
-    group_by(Participant) |>
-    normalize(select = "y") |>
-    ungroup() |>
-    merge(data |>
-      group_by(Participant) |>
-      mutate(Threshold = median(RT) + 4 * mad(RT)) |>
-      filter(Error == 0) |>
-      summarize(Threshold = mean(Threshold))) |>
-    mutate(Outlier = ifelse(x >= Threshold, TRUE, FALSE)) |>
-    ggplot(aes(x = x, y = y)) +
-    geom_area(data = normalize(estimate_density(filter(data, RT < 10), select = "RT"), select = "y"), alpha = 0.2) +
-    geom_line(aes(color = Participant, linetype = Outlier), alpha = 0.2) +
-    geom_vline(xintercept = c(125), linetype = "dashed", color = "red") +
-    scale_color_material_d("rainbow", guide = "none") +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    guides(linetype = "none") +
-    coord_cartesian(xlim = c(0, 5)) +
-    theme_modern() +
-    theme(axis.text.y = element_blank()) +
-    labs(y = "", x = "Reaction Time (s)")
-
-  p2 <- data |>
-    group_by(Participant) |>
-    summarize(Outlier = sum(Outlier) / n()) |>
-    mutate(Participant = fct_reorder(Participant, Outlier)) |>
-    ggplot(aes(x = Participant, y = Outlier)) +
-    geom_bar(stat = "identity", aes(fill = Participant)) +
-    scale_fill_material_d("rainbow", guide = "none") +
-    scale_x_discrete(expand = c(0, 0)) +
-    scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
-    see::theme_modern() +
-    theme(axis.text.x = element_blank()) +
-    labs(y = "Percentage of outlier trials")
-
-  text <- paste0(
-    "We removed ",
-    sum(data$Outlier),
-    " (",
-    insight::format_value(sum(data$Outlier) / nrow(data), as_percent = TRUE),
-    ") outlier trials (150 ms < RT < 4 MAD above median)."
-  )
-
-  data <- filter(data, Outlier == FALSE)
-  data$Outlier <- NULL
-
-  list(p = p1 / p2, data = data, text = text)
-}
 ```
 
-```{r p_outliers_RTtrials}
-#| results: 'asis'
-#| eval: false
+## Illusion Game Trials
 
-# out <- check_trials(data_ig)
-# out$text
-# out$p
-# data_ig <- out$data
+The removal of individual trials will also be performed (RT < 0.2 s or > 5 s; Thériault et al., 2024) to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies.
+
+```{r}
+data_ig <- filter(data_ig, RT > 0.2 & RT < 5)
 ```
 
-## Manipulation/Attention Checks
+## PCS
 

### 2025-05-20 analysis/1_cleaning.qmd
minor changes to scripts

@@
 ## Score Computation
+
+### PCS scale score
+
+**Note:** Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score.
+The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
+The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.
+
+```{r}
+data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
+data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
+data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
+data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
+
+```
+
+### PCS Reliability
+
+```{r}
+
+pcs_rel <- data_ppt |>
+  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
+  psych::alpha()
+
+```
 

### 2025-05-20 analysis/2_analysis.qmd
minor changes to scripts

@@
   )
-
-```
-
-## PCS
-
-### Compute subjective scale score
-
-**Note:** Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score.
-The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
-The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.
-
-```{r}
-data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
-data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
-data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
-data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
-
-```
-
-### Reliability
-
-```{r}
-
-pcs_rel <- data_ppt |>
-  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
-  psych::alpha()
 

### 2025-05-21 analysis/1_cleaning.qmd
lesgooo @AnafNeves

--- 2025-05-20 analysis/1_cleaning.qmd
@@
 execute: 
-  cache: true
+  cache: false
 ---
@@
 
-# path_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
-path_ig <- read.csv("../data/rawdata_illusion.csv")
-
-data_ig <- path_ig |>
-  group_by(Illusion_Type) |> 
-  mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
+# read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
+data_ig <- read.csv("../data/rawdata_illusion.csv") |>
+  mutate(
+    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
     Task_Difficulty = abs(Illusion_Difference),
-         Condition_Illusion = datawizard::categorize(
-           Illusion_Strength, split="quantile", n_groups=4,
-           labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
-         Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
-  ungroup()
-
+    Condition_Illusion = case_when(
+        Illusion_Effect == "Congruent" ~ "Congruent",
+        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
+        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
+    Condition_Difficulty = datawizard::categorize(
+      Task_Difficulty, 
+      split = "quantile", 
+      n_groups = 2, 
+      labels = c("Hard", "Easy")),
+    .by = "Illusion_Type"
+  )
 ```
@@
 
-## Score Computation
-
-### IPIP6
-
-```{r}
-ipip6 <- select(data_ppt, starts_with("ipip_"))
-ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]
-
-data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("ext", names(ipip6))])
-data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("cons", names(ipip6))])
-data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("neuro", names(ipip6))])
-data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("open", names(ipip6))])
-data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("honhum", names(ipip6))])
-data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("agr", names(ipip6))])
-```
-
-### PID-5
-
-```{r}
-pid5 <- select(data_ppt, starts_with("pid_"))
-
-data_ppt$PID5_Disinhibition <- rowMeans(pid5[grepl("Dis", names(pid5))])
-data_ppt$PID5_Detachment <- rowMeans(pid5[grepl("Det", names(pid5))])
-data_ppt$PID5_NegativeAffect <- rowMeans(pid5[grepl("NegAff", names(pid5))])
-data_ppt$PID5_Antagonism <- rowMeans(pid5[grepl("Ant", names(pid5))])
-data_ppt$PID5_Psychoticism <- rowMeans(pid5[grepl("Psy", names(pid5))])
-```
-
-### Experiment Duration
+## Experiment Duration
 
@@
         axis.text.y = element_blank()) 
+```
+
+## Personality Questionnaires Computation
+
+### IPIP6
+
+```{r}
+ipip6 <- select(data_ppt, starts_with("ipip_"))
+ipip6[grepl("_R", names(ipip6))] <- 8 - ipip6[grepl("_R", names(ipip6))]
+
+data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Ext", names(ipip6))])
+data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("Cons", names(ipip6))])
+data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("Neuro", names(ipip6))])
+data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("Open", names(ipip6))])
+data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("HonHum", names(ipip6))])
+data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("Agr", names(ipip6))])
+```
+
+### PID-5
+
+```{r}
+pid5 <- select(data_ppt, starts_with("pid_"))
+
+data_ppt$PID5_Disinhibition <- rowMeans(pid5[grepl("Dis", names(pid5))])
+data_ppt$PID5_Detachment <- rowMeans(pid5[grepl("Det", names(pid5))])
+data_ppt$PID5_NegativeAffect <- rowMeans(pid5[grepl("NegAff", names(pid5))])
+data_ppt$PID5_Antagonism <- rowMeans(pid5[grepl("Ant", names(pid5))])
+data_ppt$PID5_Psychoticism <- rowMeans(pid5[grepl("Psy", names(pid5))])
 ```
@@
 
-```{r}
-data_outliers <- df |>
-  mutate(Outlier = ifelse(RT >= 10, TRUE, FALSE)) |>
-  group_by(Participant) |>
-  mutate(Outlier = ifelse(RT < 0.150 | standardize(RT, robust = TRUE) > 4, TRUE, Outlier)) |>
-  ungroup()
-```
-
 ## Illusion Game Trials
 
-The removal of individual trials will also be performed (RT < 0.2 s or > 5 s; Thériault et al., 2024) to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies.
+We removed individual extreme trials (RT < 0.2 s or > 5 s; Thériault et al., 2024) to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies.
 
@@
 ```{r}
-
-# outliers$Hello <- data_ppt |>
-#   dplyr::filter(pcs_hello != "Hello")
-# 
-# outliers$Hello <- outliers$Hello$Participant
-
+outliers_pcs1 <- data_ppt[data_ppt$pcs_hello != "Hello", "Participant"]
+outliers_pcs2 <- data_ppt[data_ppt$pcs_noballs == TRUE, "Participant"]
+outliers_pcs3 <- data_ppt[data_ppt$pcs_press < 5, "Participant"]
+outliers_pcs <- unique(c(outliers_pcs1, outliers_pcs2, outliers_pcs3))
 ```
 
-We removed `r length(outliers$Hello)` (`r insight::format_percent(length(outliers$Hello) / nrow(data_ppt))`) participants that did not reported hearing "Hello" at the beggining of the experiment.
+We removed `r length(outliers_pcs)` (`r insight::format_percent(length(outliers_pcs) / nrow(data_ppt))`) participants that either did not reported hearing "Hello" at the beginning of the experiment, responded that no balls were presented in the screen, or pressed less than 5 times when asked to do so.
 
 ```{r}
-# remove after more pilots
-# outliers$Balls <- data_ppt |>
-#   dplyr::filter(pcs_noballs == TRUE) # remove after more pilots
-# 
-# # outliers$balls <- data_ppt |>
-# #   dplyr::filter(pcs_noballs == "TRUE")
-# 
-# outliers$Balls <- outliers$Balls$Participant
-```
-
-We removed `r length(outliers$Balls)` (`r insight::format_percent(length(outliers$Balls) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.
-
-```{r}
-
-# outliers$Press <- data_ppt |>
-#   dplyr::filter(pcs_press < 5)
-# 
-# outliers$Press <- outliers$Press$Participant
-```
-
-We removed `r length(outliers$Press)` (`r insight::format_percent(length(outliers$Press) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.
-
-```{r}
-data_ppt <- filter(data_ppt, !Participant %in% outliers$Hello)
-data_ppt <- filter(data_ppt, !Participant %in% outliers$Balls)
-data_ppt <- filter(data_ppt, !Participant %in% outliers$Press)
-
-data_ig <- filter(data_ig, !Participant %in% outliers$Hello)
-data_ig <- filter(data_ig, !Participant %in% outliers$Balls)
-data_ig <- filter(data_ig, !Participant %in% outliers$Press)
+# data_ppt <- filter(data_ppt, !Participant %in% outliers_pcs)
+# data_ig <- filter(data_ig, !Participant %in% outliers_pcs)
 ```
@@
 
-```{r}
-# data_ppt<- filter(data_ppt, !Participant %in% outliers)
-# data_ig <- filter(data_ig, Participant %in% data_ppt$Participant)
-```
+The final sample includes `r report::report_participants(data_ppt, age="Age", sex="Sex", education="Education")`.
 
-### Age
+## Age
 
 ```{r}
-p_age <- estimate_density(data_ppt$Age) |>
-  normalize(select = y) |>
-  mutate(y = y * 86) |>  # To match the binwidth
-  ggplot(aes(x = x)) +
-  geom_histogram(data=data_ppt, aes(x = Age, fill=Gender), bins=28) +
-  # geom_line(aes(y = y), color = "orange", linewidth=2) +
+p_age <- data_ppt |>  # To match the binwidth
+  ggplot(aes(x = data_ppt)) +
+  geom_histogram(aes(x = Age, fill=Gender), bins=28) +
   geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth=1.5) +
-  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
   scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
@@
 
-### Education
+## Education
 
@@
 
-### Ethnicity
+## Ethnicity
 
@@
 
-### Summary
-
-The final sample includes `r report::report_participants(data_ppt, age="Age", sex="Sex", education="Education")`.
-
-## Save
+# Save
 
@@
 write.csv(data_ppt, "../data/data_participants.csv", row.names = FALSE)
-write.csv(data_ig, "../data/data_IllusionGame.csv", row.names = FALSE)
+write.csv(data_ig, "../data/data_illusion.csv", row.names = FALSE)
 ```

### 2025-05-21 analysis/2_analysis.qmd
lesgooo @AnafNeves

@@
 
-data_ig <- read.csv("../data/data_IllusionGame.csv") |>
-  group_by(Illusion_Type) |> 
-  mutate(Illusion_Side = sign(Illusion_Difference), 
-    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
-    Task_Difficulty = abs(Illusion_Difference),
-      Illusion_Group = case_when(
-        Illusion_Effect == "Congruent" ~ "Congruent",
-        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
-        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
-    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
-  ungroup() |>
+data_ig <- read.csv("../data/data_illusion.csv") |>
   mutate(
     Block = as.factor(Block),
-    Illusion_Side = as.factor(Illusion_Side),
     Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
@@
 
-## Illusion Sensitivity Scores
+## Score Computation
 
-### Error Rate
+### PCS
+
+Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score.
+The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
+The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.
 
 ```{r}
-errorrate <- data_ig |>
-  group_by(Participant, Illusion_Type, Illusion_Group) |>
-  summarize(ErrorRate = sum(Error) / n()) |>
-  ungroup() |>
-  as.data.frame()
+pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")
+
+data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
+data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
 ```
 
-### IES
-
-**Note.** The IES is calculated by dividing the *average correct* RT by the *proportion of correct responses*, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent
+### Illusion Game
 
 ```{r}
-
-IES <- data_ig |>
-  group_by(Participant, Illusion_Type, Illusion_Group) |>
-  summarise(IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true"))) |>
-  ungroup() |>
-  as.data.frame()
+ig_scores <- data_ig |>
+  summarize(ErrorRate = sum(Error) / n(),
+            IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
+            .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")) |> 
+  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"), 
+              values_from = c("ErrorRate", "IES")) 
 ```
@@
 ```{r}
-
-data_vs <- left_join(IES, errorrate, by= c("Participant", "Illusion_Type", "Illusion_Group"))
-```
-
-### Correlations between Strong and Mild
-
-**Note**.
-If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.
-
-```{r}
-
-compute_correlation <- function(data, illusion_type) {
-  illusion_data <- data |> dplyr::filter(Illusion_Group != "Congruent", Illusion_Type == illusion_type) |>
-    pivot_wider(id_cols = Participant,
-      names_from = Illusion_Group,
-      values_from = c(ErrorRate, IES),
-      names_sep = "_")
-
-  ER_Incongruent <- BayesFactor::correlationBF(
-    y = illusion_data$ErrorRate_Mild, 
-    x = illusion_data$ErrorRate_Strong
-  )
-
-  IES_Incongruent <- BayesFactor::correlationBF(
-    y = illusion_data$IES_Mild, 
-    x = illusion_data$IES_Strong
-  )
-
-  return(list(
-    ER_Incongruent = ER_Incongruent,
-    IES_Incongruent = IES_Incongruent
-  ))
-}
-
-```
-
-```{r}
-
-cor_ebbinghaus <- compute_correlation(data_vs, "Ebbinghaus")
-cor_mullerlyer <- compute_correlation(data_vs, "MullerLyer")
-cor_verticalhorizontal <- compute_correlation(data_vs, "VerticalHorizontal")
-```
-
-```{r}
-
-er_ebbinghaus_post<- describe_posterior(cor_ebbinghaus$ER_Incongruent)
-er_mullerlyer_post <- describe_posterior(cor_mullerlyer$ER_Incongruent)
-er_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$ER_Incongruent)
-```
-
-```{r}
-
-ies_ebbinghaus_post <- describe_posterior(cor_ebbinghaus$IES_Incongruent)
-ies_mullerlyer_post <- describe_posterior(cor_mullerlyer$IES_Incongruent)
-ies_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$IES_Incongruent)
-
-```
-
-### Colapsed scores
-
-```{r}
-
-# collapsed_groups_data <- data_vs |>
-#  filter(Illusion_Group %in% c("Mild", "Strong")) |>
-#   group_by(Participant, Illusion_Type) |>
-#   summarise(
-#     Illusion_Group = "Incongruent",
-#     ErrorRate = mean(ErrorRate, na.rm = TRUE),
-#     IES = mean(IES, na.rm = TRUE),
-#     .groups = "drop"
-#   )
-
-# data_vs <- bind_rows(data_vs, collapsed_groups_data) |>
-#   arrange(Participant, Illusion_Type, factor(Illusion_Group, levels = c("Mild", "Strong", "Congruent", "Incongruent"))) |>
-#   filter(Illusion_Group == c( "Congruent", "Incongruent"))
-
-```
-
-```{r}
-data_vs_wide <- data_vs|>
-  select(Participant, Illusion_Type, Illusion_Group, ErrorRate, IES) |>
-  unite("group", Illusion_Type, Illusion_Group, sep = "_") |>
-  pivot_wider(names_from = group, values_from = c(ErrorRate, IES)) |>
-  as.data.frame()
-```
-
-### Illusion Manipulation Checks
-
-**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended.
-
-```{r}
-compute_ttest <- function(data, illusion_type) {
-  illusion_data <- data |> dplyr::filter(Illusion_Type == illusion_type)
-
-  ERcongruent_vs_mild <- BayesFactor::ttestBF(
-    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
-    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
-    paired = TRUE
-  )
-
-  IEScongruent_vs_mild <- BayesFactor::ttestBF(
-    x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
-    y = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
-    paired = TRUE
-  )
-
-  ERmild_vs_strong <- BayesFactor::ttestBF(
-    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
-    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Strong"],
-    paired = TRUE
-  )
-
-  IESmild_vs_strong <- BayesFactor::ttestBF(
-    x = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
-    y = illusion_data$IES[illusion_data$Illusion_Group == "Strong"],
-    paired = TRUE
-  )
-
-  # Congruent vs Incongruent if correlation high between mild and strong
-
-  # ERcong_vs_incon < BayesFactor::ttestBF(
-  #   x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
-  #   y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Incongruent"],
-  #   paired = TRUE
-  # )
-  # 
-  # IEScong_vs_incon <- BayesFactor::ttestBF(
-  #   x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
-  #   y = illusion_data$IES[illusion_data$Illusion_Group == "Incongruent"],
-  #   paired = TRUE
-  # )
-
-  return(list(
-    ERcongruent_vs_mild = ERcongruent_vs_mild,
-    ERmild_vs_strong = ERmild_vs_strong,
-    IEScongruent_vs_mild = IEScongruent_vs_mild,
-    IESmild_vs_strong = IESmild_vs_strong
-    # ERcong_vs_incon = ERcong_vs_incon,
-    # IEScong_vs_incon = IEScong_vs_incon
-  ))
-}
-
-```
-
-```{r}
-
-ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
-ttest_mullerlyer <- compute_ttest(data_vs, "MullerLyer")
-ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
-```
-
-```{r}
-
-er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
-er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
-# er_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$ERcong_vs_incon)
-
-er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
-er_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
-# er_mullerlyer_post_ci <- describe_posterior(ttest_mullerlyer$ERcong_vs_incon)
-
-er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
-er_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$ERmild_vs_strong)
-# er_verticalhorizontal_post_ci <- describe_posterior(ttestcor_verticalhorizontal$ERcong_vs_incon)
-
-```
-
-```{r}
-
-IES_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$IEScongruent_vs_mild)
-IES_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$IESmild_vs_strong)
-# IES_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)
-
-IES_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$IEScongruent_vs_mild)
-IES_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$IESmild_vs_strong)
-# IES_mullerlyer_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)
-
-IES_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$IEScongruent_vs_mild)
-IES_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$IESmild_vs_strong)
-# IES_verticalhorizontal_post_ci <- describe_posterior(ttest_verticalhorizontal$IEScong_vs_incon)
-```
-
-### Reliability
-
-**Note**. Reliability analyses are conducted on all resulting indices.
-
-#### Split-half
-
-```{r}
-
-compute_all_split_half <- function(df, score_vars = c("ErrorRate", "IES")) {
-
-  conditions <- c("Congruent", "Strong", "Mild")
-  # conditions <- c("Congruent", "Incongruent")
-  illusions <- c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")
-  pairs <- list(
-    c("Ebbinghaus", "MullerLyer"),
-    c("VerticalHorizontal", "MullerLyer"),
-    c("VerticalHorizontal", "Ebbinghaus")
-  )
-
-  # Loop through score vars and compute
-  results <- list()
-
-  for (score in score_vars) {
-    cat("\n###", score, "\n")
-
-    for (cond in conditions) {
-      cat("\n##", cond, "\n")
-
-      for (pair in pairs) {
-        var1 <- paste(score, pair[1], cond, sep = "_")
-        var2 <- paste(score, pair[2], cond, sep = "_")
-
-        if (all(c(var1, var2) %in% colnames(df))) {
-          cat("Pair:", var1, "vs", var2, "\n")
-          res <- performance::item_split_half(df[, c(var1, var2)])
-          results[[paste(score, cond, paste(pair, collapse = "_"), sep = "_")]] <- res
-        } else {
-          warning(paste("Missing columns:", var1, "or", var2))
-        }
+corrs <- data.frame()
+for(outcome in c("ErrorRate", "IES")) {
+  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
+    for(cond in c("Congruent", "Mild", "Strong")) {
+      corrs <- correlation::correlation(
+        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Easy")),
+        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Hard"))) |> 
+        mutate(Outcome = outcome) |> 
+        rbind(corrs)
     }
@@
 
-  return()
+# IF correlations > 0.5:
+ig_scores <- data_ig |>
+  summarize(ErrorRate = sum(Error) / n(),
+            IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
+            .by = c("Participant", "Illusion_Type", "Condition_Illusion")) |> 
+  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion"), 
+              values_from = c("ErrorRate", "IES")) 
+```
+
+#### Manipulation Check
+
+Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended.
+
+```{r}
+ttests <- data.frame()
+for(outcome in c("ErrorRate", "IES")) {
+  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
+    for(cond in c("Congruent", "Mild")) {
+      # Add loop for Hard and Easy if need be
+      x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond)]]
+      if(cond == "Congruent") {
+        x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild")]]
+      } else {
+        x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong")]]
+      }
+      ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |> 
+        parameters::parameters() |> 
+        mutate(
+          Outcome = outcome,
+          Illusion_Type = ill_type,
+          Condition_Illusion = cond
+        ) |> 
+        rbind(ttests)
+    }
+  }
 }
 
-split_half_results  <- compute_all_split_half(data_vs_wide)
-
+ttests |> 
+  select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |> 
+  arrange(Illusion_Type, Condition_Illusion, Outcome)
 ```
 
-#### Inter-illusion Reliability
+## Reliability
 
-Error Rate
+### PCS
 
 ```{r}
-
-con_rel_er <- data_vs_wide |>
-  select(matches("ErrorRate.*_Congruent")) |>
+data_ppt[, pcs_items] |>
   psych::alpha()
 
-mild_rel_er <- data_vs_wide |>
-  select(matches("ErrorRate.*_mild")) |>
-  psych::alpha()
-
-strong_rel_er <- data_vs_wide |>
-  select(matches("ErrorRate.*_strong")) |>
-  psych::alpha()
-
-# inc_rel_er <- data_vs_wide |>
-#   select(matches("ErrorRate.*_Inongruent")) |>
-#   psych::alpha()
-
+data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
 ```
 
+### Illusion Game
+
+::: {.panel-tabset}
+
+#### Error Rate - Mild
+
 ```{r}
-
-con_rel_IES <- data_vs_wide |>
-  select(matches("IES.*_Congruent")) |>
+ig_scores |>
+  select(matches("ErrorRate_.*_Mild.*")) |>
   psych::alpha()
-
-mild_rel_IES <- data_vs_wide |>
-  select(matches("IES.*_mild")) |>
-  psych::alpha()
-
-strong_rel_IES <- data_vs_wide |>
-  select(matches("IES.*_strong")) |>
-  psych::alpha()
-
-# inc_rel_IES <- data_vs_wide |>
-#   select(matches("IES.*_Inongruent")) |>
-#   psych::alpha()
-
 ```
 
-## PCS and Illusion Sensitivity
+#### Error Rate - Strong
 
-### Error Rate
-
-#### Ebbinghaus 
 ```{r}
-
-cor_pcs_er_ebb_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_Ebbinghaus_Congruent)
-
-er_pcs_ebb_con_post <- describe_posterior(cor_pcs_er_ebb_con)
-
-cor_pcs_er_ebb_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_Ebbinghaus_Mild)
-
-er_pcs_ebb_mild_post <- describe_posterior(cor_pcs_er_ebb_m)
-
-cor_pcs_er_ebb_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_Ebbinghaus_Strong)
-
-er_pcs_ebb_str_post <- describe_posterior(cor_pcs_er_ebb_s)
-
-# Incongruent
-# cor_pcs_er_ebb_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
-# 
-# er_pcs_ebb_inc_post <- describe_posterior(cor_pcs_er_ebb_inc)
-
+ig_scores |>
+  select(matches("ErrorRate_.*_Strong.*")) |>
+  psych::alpha()
 ```
 
-#### Muller-Lyer
+#### IES - Mild
+
 ```{r}
-
-cor_pcs_er_mul_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_MullerLyer_Congruent)
-
-er_pcs_mul_con_post <- describe_posterior(cor_pcs_er_mul_con)
-
-cor_pcs_er_mul_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_MullerLyer_Mild)
-
-er_pcs_mul_mild_post <- describe_posterior(cor_pcs_er_mul_m)
-
-cor_pcs_er_mul_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_MullerLyer_Strong)
-
-er_pcs_mul_str_post <- describe_posterior(cor_pcs_er_mul_s)
-
-# Incongruent
-# cor_pcs_er_mul_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
-# 
-# er_pcs_mul_inc_post <- describe_posterior(cor_pcs_er_mul_inc)
-
+ig_scores |>
+  select(matches("IES_.*_Mild.*")) |>
+  psych::alpha()
 ```
 
-#### VerticalHorizontal
+#### IES - Strong
+
 ```{r}
-
-cor_pcs_er_vh_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_VerticalHorizontal_Congruent)
-
-er_pcs_vh_con_post <- describe_posterior(cor_pcs_er_vh_con)
-
-cor_pcs_er_vh_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_VerticalHorizontal_Mild)
-
-er_pcs_vh_mild_post <- describe_posterior(cor_pcs_er_vh_m)
-
-cor_pcs_er_vh_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$ErrorRate_VerticalHorizontal_Strong)
-
-er_pcs_vh_str_post <- describe_posterior(cor_pcs_er_vh_s)
-
-# Incongruent
-# cor_pcs_er_vh_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
-# 
-# er_pcs_vh_inc_post <- describe_posterior(cor_pcs_er_vh_inc)
+ig_scores |>
+  select(matches("IES_.*_Strong.*")) |>
+  psych::alpha()
 ```
 
-### IES
+:::
 
-#### Ebbinghaus 
+## Illusion Sensitivity vs. PCS
+
 ```{r}
+features <- merge(ig_scores, data_ppt[c("Participant", "PCS_Score", "PID5_Psychoticism")])
 
-cor_pcs_IES_ebb_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$IES_Ebbinghaus_Congruent)
+r <- correlation::correlation(
+  select(features, -PCS_Score, -PID5_Psychoticism),
+  select(features, PCS_Score),
+  bayesian = TRUE
+)
 
-IES_pcs_ebb_con_post <- describe_posterior(cor_pcs_IES_ebb_con)
-
-cor_pcs_IES_ebb_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$IES_Ebbinghaus_Mild)
-
-IES_pcs_ebb_mild_post <- describe_posterior(cor_pcs_IES_ebb_m)
-
-cor_pcs_IES_ebb_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs,
-  x = data_vs_wide$IES_Ebbinghaus_Strong)
-
-IES_pcs_ebb_str_post <- describe_posterior(cor_pcs_IES_ebb_s)
-
-# Incongruent
-# cor_pcs_IES_ebb_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
-# 
-# IES_pcs_ebb_inc_post <- describe_posterior(cor_pcs_IES_ebb_inc)
-
+r |> 
+  format_table() |> 
+  select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |> 
+  gt::gt() 
 ```
 
-#### Muller-Lyer
+## Illusion Sensitivity vs. Psychoticism
+
 ```{r}
+r <- correlation::correlation(
+  select(features, -PCS_Score, -PID5_Psychoticism),
+  select(features, PID5_Psychoticism),
+  bayesian = TRUE
+)
 
-cor_pcs_IES_mul_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_MullerLyer_Congruent)
-
-IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_mul_con)
-
-cor_pcs_IES_mul_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_MullerLyer_Mild)
-
-IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_mul_m)
-
-cor_pcs_IES_mul_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_MullerLyer_Strong)
-
-IES_pcs_mul_str_post <- describe_posterior(cor_pcs_IES_mul_s)
-
-# Incongruent
-# cor_pcs_IES_mul_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
-# 
-# IES_pcs_mul_inc_post <- describe_posterior(cor_pcs_IES_mul_inc)
+r |> 
+  format_table() |> 
+  select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |> 
+  gt::gt()
 ```
-
-#### Vertical Horizontal
-```{r}
-
-cor_pcs_IES_vh_con <- BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_VerticalHorizontal_Congruent)
-
-IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_vh_con)
-
-cor_pcs_IES_vh_m <- BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_VerticalHorizontal_Mild)
-
-IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_vh_m)
-
-cor_pcs_IES_vh_s <-  BayesFactor::correlationBF(
-  y = data_ppt$pcs, 
-  x = data_vs_wide$IES_VerticalHorizontal_Strong)
-
-IES_pcs_vh_str_post <- describe_posterior(cor_pcs_IES_vh_s)
-
-# Incongruent
-# cor_pcs_IES_vh_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
-# 
-# IES_pcs_vh_inc_post <- describe_posterior(cor_pcs_IES_vh_inc)
-```
-
-## Correlations between Illusion scores and Psychoticism
-
-#### Ebbinghaus 
-```{r}
-
-cor_pcs_psy_ebb_con <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_Ebbinghaus_Congruent)
-
-IES_psy_ebb_con_post <- describe_posterior(cor_pcs_psy_ebb_con)
-
-cor_pcs_psy_ebb_m <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_Ebbinghaus_Mild)
-
-IES_psy_ebb_mild_post <- describe_posterior(cor_pcs_psy_ebb_m)
-
-cor_pcs_psy_ebb_s <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_Ebbinghaus_Strong)
-
-IES_psy_ebb_str_post <- describe_posterior(cor_pcs_psy_ebb_s)
-
-# Incongruent
-# cor_pcs_psy_ebb_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$IES_Ebbinghaus_Incongruent)
-# 
-# IES_psy_ebb_inc_post <- describe_posterior(cor_pcs_psy_ebb_inc)
-```
-
-#### Muller-Lyer 
-```{r}
-
-cor_pcs_psy_mul_con <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_MullerLyer_Congruent)
-
-IES_psy_mul_con_post <- describe_posterior(cor_pcs_psy_mul_con)
-
-cor_pcs_psy_mul_m <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_MullerLyer_Mild)
-
-IES_psy_mul_mild_post <- describe_posterior(cor_pcs_psy_mul_m)
-
-cor_pcs_psy_mul_s <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_MullerLyer_Strong)
-
-IES_psy_mul_str_post <- describe_posterior(cor_pcs_psy_mul_s)
-
-# Incongruent
-# cor_pcs_psy_mul_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$IES_MullerLyer_Incongruent)
-# 
-# IES_psy_mul_inc_post <- describe_posterior(cor_pcs_psy_mul_inc)
-
-```
-
-#### Vertical-Horizontal
-```{r}
-
-cor_pcs_psy_vh_con <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_VerticalHorizontal_Congruent)
-
-IES_psy_vh_con_post <- describe_posterior(cor_pcs_psy_vh_con)
-
-cor_pcs_psy_vh_m <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_VerticalHorizontal_Mild)
-
-IES_psy_vh_mild_post <- describe_posterior(cor_pcs_psy_vh_m)
-
-cor_pcs_psy_vh_s <- BayesFactor::correlationBF(
-  y = data_ppt$PID5_Psychoticism,
-  x = data_vs_wide$IES_VerticalHorizontal_Strong)
-
-IES_psy_vh_str_post <- describe_posterior(cor_pcs_psy_vh_s)
-
-# Incongruent
-# cor_pcs_psy_vh_inc <-  BayesFactor::correlationBF(
-#   y = data_ppt$pcs,
-#   x = data_vs_wide$IES_VerticalHorizontal_Incongruent)
-# 
-# IES_psy_vh_inc_post <- describe_posterior(cor_pcs_psy_vh_inc)
-
-```

### 2025-05-21 paper/R1/manuscript.qmd
updated manuscript

@@
 
-**The phenomenological control task will include several manipulation check indices to identify problematic participants.** 
+**The phenomenological control scale will include several attention checks to identify problematic participants.** 
 **The task consists of various auditory and visual exercises; at the outset, participants hear a voice say “hello” and are asked to select the corresponding phrase from multiple options (e.g., “Hello,” “Goodbye,” “How are you,” “Thank you”).**
@@
 **Pressing it fewer than five times within the allotted time indicates a failure to follow the auditory instructions.** 
-**Participants who fail any one of these manipulation checks will be excluded from further analysis.** 
-**Finally, reliability of the PCS will be assessed by computing Cronbach's alpha [@cronbach1951coefficient].**
+**Participants who fail any one of these will be excluded from further analysis.**
+**Finally, the reliability of the PCS will be assessed by computing Cronbach's alpha of its items [@cronbach1951coefficient].**
+
+Outliers in the Illusion Game will be identified based on participants’ performance. 
+**Specifically, any participant exhibiting an error rate greater than 45% for any illusion type will be considered to be responding at chance level (i.e., randomly) and flagged accordingly**. 
+**If this level of performance occurs in the first block, the entire participant will be excluded from analysis, as it suggests a failure to properly engage with the task.** 
+**However, if the elevated error rate is observed only in the second block, this will be interpreted as a loss of engagement or motivation (e.g., boredom due to task repetition).** 
+**In such cases, only the second block will be discarded, allowing for estimation of illusion sensitivity based on the valid data from the first block, albeit with reduced precision.** 
+**To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average [RT > 4 SD above the mean, based on @makowski2023] will be excluded from the analysis.**
+After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed.
 
 **To assess whether the illusions functioned as expected, stimuli will be categorized into three groups: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.** 
-**Two Bayesian t-tests will be conducted to assess differences in IES between the Congruent and Mild conditions, and between the Mild and Strong conditions.**
-**The IES is calculated by dividing the average correct RTs by the proportion of correct responses.**
-**Significant differences in these comparisons will provide evidence that the illusions functioned as intended.**
+**The two outcome measures—error rate and IES—will be computed separately for each illusion and each illusion strength group.** 
+**To evaluate differences between these conditions, two Bayesian t-tests will be conducted: one comparing the Congruent and Mild conditions, and the other comparing the Mild and Strong conditions.** 
+**Significant differences in either IES or error rate across these comparisons will be taken as evidence that the illusions operated as intended.**
 
-**The two outcome measures—error rate and IES—will be computed for each illusion and for each illusion strength group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent.**
-**Correlation will be computed between the mild and strong groups for each illusion and outcomes separately.**
-**If these correlations are high [*r* > .50, @cohen2013statistical], the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.**
+**Next, to determine whether the Mild and Strong Illusion Strength groups can be collapsed for further analysis, Bayesian correlations will be computed between them for each illusion and outcome measure.**
+**If these correlations are sufficiently high [r > .50, @cohen2013statistical], the groups will be collapsed and outcomes recomputed across all relevant trials  (i.e., by averaging across these groups).**
+**If not, the groups will be analysed separately.** 
+**This step is necessary because reaction time may not have a linear relationship with illusion strength [@makowski2023], and collapsing the data without this check may obscure meaningful differences.**
+**Finally, reliability analyses will be conducted on all resulting indices, with Cronbach’s alpha used to evaluate internal consistency across the three illusion types.**
 
-**Reliability analyses will then be conducted on all resulting indices.** **First, split-half reliability, to assess internal consistency, will be computed by correlating two equal subsets of individual scores, with high correlations expected (*r* > .5).**
-**Second, inter-illusion reliability will be evaluated using Cronbach’s alpha across the three illusions.**
+**Bayesian correlations are then computed between PCS and illusion sensitivity scores - with the resulting IES and error rate indices.** 
+Following @lush2022, we expect to collect evidence against (BF10 <= 1/3) a relationship between PCS and VI sensitivity. 
+**Additionally, Bayesian correlations will be computed between maladaptive trait facets and illusion sensitivity scores.** 
+**Based on prior research [@makowski2023], we expect to find evidence (BF10 ≥ 3) supporting a relationship between the psychoticism facet of the PID-5 and illusion sensitivity.**
 
-Illusion Game outliers will be flagged based on their RT distributions, following the same procedure as in [@makowski2023]. 
-**If the RT is collapsed to the left (i.e., has \> 1/3 of ultra-fast responses - typically \< 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).**
-In addition, the removal of individual trials will also be performed [RT \< 200 ms or \> 3 SD; following @theriault2024].
-**To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average [RT > 4 SD above the mean, based on @makowski2023] will be excluded from the analysis.**
+All Bayesian analyses will be conducted using the BayesFactor package [@BayesFactor]. 
+**For correlations, a medium shifted beta prior will be applied (r-scale parameter set to  1/3), as recommended by @morey2018baysefactor, providing a balanced approach to estimating effect sizes, without placing undue weight on larger effect sizes or artificially inflating evidence for the null hypothesis.**
+**For t-tests, the ttestBF function will be used with a medium Cauchy prior on the standardised effect size (r-scale = √2⁄2), corresponding to the default for independent samples.**
 
-After removing problematic participants and trials, the outcome measures (PC and VI sensitivity scores) will be computed and the Bayesian correlation (with medium prior on the coefficient, i.e., r-scale parameter set to 1/3) will be computed [using the *BayesFactor* package, @BayesFactor].
-**This shifted beta prior, as recommended by @morey2018baysefactor, offers a balanced approach to estimating effect sizes, without placing undue weight on larger effect sizes or artificially inflating evidence for the null hypothesis.**
-Following @lush2022, we expect to collect evidence against (BF10 \<= 1/3) a relationship between PCS and VI sensitivity. 
-**Additionally, Bayesian correlations will be computed using the BayesFactor package, employing a medium prior on the coefficient (r-scale parameter set to 1/3) to assess relationships between maladaptive trait facets and illusion sensitivity scores.**
-**Based on prior research (Makowski et al., 2023), we expect to find evidence (BF10 ≥ 3) supporting a relationship between the psychoticism facet of the PID-5 and illusion sensitivity.**
 Data analysis will be carried out using R, using *tidyverse* [@tidyverse] and *easystats* [@easystats; @parameters; @correlation; @bayestestR; @datawizard]. 

### 2025-05-21 analysis/1_cleaning.qmd
removed paths for the review

@@
 
-# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
-
 data_ppt <- read.csv("../data/rawdata_participants.csv")
 
-# read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
 data_ig <- read.csv("../data/rawdata_illusion.csv") |>
@@
             Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
-        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
+            Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"
+        ),
         Condition_Difficulty = datawizard::categorize(
@@
             n_groups = 2,
-      labels = c("Hard", "Easy")),
+            labels = c("Hard", "Easy")
+        ),
         .by = "Illusion_Type"
@@
 data_ppt |>
-  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
+    mutate(
+        Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
-         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
+        Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)
+    ) |>
     ggplot(aes(y = Participant, x = Duration)) +
@@
     guides(color = "none", shape = "none") +
-  ggside::geom_xsidedensity(fill = "grey", color=NA) +
+    ggside::geom_xsidedensity(fill = "grey", color = NA) +
     ggside::scale_xsidey_continuous(expand = c(0, 0)) +
@@
     ggside::theme_ggside_void() +
-  theme(ggside.panel.scale = .3, 
-        axis.text.y = element_blank()) 
+    theme(
+        ggside.panel.scale = .3,
+        axis.text.y = element_blank()
+    )
 ```
@@
 data_outliers <- data_ig |>
-  summarize(p_Error = sum(Error) / n(),
+    summarize(
+        p_Error = sum(Error) / n(),
         z_RT = mean(RT),
-            .by = c("Participant", "Illusion_Type", "Block")) |> 
+        .by = c("Participant", "Illusion_Type", "Block")
+    ) |>
     mutate(z_RT = as.numeric(standardize(z_RT)), .by = c("Illusion_Type", "Block")) |>
@@
     ggplot(aes(x = p_Error)) +
-  geom_density(aes(color=Illusion_Type)) +
-  geom_vline(xintercept = 0.45, linetype = "dashed", color="red")
+    geom_density(aes(color = Illusion_Type)) +
+    geom_vline(xintercept = 0.45, linetype = "dashed", color = "red")
 
@@
     ggplot(aes(x = z_RT)) +
-  geom_density(aes(color=Illusion_Type)) +
-  geom_vline(xintercept = 4, linetype = "dashed", color="red")
+    geom_density(aes(color = Illusion_Type)) +
+    geom_vline(xintercept = 4, linetype = "dashed", color = "red")
 
@@
     ggplot(aes(x = data_ppt)) +
-  geom_histogram(aes(x = Age, fill=Gender), bins=28) +
-  geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth=1.5) +
-  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
+    geom_histogram(aes(x = Age, fill = Gender), bins = 28) +
+    geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth = 1.5) +
+    scale_fill_manual(values = c("Male" = "#64B5F6", "Female" = "#F06292", "Other" = "orange")) +
     scale_x_continuous(expand = c(0, 0)) +
@@
     geom_bar(aes(fill = Education)) +
-  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
+    scale_y_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +
     scale_fill_viridis_d(guide = "none") +

### 2025-05-21 analysis/2_analysis.qmd
removed paths for the review

@@
 library(psych)
-
-# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")
 
@@
 ig_scores <- data_ig |>
-  summarize(ErrorRate = sum(Error) / n(),
+    summarize(
+        ErrorRate = sum(Error) / n(),
         IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
-            .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")) |> 
-  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"), 
-              values_from = c("ErrorRate", "IES")) 
+        .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")
+    ) |>
+    pivot_wider(
+        names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"),
+        values_from = c("ErrorRate", "IES")
+    )
 ```
@@
 corrs <- data.frame()
-for(outcome in c("ErrorRate", "IES")) {
-  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
-    for(cond in c("Congruent", "Mild", "Strong")) {
+for (outcome in c("ErrorRate", "IES")) {
+    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
+        for (cond in c("Congruent", "Mild", "Strong")) {
             corrs <- correlation::correlation(
-        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Easy")),
-        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Hard"))) |> 
+                select(ig_scores, select = paste0(outcome, "_", ill_type, "_", cond, "_Easy")),
+                select(ig_scores, select = paste0(outcome, "_", ill_type, "_", cond, "_Hard"))
+            ) |>
                 mutate(Outcome = outcome) |>
@@
 ig_scores <- data_ig |>
-  summarize(ErrorRate = sum(Error) / n(),
+    summarize(
+        ErrorRate = sum(Error) / n(),
         IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
-            .by = c("Participant", "Illusion_Type", "Condition_Illusion")) |> 
-  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion"), 
-              values_from = c("ErrorRate", "IES")) 
+        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
+    ) |>
+    pivot_wider(
+        names_from = c("Illusion_Type", "Condition_Illusion"),
+        values_from = c("ErrorRate", "IES")
+    )
 ```
@@
 ttests <- data.frame()
-for(outcome in c("ErrorRate", "IES")) {
-  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
-    for(cond in c("Congruent", "Mild")) {
+for (outcome in c("ErrorRate", "IES")) {
+    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
+        for (cond in c("Congruent", "Mild")) {
             # Add loop for Hard and Easy if need be
             x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond)]]
-      if(cond == "Congruent") {
+            if (cond == "Congruent") {
                 x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild")]]

### 2025-05-21 analysis/preprocessing.R

@@ -1,7 +1,7 @@
 library(jsonlite)
 library(progress)
 
-path <- "C:/Users/asf25/Box/IllusionPCS/"
+path <- "path"
 
 # JsPsych experiment ------------------------------------------------------
 
@@ -16,11 +16,11 @@ alldata_ig <- data.frame()
 for (file in files) {
   progbar$tick()
   rawdata <- read.csv(paste0(path, "/", file))
-  
-  
+
+
   # Initialize participant-level data
   dat <- rawdata[rawdata$screen == "browser_info", ]
-  
+
   data_ppt <- data.frame(
     Participant = sub("\\.csv$", "", file),
     Recruitment = dat$researcher,
@@ -33,37 +33,37 @@ for (file in files) {
     Screen_Width = dat$screen_width,
     Screen_Height = dat$screen_height
   )
-  
-  
-  if("prolific_id" %in% colnames(dat)){
+
+
+  if ("prolific_id" %in% colnames(dat)) {
     data_ppt$Prolific_ID <- dat$prolific_id
   }
-  
+
   # Demographics
   demog <- jsonlite::fromJSON(rawdata[rawdata$screen == "demographic_questions", ]$response)
-  
-  #Education
+
+  # Education
   demog$Education <- ifelse(demog$Education == "other", demog$`Education-Comment`, demog$Education)
   demog$`Education-Comment` <- NULL
-  
-  #Discipline
+
+  # Discipline
   demog$Discipline <- ifelse(!is.null(demog$Discipline), demog$Discipline, NA)
   demog$Discipline <- ifelse(demog$Discipline == "other", demog$`Discipline-Comment`, demog$Discipline)
   demog$`Discipline-Comment` <- NULL
-  
-  #Student
+
+  # Student
   demog$Student <- ifelse(!is.null(demog$Student), demog$Student, NA)
-  
-  #ethnicity
+
+  # ethnicity
   demog$Ethnicity <- ifelse(!is.null(demog$Ethnicity), demog$Ethnicity, NA)
   demog$Ethnicity <- ifelse(demog$Ethnicity == "other", demog$`Ethnicity-Comment`, demog$Ethnicity)
   demog$`Ethnicity-Comment` <- NULL
-  
+
   demog <- as.data.frame(demog)
   data_ppt <- cbind(data_ppt, demog)
-  
+
[…]
